---
title: "Prepare chla data and vertically align to same depths as (pH) climatology"
author: "David Stappard & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r set_options_global, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

# Task

This markdown file reads previously created pH climatology file and uses that as the definition of the depth levels that the Argo chlorophyll a data should be aligned to. Previously created BGC data (bgc_data.rds) and metadata (bgc_metadata.rds) are loaded from the BGC preprocessed folder.

Base data qc flags are checked to ensure that the float position, pressure measurements and chlorophyll a measurements are good. Pressure is used to derive the depth of each measurement. The chlorophyll a profile is checked to ensure that significant gaps (specified by the opt_gap_limit, opt_gap_min_depth and opt_gap_max_depth) do not exist. Profiles are assigned a profile_range field that identifies the depth 1 = 614 m, 2 = 1225 m and 3 = 1600 m.

The float chlorophyll a profiles are then aligned using the spline function to match the depth levels of the climatology resulting in data frame bgc_data_chla_interpolated_clean.



```{r loading_libraries, include=FALSE}

library(tidyverse)
library(argodata)
library(lubridate)
library(gridExtra)
library(gsw)

```

## Set directories

location of pre-prepared data

```{r set_updata_root_directory, include=FALSE}

path_argo <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/bgc_argo'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")

```

## Set options

Define options that are used to determine profiles that we will us in the ongoing analysis

```{r set_options}

# Options

# opt_profile_depth_range
# The profile must have at least one chla reading at a depth <= opt_profile_depth_range[1, ]
# The profile must have at least one chla reading at a depth >= opt_profile_depth_range[2, ].
# In addition if the profile depth does not exceed the min(opt_profile_depth_range[2, ]) (i.e. 600) it will be removed.
profile_range <- c(1, 2, 3)
min_depth <- c(10, 10, 10)
max_depth <- c(614, 1225, 1600)
opt_profile_depth_range <- data.frame(profile_range, min_depth, max_depth)

# opt_gap...
# The profile should not have a gap greater that opt_gap_limit within the range defined by opt_gap_min_depth and opt_gap_max_depth
opt_gap_limit <- c(28, 55, 110)
opt_gap_min_depth <- c(0, 400, 1000)
opt_gap_max_depth <- c(400, 1000, 1600)

# opt_measure_label, opt_xlim and opt_xbreaks are associated formatting
opt_measure_label <- expression("chlorophyll a ( mg m"^"-3"~")")
opt_xlim <- c(-0.5, 2.0)
opt_xbreaks <- c(-0.5, 0, 0.5, 1.0, 1.5, 2.0)

# opt_n_prof_sel
# The selection criteria that is used against n_prof, here set to 1
# Description of n_prof usage is provided at https://argo.ucsd.edu/data/data-faq/version-3-profile-files/ the next two lines are from that page.
#     The main Argo CTD profile is stored in N_PROF=1. All other parameters (including biogeochemical parameters) that are measured 
#     with the same vertical sampling scheme and at the same location and time as the main Argo CTD profile are also stored in N_PROF=1.
opt_n_prof_sel = 1

```


## read climatology

read pH climatology, values are provided at set depths

```{r read_climatology}

# climatology values (clim_ph) available for lat, lon, month and pressure
ucsd_clim <- read_rds(file = paste0(path_argo_preprocessed, '/ucsd_ph_clim.rds'))

# What is the max depth we are interested in
opt_profile_max_depth <- max(opt_profile_depth_range$max_depth)

# existing depth levels that we will align to
target_depth_levels <-   ucsd_clim %>%
  filter(depth <= opt_profile_max_depth) %>%
  rename(target_depth = depth) %>%
  distinct(target_depth)


rm(ucsd_clim)
gc()

```

## read chla data

read chla profile and carry out basic checks, good data.

```{r read_profile_chla_data}

# base data and associated metadata
bgc_data <- read_rds(file = paste0(path_argo_preprocessed, '/bgc_data.rds'))
bgc_metadata <- read_rds(file = paste0(path_argo_preprocessed, '/bgc_metadata.rds'))

# Select relevant field from metadata ready to join to bgc_data
bgc_metadata_select <- bgc_metadata %>%
  filter(position_qc == 1) %>%
  select(file_id,
         date,
         lat,
         lon) %>%
  mutate(year = year(date),
         month = month(date),
         .after = date)

# we drive alignment from pressure and chla data
# conditions 
# n_prof == 1
# pres_adjusted_qc %in% c(1, 8) - pressure data marked as good
# chla_adjusted_qc %in% c(1, 8) - chla data marked as good
# !is.na(pres_adjusted) - pressure value must be present
# !is.na(chla_adjusted) - chla value must be present
bgc_data_chla <- bgc_data %>%
  filter(
    pres_adjusted_qc %in% c(1, 8) &
      chla_adjusted_qc %in% c(1, 8) &
      n_prof == opt_n_prof_sel &
      !is.na(pres_adjusted) &
      !is.na(chla_adjusted)
  ) %>%
  select(file_id,
         pres_adjusted,
         chla_adjusted)

# join with metadata information and calculate depth field
bgc_data_chla <- inner_join(bgc_metadata_select %>% select(file_id, lat),
                          bgc_data_chla) %>%
  mutate(depth = gsw_z_from_p(pres_adjusted, latitude = lat) * -1.0,
         .before = pres_adjusted) %>%
  select(-c(lat, pres_adjusted))

# ensure we have a depth, and chla_adjusted for all rows in bgc_data_chla
bgc_data_chla <- bgc_data_chla %>%
                  filter(!is.na(depth) & !is.na(chla_adjusted))

# clean up working tables
rm(bgc_data, bgc_metadata)
gc()

```

## Profile limits

Apply the rules that are determined by options set in set_options.
Profile must cover a set range and not contain gaps.

```{r profile_limits_chla}

# Determine profile min and max depths
bgc_profile_limits <- bgc_data_chla %>%
  group_by(file_id) %>%
  summarise(
    min_depth = min(depth),
    max_depth = max(depth),
  ) %>%
  ungroup()

# The profile much match at least one of the range criteria
force_min <- min(opt_profile_depth_range$min_depth)
force_max <- min(opt_profile_depth_range$max_depth)

# Apply profile min and max restrictions
bgc_apply_limits <- bgc_profile_limits %>%
  filter(
    min_depth <= force_min &
    max_depth >= force_max
    )

# Ensure working data set only contains profiles that have confrormed to the range test
bgc_data_chla <- right_join(bgc_data_chla,
                          bgc_apply_limits %>% select(file_id))


# Add profile type field and set all to 1.  
# All profile that meet the minimum requirements are profile_range = 1
bgc_data_chla <- bgc_data_chla %>%
  mutate(profile_range = 1)

for (i in 2:nrow(opt_profile_depth_range)) {

  #i = 3
  range_min <- opt_profile_depth_range[i,'min_depth']
  range_max <- opt_profile_depth_range[i,'max_depth']

  # Apply profile min and max restrictions
  bgc_apply_limits <- bgc_profile_limits %>%
    filter(min_depth <= range_min &
             max_depth >= range_max) %>%
    select(file_id) %>%
    mutate (profile_range = i)

  # Update profile range to i for these profiles
  # bgc_data_temp <- full_join(bgc_data_temp, bgc_apply_limits) %>%
  #                         filter(!is.na(min_depth))
  bgc_data_chla <-
    bgc_data_chla %>% rows_update(bgc_apply_limits, by = "file_id")
  
}

# Find the gaps within the profiles
profile_gaps <- full_join(bgc_data_chla,
                          opt_profile_depth_range) %>%
  filter(depth >= min_depth & 
           depth <= max_depth) %>% 
  select(file_id,
         depth) %>%
  arrange(file_id, depth) %>%
  group_by(file_id) %>%
  mutate(gap = depth - lag(depth, default = 0)) %>%
  ungroup()

# Ensure we do not have gaps in the profile that invalidate it 
for (i_gap in opt_gap_limit) {

  # The limits to be applied in that pass of for loop
  # i_gap <- opt_gap_limit[3]
  i_gap_min = opt_gap_min_depth[which(opt_gap_limit == i_gap)]
  i_gap_max = opt_gap_max_depth[which(opt_gap_limit == i_gap)]
  
  # Which gaps are greater than i_gap
  profile_gaps_remove <- profile_gaps %>%
    filter(gap > i_gap) %>%
    filter(depth >= i_gap_min & depth <= i_gap_max) %>%
    distinct(file_id) %>% 
    pull()
  
  # Remonve these profiles from working data set
  bgc_data_chla <- bgc_data_chla %>% 
    filter(!file_id %in% profile_gaps_remove)

}

# clean up working tables
rm(bgc_profile_limits, profile_gaps, profile_gaps_remove, bgc_apply_limits)
gc()

```

## Vertical align chla

We have a set of chla profiles that match our criteria we now need to align that data set to match the 
depth that are in target_depth_range, this will match the range of climatology values in ucsd_clim

```{r vertical_align_chla}

# create unique combinations of file_id and profile ranges
profile_range_file_id <- 
  bgc_data_chla %>% 
  distinct(file_id, profile_range)


# select variable of interest and prepare target_depth field
bgc_data_chla_clean <- bgc_data_chla %>%
  select(-profile_range) %>%
  mutate(target_depth = depth, .after = depth)

rm(bgc_data_chla)
gc()

# create all possible combinations of location, month and depth levels for interpolation
target_depth_grid <-
  expand_grid(
    target_depth_levels,
    profile_range_file_id
  )

# Constrain target_depth_grid to profile depth range
target_depth_grid <-
  left_join(target_depth_grid, opt_profile_depth_range) %>%
  filter(target_depth <= max_depth)

target_depth_grid <- target_depth_grid %>%
  select(target_depth,
         file_id)

# extend chla depth vectors with target depths
bgc_data_chla_extended <-
  full_join(bgc_data_chla_clean, target_depth_grid) %>%
  arrange(file_id, target_depth)

rm(bgc_data_chla_clean)
gc()

# predict spline interpolation on adjusted depth grid for chla location and month
bgc_data_chla_interpolated <-
  bgc_data_chla_extended %>%
  group_by(file_id) %>%
  mutate(chla_spline = spline(target_depth, chla_adjusted,
                                method = "natural",
                                xout = target_depth)$y) %>%
  ungroup()

rm(bgc_data_chla_extended)
gc()

# subset interpolated values on target depth range
bgc_data_chla_interpolated_clean <- 
  inner_join(target_depth_levels, bgc_data_chla_interpolated)

rm(bgc_data_chla_interpolated)
gc()

# select columns and rename to initial names
bgc_data_chla_interpolated_clean <-
  bgc_data_chla_interpolated_clean %>%
  select(file_id,
         depth = target_depth,
         chla = chla_spline)

# merge with profile range
bgc_data_chla_interpolated_clean <-
  full_join(bgc_data_chla_interpolated_clean,
            profile_range_file_id)

# merge with meta data
bgc_data_chla_interpolated_clean <-
  left_join(bgc_data_chla_interpolated_clean,
            bgc_metadata_select)

```


## Write files

Write the interpolated chla profiles that map onto depth levels.
```{r write_profile_chla_data}

# Write files
bgc_data_chla_interpolated_clean %>%
  write_rds(file = paste0(path_argo_preprocessed, "/chla_bgc_va.rds"))

# Rename so that names match if just reading existing files
chla_bgc_va <- bgc_data_chla_interpolated_clean

rm(bgc_data_chla_interpolated_clean)
gc()

```

## read files

Read files that were previously created ready for analysis
```{r read_chla_va}

# read files
chla_bgc_va <- read_rds(file = paste0(path_argo_preprocessed, "/chla_bgc_va.rds"))

```

## Analysis

```{r set_global_theme, include=FALSE}

theme_set(theme_bw())

```

## chla mean profile

```{r plot_overall_mean_chla_profiles, fig.asp=1}

max_depth_1 <- opt_profile_depth_range[1, "max_depth"]
max_depth_2 <- opt_profile_depth_range[2, "max_depth"]
max_depth_3 <- opt_profile_depth_range[3, "max_depth"]

# Profiles to 600m
chla_overall_mean_1 <- chla_bgc_va %>% 
  filter(profile_range %in% c(1, 2, 3) & depth >= 10 & depth <= max_depth_1) %>%
  group_by(depth) %>% 
  summarise(count_measures = n(),
            chla_mean = mean(chla, na.rm = TRUE),
            chla_sd = sd(chla, na.rm = TRUE))

chla_year_mean_1 <- chla_bgc_va %>% 
  filter(profile_range %in% c(1, 2, 3) & depth >= 10 & depth <= max_depth_1) %>%
  group_by(year, depth) %>% 
  summarise(count_measures = n(),
            chla_mean = mean(chla, na.rm = TRUE),
            chla_sd = sd(chla, na.rm = TRUE))

# Profiles to 1200m
chla_overall_mean_2 <- chla_bgc_va %>% 
  filter(profile_range %in% c(2, 3) & depth >= 10 & depth <= max_depth_2) %>%
  group_by(depth) %>% 
  summarise(count_measures = n(),
            chla_mean = mean(chla, na.rm = TRUE),
            chla_sd = sd(chla, na.rm = TRUE))

chla_year_mean_2 <- chla_bgc_va %>% 
  filter(profile_range %in% c(2, 3) & depth >= 10 & depth <= max_depth_2) %>%
  group_by(year, depth) %>% 
  summarise(count_measures = n(),
            chla_mean = mean(chla, na.rm = TRUE),
            chla_sd = sd(chla, na.rm = TRUE))

# Profiles to 1500m
chla_overall_mean_3 <- chla_bgc_va %>% 
  filter(profile_range %in% c(3) & depth >= 10 & depth <= max_depth_3) %>%
  group_by(depth) %>% 
  summarise(count_measures = n(),
            chla_mean = mean(chla, na.rm = TRUE),
            chla_sd = sd(chla, na.rm = TRUE))

chla_year_mean_3 <- chla_bgc_va %>% 
  filter(profile_range %in% c(3) & depth >= 10 & depth <= max_depth_3) %>%
  group_by(year, depth) %>% 
  summarise(count_measures = n(),
            chla_mean = mean(chla, na.rm = TRUE),
            chla_sd = sd(chla, na.rm = TRUE))

# All years
chla_overall_mean_1 %>% 
  ggplot()+
  geom_path(aes(x = chla_mean,
                y = depth))+
  geom_ribbon(aes(
    xmax = chla_mean + chla_sd,
    xmin = chla_mean - chla_sd,
    y = depth
  ),
  alpha = 0.2) +
  scale_y_reverse()+
  coord_cartesian(xlim = opt_xlim)+
  scale_x_continuous(breaks = opt_xbreaks)+
  labs(
    title = paste0('Overall mean chlorophyll a to ', max_depth_1, 'm'), 
    x=opt_measure_label, 
    y='depth (m)'
  )

chla_overall_mean_2 %>% 
  ggplot()+
  geom_path(aes(x = chla_mean,
                y = depth))+
  geom_ribbon(aes(
    xmax = chla_mean + chla_sd,
    xmin = chla_mean - chla_sd,
    y = depth
  ),
  alpha = 0.2) +
  scale_y_reverse()+
  coord_cartesian(xlim = opt_xlim)+
  scale_x_continuous(breaks = opt_xbreaks)+
  labs(
    title = paste0('Overall mean chlorophyll a to ', max_depth_2, 'm'), 
    x=opt_measure_label, 
    y='depth (m)'
  )

chla_overall_mean_3 %>% 
  ggplot()+
  geom_path(aes(x = chla_mean,
                y = depth))+
  geom_ribbon(aes(
    xmax = chla_mean + chla_sd,
    xmin = chla_mean - chla_sd,
    y = depth
  ),
  alpha = 0.2) +
  scale_y_reverse()+
  coord_cartesian(xlim = opt_xlim)+
  scale_x_continuous(breaks = opt_xbreaks)+
  labs(
    title = paste0('Overall mean chlorophyll a to ', max_depth_3, 'm'), 
    x=opt_measure_label, 
    y='depth (m)'
  )

# by years
chla_year_mean_1 %>% 
  ggplot()+
  geom_path(aes(x = chla_mean,
                y = depth))+
  geom_ribbon(aes(
    xmax = chla_mean + chla_sd,
    xmin = chla_mean - chla_sd,
    y = depth
  ),
  alpha = 0.2) +
  scale_y_reverse()+
  facet_wrap(~year)+
  coord_cartesian(xlim = opt_xlim)+
  scale_x_continuous(breaks = opt_xbreaks)+
  labs(
    title = paste0('Yearly overall mean chlorophyll a to ', max_depth_1, 'm'),
    x = opt_measure_label,
    y = 'depth (m)'
  )

chla_year_mean_2 %>% 
  ggplot()+
  geom_path(aes(x = chla_mean,
                y = depth))+
  geom_ribbon(aes(
    xmax = chla_mean + chla_sd,
    xmin = chla_mean - chla_sd,
    y = depth
  ),
  alpha = 0.2) +
  scale_y_reverse()+
  facet_wrap(~year)+
  coord_cartesian(xlim = opt_xlim)+
  scale_x_continuous(breaks = opt_xbreaks)+
  labs(
    title = paste0('Yearly overall mean chlorophyll a to ', max_depth_2, 'm'),
    x = opt_measure_label,
    y = 'depth (m)'
  )

chla_year_mean_3 %>% 
  ggplot()+
  geom_path(aes(x = chla_mean,
                y = depth))+
  geom_ribbon(aes(
    xmax = chla_mean + chla_sd,
    xmin = chla_mean - chla_sd,
    y = depth
  ),
  alpha = 0.2) +
  scale_y_reverse()+
  facet_wrap(~year)+
  coord_cartesian(xlim = opt_xlim)+
  scale_x_continuous(breaks = opt_xbreaks)+
  labs(
    title = paste0('Yearly overall mean chlorophyll a to ', max_depth_3, 'm'),
    x = opt_measure_label,
    y = 'depth (m)'
  )

```

## Profile counts

Details of the number of profiles and to which depths over the analysis period
```{r chla_histogram, fig.asp=1}

chla_histogram <- chla_bgc_va %>%
  group_by(year, profile_range = as.character(profile_range)) %>%
  summarise(num_profiles = n_distinct(file_id)) %>%
  ungroup()

chla_histogram %>%
  ggplot() +
  geom_bar(
    aes(
      x = year,
      y = num_profiles,
      fill = profile_range,
      group = profile_range
    ),
    position = "stack",
    stat = "identity"
  ) +
  scale_fill_viridis_d() +
  labs(title = "chlorophyll a profiles per year and profile range",
       x = "year",
       y = "profile count",
       fill = "profile range")

```
