---
title: "Prepare pH data and vertically align to climatology"
author: "David Stappard & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r set_options_global, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

# Task

Load existing bgc_data.rds file and align profiles to a common vertical pattern.

```{r loading_libraries, include=FALSE}

library(tidyverse)
library(argodata)
library(lubridate)
library(gridExtra)
library(gsw)

```

## Set directories

location of pre-prepared data

```{r set_updata_root_directory, include=FALSE}

path_argo <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/bgc_argo'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")

```

## Set options

Define options that are used to determine profiles that we will us in the ongoing analysis

```{r set_options}

# Options

# opt_profile_depth_range
# The profile must have at least one pH reading at a depth <= opt_profile_depth_range[1, ]
# The profile must have at least one pH reading at a depth >= opt_profile_depth_range[2, ].
# In addition if the profile depth does not exceed the min(opt_profile_depth_range[2, ]) (i.e. 600) it will be removed.
profile_range <- c(1, 2, 3)
min_depth <- c(10, 10, 10)
max_depth <- c(614, 1225, 1600)
opt_profile_depth_range <- data.frame(profile_range, min_depth, max_depth)

# opt_gap...
# The profile should not have a gap greater that opt_gap_limit within the range defined by opt_gap_min_depth and opt_gap_max_depth
opt_gap_limit <- c(28, 55, 110)
opt_gap_min_depth <- c(0, 400, 1000)
opt_gap_max_depth <- c(400, 1000, 1600)

# opt_measure
# opt_measure = 1 analysis is done using pH
# opt_measure = 2 analysis is done using h_plus
# opt_measure_label, opt_xlim and opt_xbreaks are associated formatting
opt_measure <- 2
if (opt_measure == 1){
  opt_measure_label <- "pH anomaly"
  opt_xlim <- c(-0.08, 0.08)
  opt_xbreaks <- c(-0.08, -0.04, 0, 0.04, 0.08)
} else {
  opt_measure_label <- expression("[H]"^"+" ~ "anomaly")
  opt_xlim <- c(-2e-9, 2e-9)
  opt_xbreaks <- c(-2e-9, -1e-9, 0, 1e-9, 2e-9)
}

```


## read climatology

read pH climatology, values are provided at set depths

```{r read_climatology}

# climatology values (clim_ph) available for lat, lon, month and pressure
ucsd_clim <- read_rds(file = paste0(path_argo_preprocessed, '/ucsd_ph_clim.rds'))

# Select fields of interest
ucsd_clim <- ucsd_clim %>% 
  select(
    month,
    lat,
    lon,
    depth,
    clim_pH
  ) %>%
  mutate(
    clim_h_plus = 10^-clim_pH
  )

#What is the max depth we are interested in
opt_profile_max_depth <- max(opt_profile_depth_range$max_depth)

# existing depth levels that we will align to
target_depth_range <-   distinct(ucsd_clim %>%
                              filter(depth <= opt_profile_max_depth) %>% 
                              mutate(target_depth = depth) %>%
                              select(target_depth))

```

## read pH data

read pH profile and carry out basic checks, good data.

```{r read_profile_pH_data}

# base data and associated metadata
bgc_data <- read_rds(file = paste0(path_argo_preprocessed, '/bgc_data.rds'))
bgc_metadata <- read_rds(file = paste0(path_argo_preprocessed, '/bgc_metadata.rds'))

# Select relevant field from metadata ready to join to bgc_data
bgc_metadata_select <- bgc_metadata %>%
                            filter (position_qc == 1) %>%
                            select(
                              file_id,
                              date,
                              lat,
                              lon
                              ) %>%
                            mutate(year = year(date), month = month(date), .after = date)

# we drive alignment from pressure and pH data
# conditions 
# n_prof == 1
# pres_adjusted_qc %in% c(1, 8) - pressure data marked as good
# ph_in_situ_total_adjusted_qc %in% c(1, 8) - pH data marked as good
# !is.na(pres_adjusted) - pressure value must be present
# !is.na(ph_in_situ_total_adjusted) - pH value must be present
bgc_data_pH <- bgc_data %>%
                  filter(pres_adjusted_qc %in% c(1, 8) & 
                           ph_in_situ_total_adjusted_qc %in% c(1, 8) & 
                           n_prof == 1 & !is.na(pres_adjusted) &
                           !is.na(ph_in_situ_total_adjusted)) %>%
                  select(
                    file_id,
                    pres_adjusted,
                    ph_in_situ_total_adjusted
                  )

# join with metadata information and calculate depth field
bgc_data_pH <- full_join(bgc_metadata_select, bgc_data_pH) %>%
                          select(
                            file_id,
                            lat, 
                            lon, 
                            date,
                            year, 
                            month,
                            pres_adjusted,
                            ph_in_situ_total_adjusted
                          ) %>%
                          mutate(depth = gsw_z_from_p(pres_adjusted, latitude = lat) * -1.0,
                                 .before = pres_adjusted) %>%
                          filter(!is.na(file_id))

# ensure we have a depth, pres_adjusted and ph_in_situ_total_adjusted for all rows in bgc_data_pH
bgc_data_pH <- bgc_data_pH %>%
                  filter(!is.na(depth) & !is.na(pres_adjusted) & !is.na(ph_in_situ_total_adjusted))

# clean up working tables
rm(bgc_data, bgc_metadata_select)
```

## Profile limits

Apply the rules that are determined by options set in set_options.
Profile must cover a set range and not contain gaps.

```{r profile_limits_pH}

# Determine profile min and max depths
bgc_profile_limits <- bgc_data_pH %>%
  group_by(file_id) %>%
  summarise(
    min_depth = min(depth),
    max_depth = max(depth),
  ) %>%
  ungroup()

# The profile much match at least one of teh range criteria
force_min <- min(opt_profile_depth_range$min_depth)
force_max <- min(opt_profile_depth_range$max_depth)

# Apply profile min and max restrictions
bgc_apply_limits <- bgc_profile_limits %>%
  filter(
    min_depth <= force_min &
    max_depth >= force_max
    )

# Ensure working data set only contains profiles that have confrormed to the range test
bgc_data_pH <- full_join(bgc_data_pH, bgc_apply_limits) %>%
                        filter(!is.na(min_depth))

# Add profile type field and set all to 1.  
# All profile that meet the minimum requirements are profile_range = 1
bgc_data_pH <- bgc_data_pH %>%
  mutate(profile_range = 1)

for (i in 2:nrow(opt_profile_depth_range)) {

  #i = 3
  range_min <- opt_profile_depth_range[i,'min_depth']
  range_max <- opt_profile_depth_range[i,'max_depth']

  # Apply profile min and max restrictions
  bgc_apply_limits <- bgc_profile_limits %>%
    filter(
      min_depth <= range_min &
      max_depth >= range_max
      ) %>%
    select(
      file_id
      ) %>%
    mutate (
      profile_range = i 
    )

  # Update profile range to i for these profiles
  # bgc_data_temp <- full_join(bgc_data_temp, bgc_apply_limits) %>%
  #                         filter(!is.na(min_depth))
  bgc_data_pH <- bgc_data_pH %>% rows_update(bgc_apply_limits, by = "file_id")

}

# Find the gaps within the profiles
profile_gaps <- bgc_data_pH %>%
  select(
    file_id,
    depth
    ) %>%
  arrange(file_id, depth) %>%
  group_by(file_id) %>%
  mutate(gap = depth - lag(depth, default = 0)) %>%
  ungroup()

# Ensure we do not have gaps in the profile that invalidate it 
for (i_gap in opt_gap_limit) {

  # The limits to be applied in that pass of for loop
  # i_gap <- opt_gap_limit[3]
  i_gap_min = opt_gap_min_depth[which(opt_gap_limit == i_gap)]
  i_gap_max = opt_gap_max_depth[which(opt_gap_limit == i_gap)]
  
  # Which gaps are greater than i_gap
  profile_gaps_remove <- profile_gaps %>%
    filter(gap > i_gap) %>%
    filter(depth >= i_gap_min & depth <= i_gap_max) %>%
    select(file_id, gap)
  
  # Remonve these profiles from working data set
  bgc_data_pH <- full_join(bgc_data_pH, profile_gaps_remove) %>%
                                  filter(is.na(gap)) %>%
                                  select(-c(gap))
}

# Ensure primary fields are all present
bgc_data_pH <- bgc_data_pH %>%
  filter(!is.na(lat) & !is.na(lon) & !is.na(date) & !is.na(depth)) 

# clean up working tables
rm(bgc_profile_limits, profile_gaps, profile_gaps_remove, bgc_apply_limits)

```

## Vertical align pH

We have a set of pH profiles that match our criteria we now need to align that data set to match the 
depth that are in target_depth_range, this will match the range of climatology values in ucsd_clim

```{r vertical_align_pH}

# select variable of interest and prepare target_depth field
bgc_data_pH_clean <- bgc_data_pH %>%
                            select(
                              file_id, 
                              lat,
                              lon,
                              date,
                              depth,
                              profile_range,
                              pH = ph_in_situ_total_adjusted
                            ) %>%
                            mutate(
                              target_depth = depth, .after = depth
                            )

# create all possible combinations of location, month and depth levels for interpolation
target_depth_grid <-
  expand_grid(target_depth_range,
              bgc_data_pH_clean %>% distinct(file_id, lat, lon, date, profile_range))

# Take account of the profile_range field to determine the profile_range of target_depth_grid
target_depth_grid <- left_join(target_depth_grid, opt_profile_depth_range) %>%
        filter(target_depth <= max_depth)

target_depth_grid <- target_depth_grid %>% 
                        select(
                          target_depth, 
                          file_id, 
                          lat, 
                          lon,
                          date,
                          profile_range
                          )

# extend pH depth vectors with target depths
bgc_data_pH_extended <- full_join(bgc_data_pH_clean, target_depth_grid) %>% 
  arrange(file_id, lat, lon, date, target_depth)

# predict spline interpolation on adjusted depth grid for pH location and month
bgc_data_pH_interpolated <-
  bgc_data_pH_extended %>%
  group_by(file_id, lat, lon, date) %>%
  mutate(pH_spline = spline(target_depth, pH,
                                method = "natural",
                                xout = target_depth)$y) %>%
  ungroup()

# subset interpolated values on target depth range
bgc_data_pH_interpolated_clean <- 
  inner_join(target_depth_grid, bgc_data_pH_interpolated)

# select columns and rename to initial names
bgc_data_pH_interpolated_clean <-
  bgc_data_pH_interpolated_clean %>%
  select(
    file_id, 
    lat, 
    lon, 
    date,
    profile_range,
    depth = target_depth,
    pH = pH_spline)

# Create year and month fields
bgc_data_pH_interpolated_clean <- bgc_data_pH_interpolated_clean %>%
  mutate(year = year(date),
         month = month(date),
         .after = date) %>%
  mutate(
    h_plus = 10^-pH
  )

# clean up working tables
rm(bgc_data_pH_extended, bgc_data_pH_interpolated)

```

## Create anomaly profiles

Create anomaly profiles as observed - climatology
```{r pH_anomaly}

# Create bgc_temp_anomaly, but only where we have an climatology temperature
bgc_pH_anomaly <- left_join(bgc_data_pH_interpolated_clean, ucsd_clim) %>%
                        filter(!is.na(clim_pH) & lat <= -30)

# Calculate the anomaly temperature
bgc_pH_anomaly <- bgc_pH_anomaly %>%
                      mutate(
                        anomaly_pH = pH - clim_pH,
                        anomaly_h_plus = h_plus - clim_h_plus
                      )

# -----------------------------------------------------------------------------
# Climatology check
# -----------------------------------------------------------------------------
# It is possible even though we have observational measures to full depth the climatology may not matach all depth.
# These profiles will be removed from both temp_bgc_va and temp_anomaly_va

# Anomaly max depth
bgc_pH_anomaly_depth_check <- bgc_pH_anomaly %>%
                                group_by(
                                  file_id,
                                  profile_range
                                  ) %>%
                                summarise(
                                  min_pdepth = min(depth),
                                  max_pdepth = max(depth)
                                  ) %>%
                                ungroup()

# Add the required min depth and max depth
bgc_pH_anomaly_depth_check <- left_join(bgc_pH_anomaly_depth_check, opt_profile_depth_range)

# This profiles do not match the depth required by the profile_range
min_depth_check <- min(target_depth_grid$target_depth)
remove_profiles <- bgc_pH_anomaly_depth_check %>%
  filter((max_pdepth < max_depth) | (min_pdepth > min_depth_check)) %>%
  select(file_id, max_depth)

# remove from both bgc_data_pH_interpolated_clean and bgc_pH_anomaly
bgc_data_pH_interpolated_clean <- full_join(bgc_data_pH_interpolated_clean, remove_profiles) %>%
            filter(is.na(max_depth)) %>%
            select(-c(max_depth))

bgc_pH_anomaly <- full_join(bgc_pH_anomaly, remove_profiles) %>%
            filter(is.na(max_depth)) %>%
            select(-c(max_depth))

# clean up working tables
#rm(bgc_pH_anomaly_depth_check, remove_profiles)

```

## Write files

Write the climatology that maps onto depth levels, interpolated pH profiles that map onto depth levels and resulting anomaly files.
```{r write_profile_pH_data}

# Write files
bgc_data_pH_interpolated_clean %>%
  write_rds(file = paste0(path_argo_preprocessed, "/pH_bgc_va.rds"))

ucsd_clim %>%
  write_rds(file = paste0(path_argo_preprocessed, "/pH_clim_va.rds"))

bgc_pH_anomaly %>%
  write_rds(file = paste0(path_argo_preprocessed, "/pH_anomaly_va.rds"))

# Rename so that names match if just reading existing files
pH_bgc_va <- bgc_data_pH_interpolated_clean
pH_clim_va <- ucsd_clim
pH_anomaly_va <- bgc_pH_anomaly

rm(bgc_data_pH_interpolated_clean, ucsd_clim, bgc_pH_anomaly)


```

## read files

Read files that were previously created ready for analysis
```{r read_pH_va}

# read files
pH_bgc_va <- read_rds(file = paste0(path_argo_preprocessed, "/pH_bgc_va.rds"))

pH_clim_va <- read_rds(file = paste0(path_argo_preprocessed, "/pH_clim_va.rds"))

pH_anomaly_va <- read_rds(file = paste0(path_argo_preprocessed, "/pH_anomaly_va.rds"))

```

## Analysis

```{r set_global_theme, include=FALSE}

theme_set(theme_bw())
```

## pH anomaly

```{r plot_overall_mean_anomaly_profiles, fig.asp=1}

max_depth_1 <- opt_profile_depth_range[1, "max_depth"]
max_depth_2 <- opt_profile_depth_range[2, "max_depth"]
max_depth_3 <- opt_profile_depth_range[3, "max_depth"]

# Profiles to 600m
if (opt_measure == 1){
  anomaly_overall_mean_1 <- pH_anomaly_va %>% 
    filter(profile_range %in% c(1, 2, 3) & depth <= max_depth_1) %>%
    group_by(depth) %>% 
    summarise(count_anomaly = n(),
              anomaly_mean = mean(anomaly_pH, na.rm = TRUE),
              anomaly_sd = sd(anomaly_pH, na.rm = TRUE))
  
  anomaly_year_mean_1 <- pH_anomaly_va %>% 
    filter(profile_range %in% c(1, 2, 3) & depth <= max_depth_1) %>%
    group_by(year, depth) %>% 
    summarise(count_anomaly = n(),
              anomaly_mean = mean(anomaly_pH, na.rm = TRUE),
              anomaly_sd = sd(anomaly_pH, na.rm = TRUE))
  
  # Profiles to 1200m
  anomaly_overall_mean_2 <- pH_anomaly_va %>% 
    filter(profile_range %in% c(2, 3) & depth <= max_depth_2) %>%
    group_by(depth) %>% 
    summarise(count_anomaly = n(),
              anomaly_mean = mean(anomaly_pH, na.rm = TRUE),
              anomaly_sd = sd(anomaly_pH, na.rm = TRUE))
  
  anomaly_year_mean_2 <- pH_anomaly_va %>% 
    filter(profile_range %in% c(2, 3) & depth <= max_depth_2) %>%
    group_by(year, depth) %>% 
    summarise(count_anomaly = n(),
              anomaly_mean = mean(anomaly_pH, na.rm = TRUE),
              anomaly_sd = sd(anomaly_pH, na.rm = TRUE))
  
  # Profiles to 1500m
  anomaly_overall_mean_3 <- pH_anomaly_va %>% 
    filter(profile_range %in% c(3) & depth <= max_depth_3) %>%
    group_by(depth) %>% 
    summarise(count_anomaly = n(),
              anomaly_mean = mean(anomaly_pH, na.rm = TRUE),
              anomaly_sd = sd(anomaly_pH, na.rm = TRUE))
  
  anomaly_year_mean_3 <- pH_anomaly_va %>% 
    filter(profile_range %in% c(3) & depth <= max_depth_3) %>%
    group_by(year, depth) %>% 
    summarise(count_anomaly = n(),
              anomaly_mean = mean(anomaly_pH, na.rm = TRUE),
              anomaly_sd = sd(anomaly_pH, na.rm = TRUE))
} else {
  anomaly_overall_mean_1 <- pH_anomaly_va %>% 
    filter(profile_range %in% c(1, 2, 3) & depth <= max_depth_1) %>%
    group_by(depth) %>% 
    summarise(count_anomaly = n(),
              anomaly_mean = mean(anomaly_h_plus, na.rm = TRUE),
              anomaly_sd = sd(anomaly_h_plus, na.rm = TRUE))
  
  anomaly_year_mean_1 <- pH_anomaly_va %>% 
    filter(profile_range %in% c(1, 2, 3) & depth <= max_depth_1) %>%
    group_by(year, depth) %>% 
    summarise(count_anomaly = n(),
              anomaly_mean = mean(anomaly_h_plus, na.rm = TRUE),
              anomaly_sd = sd(anomaly_h_plus, na.rm = TRUE))
  
  # Profiles to 1200m
  anomaly_overall_mean_2 <- pH_anomaly_va %>% 
    filter(profile_range %in% c(2, 3) & depth <= max_depth_2) %>%
    group_by(depth) %>% 
    summarise(count_anomaly = n(),
              anomaly_mean = mean(anomaly_h_plus, na.rm = TRUE),
              anomaly_sd = sd(anomaly_h_plus, na.rm = TRUE))
  
  anomaly_year_mean_2 <- pH_anomaly_va %>% 
    filter(profile_range %in% c(2, 3) & depth <= max_depth_2) %>%
    group_by(year, depth) %>% 
    summarise(count_anomaly = n(),
              anomaly_mean = mean(anomaly_h_plus, na.rm = TRUE),
              anomaly_sd = sd(anomaly_h_plus, na.rm = TRUE))
  
  # Profiles to 1500m
  anomaly_overall_mean_3 <- pH_anomaly_va %>% 
    filter(profile_range %in% c(3) & depth <= max_depth_3) %>%
    group_by(depth) %>% 
    summarise(count_anomaly = n(),
              anomaly_mean = mean(anomaly_h_plus, na.rm = TRUE),
              anomaly_sd = sd(anomaly_h_plus, na.rm = TRUE))
  
  anomaly_year_mean_3 <- pH_anomaly_va %>% 
    filter(profile_range %in% c(3) & depth <= max_depth_3) %>%
    group_by(year, depth) %>% 
    summarise(count_anomaly = n(),
              anomaly_mean = mean(anomaly_h_plus, na.rm = TRUE),
              anomaly_sd = sd(anomaly_h_plus, na.rm = TRUE)) %>%
    ungroup()
}

# year_mean_3 <- anomaly_year_mean_3 %>%
#   group_by(year_num = year) %>%
#   summarise(year_mean = mean(anomaly_mean, na.rm = TRUE)) %>%
#   ungroup()
# 
# anomaly_year_mean_3 <- full_join(anomaly_year_mean_3, year_mean_3)

# All years
anomaly_overall_mean_1 %>% 
  ggplot()+
  geom_path(aes(x = anomaly_mean,
                y = depth))+
  geom_ribbon(aes(xmax = anomaly_mean + anomaly_sd,
                  xmin = anomaly_mean - anomaly_sd,
                  y = depth),
              alpha = 0.2)+
  geom_vline(xintercept = 0)+
  scale_y_reverse()+
  # scale_color_manual(values = HNL_colors)+
  # scale_fill_manual(values = HNL_colors)+
  coord_cartesian(xlim = opt_xlim)+
  scale_x_continuous(breaks = opt_xbreaks)+
  labs(title = paste0('Overall mean anomaly profiles to ', max_depth_1, 'm'), x=opt_measure_label, y='depth (m)')

anomaly_overall_mean_2 %>% 
  ggplot()+
  geom_path(aes(x = anomaly_mean,
                y = depth))+
  geom_ribbon(aes(xmax = anomaly_mean + anomaly_sd,
                  xmin = anomaly_mean - anomaly_sd,
                  y = depth),
              alpha = 0.2)+
  geom_vline(xintercept = 0)+
  scale_y_reverse()+
  # scale_color_manual(values = HNL_colors)+
  # scale_fill_manual(values = HNL_colors)+
  coord_cartesian(xlim = opt_xlim)+
  scale_x_continuous(breaks = opt_xbreaks)+
  labs(title = paste0('Overall mean anomaly profiles to ', max_depth_2, 'm'), x=opt_measure_label, y='depth (m)')

anomaly_overall_mean_3 %>% 
  ggplot()+
  geom_path(aes(x = anomaly_mean,
                y = depth))+
  geom_ribbon(aes(xmax = anomaly_mean + anomaly_sd,
                  xmin = anomaly_mean - anomaly_sd,
                  y = depth),
              alpha = 0.2)+
  geom_vline(xintercept = 0)+
  scale_y_reverse()+
  # scale_color_manual(values = HNL_colors)+
  # scale_fill_manual(values = HNL_colors)+
  coord_cartesian(xlim = opt_xlim)+
  scale_x_continuous(breaks = opt_xbreaks)+
  labs(title = paste0('Overall mean anomaly profiles to ', max_depth_3, 'm'), x=opt_measure_label, y='depth (m)')

# by years
anomaly_year_mean_1 %>% 
  ggplot()+
  geom_path(aes(x = anomaly_mean,
                y = depth))+
  geom_ribbon(aes(xmax = anomaly_mean + anomaly_sd,
                  xmin = anomaly_mean - anomaly_sd,
                  y = depth),
              alpha = 0.2)+
  geom_vline(xintercept = 0)+
  scale_y_reverse()+
  # scale_color_manual(values = HNL_colors)+
  # scale_fill_manual(values = HNL_colors)+
  facet_wrap(~year)+
  coord_cartesian(xlim = opt_xlim)+
  scale_x_continuous(breaks = opt_xbreaks)+
  labs(title = paste0('Yearly mean anomaly profiles to ', max_depth_1, 'm'), x=opt_measure_label, y='depth (m)')

anomaly_year_mean_2 %>% 
  ggplot()+
  geom_path(aes(x = anomaly_mean,
                y = depth))+
  geom_ribbon(aes(xmax = anomaly_mean + anomaly_sd,
                  xmin = anomaly_mean - anomaly_sd,
                  y = depth),
              alpha = 0.2)+
  geom_vline(xintercept = 0)+
  scale_y_reverse()+
  # scale_color_manual(values = HNL_colors)+
  # scale_fill_manual(values = HNL_colors)+
  facet_wrap(~year)+
  coord_cartesian(xlim = opt_xlim)+
  scale_x_continuous(breaks = opt_xbreaks)+
  labs(title = paste0('Yearly mean anomaly profiles to ', max_depth_2, 'm'), x=opt_measure_label, y='depth (m)')


#min_year = min(year_mean_3$year_num)

anomaly_year_mean_3 %>% 
  ggplot()+
  geom_path(aes(x = anomaly_mean,
                y = depth))+
  geom_ribbon(aes(xmax = anomaly_mean + anomaly_sd,
                  xmin = anomaly_mean - anomaly_sd,
                  y = depth),
              alpha = 0.2)+
  #geom_vline(xintercept = year_mean_3$year_mean[year[1]-2013])+
  geom_vline(xintercept = 0)+
  scale_y_reverse()+
  # scale_color_manual(values = HNL_colors)+
  # scale_fill_manual(values = HNL_colors)+
  facet_wrap(~year)+
  coord_cartesian(xlim = opt_xlim)+
  scale_x_continuous(breaks = opt_xbreaks)+
  labs(title = paste0('Yearly mean anomaly profiles to ', max_depth_3, 'm'), x=opt_measure_label, y='depth (m)')

#rm(anomaly_overall_mean)
```

## Profile counts

Details of the number of profiles and to which depths over the analysis period
```{r pH_histogram, fig.asp=1}

pH_histogram <- pH_bgc_va %>%
  group_by(year, profile_range = as.character(profile_range)) %>%
  summarise(num_profiles = n_distinct(file_id)) %>%
  ungroup()

pH_histogram %>% 
  ggplot()+
  geom_bar(aes(x = year, y = num_profiles, fill = profile_range, group = profile_range), position = "stack", stat="identity")+
  scale_fill_viridis_d()+
  labs(title = "pH profiles per year and profile range", x = "year", y = "profile count", fill = "profile range")


```
