---
title: "Prepare core temperature data and vertically align to climatology"
author: "David Stappard & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r set_options_global, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

# Task

Load existing core_data.rds file and align profiles to a common vertical pattern.

```{r loading_libraries, include=FALSE}

library(tidyverse)
library(argodata)
library(lubridate)
library(gridExtra)
library(gsw)
library(ggforce)

```

## Set directories

location of pre-prepared data

```{r set_updata_root_directory, include=FALSE}

path_argo <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/bgc_argo'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")

path_core_argo <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/core_argo_r_argodata'
path_core_preprocessed <- paste0(path_core_argo, "/preprocessed_core_data")

```

## Set options

Define options that are used to determine profiles that we will us in the ongoing analysis

```{r set_options}

# Options

# opt_profile_depth_range
# The profile must have at least one temperature reading at a depth <= opt_profile_depth_range[1, ]
# The profile must have at least one temperature reading at a depth >= opt_profile_depth_range[2, ].
# In addition if the profile depth does not exceed the min(opt_profile_depth_range[2, ]) (i.e. 600) it will be removed.
profile_range <- c(1, 2, 3)
min_depth <- c(7.5, 7.5, 7.5)
max_depth <- c(600, 1200, 1500)
opt_profile_depth_range <- data.frame(profile_range, min_depth, max_depth)

# opt_gap...
# The profile should not have a gap greater that opt_gap_limit within the range defined by opt_gap_min_depth and opt_gap_max_depth
opt_gap_limit <- c(28, 55, 110)
opt_gap_min_depth <- c(0, 400, 1000)
opt_gap_max_depth <- c(400, 1000, 1500)

# year to be refreshed are set by opt_min_year and opt_max_year
opt_min_year = 2013
opt_max_year = 2023

```


## read climatology

Temperature climatology has been prepared during BGC analysis

```{r read_climatology}

temp_clim_va <- read_rds(file = paste0(path_argo_preprocessed, "/temp_clim_va.rds"))

target_depth_range <- unique(temp_clim_va[c("depth")])
target_depth_range <- target_depth_range %>% select(target_depth = depth)

```

## Prepare years

Read annual core temperature and metadata file. Validate profile against limits and align to target_depth_range

```{r prepare_years}

for (target_year in opt_min_year:opt_max_year) {

#target_year = 2013

# --------------------------------------------------------------------------------
# Read data
# --------------------------------------------------------------------------------

# base data and associated metadata
core_data <- read_rds(file = paste0(path_core_preprocessed, "/", target_year, "_core_data_temp.rds"))
core_metadata <- read_rds(file = paste0(path_core_preprocessed, "/", target_year, "_core_metadata.rds"))

# Select relevant field from metadata ready to join to core_data
core_metadata_select <- core_metadata %>%
                            filter (position_qc == 1) %>%
                            select(
                              file,
                              date,
                              lat,
                              lon,
                              platform_number,
                              cycle_number
                              ) %>%
                            mutate(year = year(date), month = month(date), .after = date) %>%
                            unite('id', platform_number:cycle_number, sep = '_', remove = TRUE)

# we only want pressure and temperature data
# conditions 
# !is.na(depth) - pressure value must be present
# !is.na(temp_adjusted) - temperature value must be present
core_data_temp <- core_data %>%
                  filter(!is.na(depth) & !is.na(temp_adjusted)) %>%
                  select(
                    file,
                    depth,
                    temp_adjusted
                  )

# join with metadata information and calculate depth field
core_data_temp <- full_join(core_metadata_select, core_data_temp) %>%
                          select(
                            id,
                            lat, 
                            lon, 
                            date,
                            year, 
                            month,
                            depth,
                            temp = temp_adjusted
                          ) %>%
                          filter(!is.na(id))

# clean up working tables
rm(core_data, core_metadata, core_metadata_select)


# --------------------------------------------------------------------------------
# Profile limits
# --------------------------------------------------------------------------------

# Apply the rules that are determined by options set in set_options.
# Profile must cover a set range and not contain gaps.

# Determine profile min and max depths
core_profile_limits <- core_data_temp %>%
  group_by(id) %>%
  summarise(
    min_depth = min(depth),
    max_depth = max(depth),
  ) %>%
  ungroup()

# The profile much match at least one of teh range criteria
force_min <- min(opt_profile_depth_range$min_depth)
force_max <- min(opt_profile_depth_range$max_depth)

# Apply profile min and max restrictions
core_apply_limits <- core_profile_limits %>%
  filter(
    min_depth <= force_min &
    max_depth >= force_max
    )

# Ensure working data set only contains profiles that have confrormed to the range test
core_data_temp <- full_join(core_data_temp, core_apply_limits) %>%
                        filter(!is.na(min_depth))

# Add profile type field and set all to 1.  
# All profile that meet the minimum requirements are profile_range = 1
core_data_temp <- core_data_temp %>%
  mutate(profile_range = 1)

for (i in 2:nrow(opt_profile_depth_range)) {

  #i = 3
  range_min <- opt_profile_depth_range[i,'min_depth']
  range_max <- opt_profile_depth_range[i,'max_depth']

  # Apply profile min and max restrictions
  core_apply_limits <- core_profile_limits %>%
    filter(
      min_depth <= range_min &
      max_depth >= range_max
      ) %>%
    select(
      id
      ) %>%
    mutate (
      profile_range = i 
    )

  # Update profile range to i for these profiles
  # core_data_temp <- full_join(core_data_temp, core_apply_limits) %>%
  #                         filter(!is.na(min_depth))
  core_data_temp <- core_data_temp %>% rows_update(core_apply_limits, by = "id")

}

# Find the gaps within the profiles
profile_gaps <- core_data_temp %>%
  select(
    id,
    depth
    ) %>%
  arrange(id, depth) %>%
  group_by(id) %>%
  mutate(gap = depth - lag(depth, default = 0)) %>%
  ungroup()

# Ensure we do not have gaps in the profile that invalidate it 
for (i_gap in opt_gap_limit) {

  # The limits to be applied in that pass of for loop
  # i_gap <- opt_gap_limit[3]
  i_gap_min = opt_gap_min_depth[which(opt_gap_limit == i_gap)]
  i_gap_max = opt_gap_max_depth[which(opt_gap_limit == i_gap)]
  
  # Which gaps are greater than i_gap
  profile_gaps_remove <- profile_gaps %>%
    filter(gap > i_gap) %>%
    filter(depth >= i_gap_min & depth <= i_gap_max) %>%
    select(id, gap)
  
  # Remonve these profiles from working data set
  core_data_temp <- full_join(core_data_temp, profile_gaps_remove) %>%
                                  filter(is.na(gap)) %>%
                                  select(-c(gap))
}

# Ensure primary fields are all present
core_data_temp <- core_data_temp %>%
  filter(!is.na(lat) & !is.na(lon) & !is.na(date) & !is.na(depth)) 

# clean up working tables
rm(core_apply_limits, core_profile_limits, profile_gaps, profile_gaps_remove)

# --------------------------------------------------------------------------------
# vertical alignment
# --------------------------------------------------------------------------------

# We have a set of temperature profiles that match our criteria we now need to align that data set to match the 
# depth that are in target_depth_range, this will match the range of climatology values in temp_clim_va

# select variable of interest and prepare target_depth field
core_data_temp_clean <- core_data_temp %>%
                              select(
                                id, 
                                lat,
                                lon,
                                date,
                                depth,
                                profile_range,
                                temp
                              ) %>%
                              mutate(
                                target_depth = depth, .after = depth
                              )

# create all possible combinations of location, month and depth levels for interpolation
target_depth_grid <-
  expand_grid(target_depth_range,
              core_data_temp_clean %>% distinct(id, lat, lon, date, profile_range))

# Take account of the profile_range field to determine the profile_range of target_depth_grid
target_depth_grid <- left_join(target_depth_grid, opt_profile_depth_range) %>%
        filter(target_depth <= max_depth)

target_depth_grid <- target_depth_grid %>% 
                        select(
                          target_depth, 
                          id, 
                          lat, 
                          lon,
                          date,
                          profile_range
                          )

# extend temperature depth vectors with target depths
core_data_temp_extended <- full_join(core_data_temp_clean, target_depth_grid) %>% 
  arrange(id, lat, lon, date, target_depth)

# predict spline interpolation on adjusted depth grid for temperature location and month
core_data_temp_interpolated <-
  core_data_temp_extended %>%
  group_by(id, lat, lon, date) %>%
  mutate(temp_spline = spline(target_depth, temp,
                                method = "natural",
                                xout = target_depth)$y) %>%
  ungroup()

# subset interpolated values on target depth range
core_data_temp_interpolated_clean <- 
  inner_join(target_depth_grid, core_data_temp_interpolated)

# select columns and rename to initial names
core_data_temp_interpolated_clean <-
  core_data_temp_interpolated_clean %>%
  select(
    id, 
    lat, 
    lon, 
    date,
    profile_range,
    depth = target_depth,
    temp = temp_spline)

# Create year and month fields
core_data_temp_interpolated_clean <- core_data_temp_interpolated_clean %>%
  mutate(year = year(date),
         month = month(date),
         .after = date)

# clean up working tables
rm(core_data_temp_clean, core_data_temp_extended, core_data_temp_interpolated)

# --------------------------------------------------------------------------------
# Create anomaly profiles
# --------------------------------------------------------------------------------

# Create anomaly profiles as observed - climatology

# Create core_temp_anomaly, but only where we have an climatology temperature
core_temp_anomaly <- left_join(core_data_temp_interpolated_clean, temp_clim_va) %>%
                        filter(!is.na(clim_temp))

# Calculate the anomaly temperature
core_temp_anomaly <- core_temp_anomaly %>%
                      mutate(
                        anomaly = temp - clim_temp
                      )

# --------------------------------------------------------------------------------
# Write files
# --------------------------------------------------------------------------------

# Write the climatology that maps onto depth levels, interpolated temperature profiles that map onto depth levels and resulting anomaly files.

# Write files
core_data_temp_interpolated_clean %>%
  write_rds(file = paste0(path_core_preprocessed, "/", target_year, "_temp_core_va.rds"))

core_temp_anomaly %>%
  write_rds(file = paste0(path_core_preprocessed, "/", target_year, "_temp_anomaly_va.rds"))

# Rename so that names match if just reading existing files
temp_core_va <- core_data_temp_interpolated_clean
temp_anomaly_va <- core_temp_anomaly

rm(core_data_temp_interpolated_clean, core_temp_anomaly)

}

```

## Consolidate years

This process create three files in the path_argo_core_preprocessed directory that will be used for further analysis

```{r consolidate_into_allyears}
# ------------------------------------------------------------------------------
# Process temperature file
# ------------------------------------------------------------------------------
consolidated_created = 0

for (target_year in opt_min_year:opt_max_year) {

  # read the yearly file based on target_year
  temp_core_va_yr <-
  read_rds(file = paste0(path_core_preprocessed, "/", target_year, "_temp_core_va.rds"))

  # Combine into a consolidated all years file
  if (consolidated_created == 0) {
    temp_core_va <- temp_core_va_yr
    consolidated_created = 1
  } else {
    temp_core_va <- rbind(temp_core_va, temp_core_va_yr)
  }
}

# write consolidated files  
temp_core_va %>% 
  write_rds(file = paste0(path_core_preprocessed, "/temp_core_va.rds"))

# remove files to free space
rm(temp_core_va)
rm(temp_core_va_yr)
gc()

# ------------------------------------------------------------------------------
# Process anomaly file
# ------------------------------------------------------------------------------
consolidated_created = 0

for (target_year in opt_min_year:opt_max_year) {

  # read the yearly file based on target_year
  temp_anomaly_va_yr <-
  read_rds(file = paste0(path_core_preprocessed, "/", target_year, "_temp_anomaly_va.rds"))

  # Combine into a consolidated all years file
  if (consolidated_created == 0) {
    temp_anomaly_va <- temp_anomaly_va_yr
    consolidated_created = 1
  } else {
    temp_anomaly_va <- rbind(temp_anomaly_va, temp_anomaly_va_yr)
  }
}

# write consolidated files  
temp_anomaly_va %>% 
  write_rds(file = paste0(path_core_preprocessed, "/temp_anomaly_va.rds"))

# remove files to free space
rm(temp_anomaly_va)
rm(temp_anomaly_va_yr)
gc()
  

```
