
## Data preparation

```{r prep_cluster_data}

# select profile based on profile_range and the appropriate max depth
anomaly_va <- anomaly_va %>% 
  filter(profile_range == opt_profile_range & depth <= opt_max_depth[opt_profile_range])

# Simplified table ready to pivot
anomaly_va_id <- anomaly_va %>%
  select(file_id,
         depth,
         anomaly,
         year, 
         month, 
         lat, 
         lon)

# wide table with each depth becoming a column
anomaly_va_wide <- anomaly_va_id %>%
  select(file_id, depth, anomaly) %>%
  pivot_wider(names_from = depth, values_from = anomaly)

# Drop any rows with missing values N/A caused by gaps in climatology data
anomaly_va_wide <- anomaly_va_wide %>% 
  drop_na()

# Table for cluster analysis
points <- anomaly_va_wide %>%
  column_to_rownames(var = "file_id")

# normalisation?
if (opt_norm_anomaly) {
  
  # Get the maximum anomaly for each profile - the normalisation will then fit -1 to 1
  anomaly_va_id_normalised <- anomaly_va_id %>%
    group_by(file_id) %>%
    mutate(abs_ma = max(abs(anomaly))) %>%
    ungroup()
  
  # divide each anomaly by the maximum anomaly
  anomaly_va_id_normalised <- anomaly_va_id_normalised %>%
    mutate(anomaly = anomaly/abs_ma)
    
  # wide table with each depth becoming a column
  anomaly_va_wide <- anomaly_va_id_normalised %>%
    select(file_id, depth, anomaly) %>%
    pivot_wider(names_from = depth, values_from = anomaly)
  
  # Drop any rows with missing values N/A caused by gaps in climatology data
  anomaly_va_wide <- anomaly_va_wide %>% 
    drop_na()
  
  # Table for cluster analysis
  points_normalised <- anomaly_va_wide %>%
    column_to_rownames(var = "file_id")

}

```

## Cluster analysis

### Cluster means

Based on all floats regardless of surface condition.

```{r cluster_analysis_cluster_details}

for (iType in 1:2) {
  for (inum_clusters in opt_num_clusters_min[opt_profile_range]:opt_num_clusters_max[opt_profile_range]) {
    if (iType == 1) {

      set.seed(1)
      kclusts <-
        tibble(k = inum_clusters) %>%
        mutate(kclust = map(k, ~ kmeans(points, .x, iter.max = opt_max_iterations, nstart = opt_n_start)),
          tidied = map(kclust, tidy),
          glanced = map(kclust, glance),
          augmented = map(kclust, augment, points)
        )
      
      profile_id <-
        kclusts %>%
        unnest(cols = c(augmented)) %>%
        select(file_id = .rownames,
               cluster = .cluster) %>%
        mutate(file_id = as.numeric(file_id),
               cluster = as.character(cluster))
      
      # Add cluster to anomaly_va_id
      anomaly_cluster <-
        full_join(anomaly_va_id, profile_id)
      
      # Add profile_type field
      anomaly_cluster <- anomaly_cluster %>%
        mutate(profile_type = 'base')
      
      # Check null clusters
      anomaly_cluster <- anomaly_cluster %>%
        filter(!is.na(cluster))
      
      # Create table to be used for later analysis and Set the number of clusters field
      if (!exists('anomaly_cluster_all')) {
        anomaly_cluster_all <- anomaly_cluster %>%
          mutate(num_clusters = inum_clusters)
      } else {
        anomaly_cluster_all <-
          rbind(
            anomaly_cluster_all,
            anomaly_cluster %>%
              mutate(num_clusters = inum_clusters)
          )
      }
      
    } else if (iType == 2 & opt_norm_anomaly) {

      set.seed(1)
      kclusts <-
        tibble(k = inum_clusters) %>%
        mutate(kclust = map(k, ~ kmeans(points_normalised, .x, iter.max = opt_max_iterations, nstart = opt_n_start)),
          tidied = map(kclust, tidy),
          glanced = map(kclust, glance),
          augmented = map(kclust, augment, points)
        )
      
      profile_id <-
        kclusts %>%
        unnest(cols = c(augmented)) %>%
        select(file_id = .rownames,
               cluster = .cluster) %>%
        mutate(file_id = as.numeric(file_id),
               cluster = as.character(cluster))
      
      # Add cluster to anomaly_va
      anomaly_cluster_norm <-
        full_join(anomaly_va_id_normalised %>% select(-c(abs_ma)) ,
                  profile_id)
      
      # Add profile_type field
      anomaly_cluster_norm <- anomaly_cluster_norm %>%
        mutate(profile_type = 'adjusted')
      
      # Check null clusters
      anomaly_cluster_norm <- anomaly_cluster_norm %>%
        filter(!is.na(cluster))
      
      # Create table to be used for later analysis and Set the number of clusters field
      if (!exists('anomaly_cluster_all')) {
        anomaly_cluster_all <- anomaly_cluster_norm %>%
          mutate(num_clusters = inum_clusters)
      } else {
        anomaly_cluster_all <-
          rbind(
            anomaly_cluster_all,
            anomaly_cluster_norm %>%
              mutate(num_clusters = inum_clusters)
          )
      }
      
    }
    
  }
}

# Prepare to plot cluster mean
anomaly_cluster_mean <- anomaly_cluster_all %>%
  group_by(profile_type, num_clusters, cluster, depth) %>%
  summarise(
    count_cluster = n(),
    anomaly_mean = mean(anomaly, na.rm = TRUE),
    anomaly_sd = sd(anomaly, na.rm = TRUE)
  ) %>%
  ungroup()

anomaly_cluster_mean_year <- anomaly_cluster_all %>%
  group_by(profile_type, num_clusters, cluster, depth, year) %>%
  summarise(
    count_cluster = n(),
    anomaly_mean = mean(anomaly, na.rm = TRUE),
    anomaly_sd = sd(anomaly, na.rm = TRUE)
  ) %>%
  ungroup()

anomaly_year_mean <- anomaly_cluster_all %>%
  group_by(profile_type, num_clusters, cluster, year) %>%
  summarise(
    count_cluster = n(),
    anomaly_mean = mean(anomaly, na.rm = TRUE),
    anomaly_sd = sd(anomaly, na.rm = TRUE)
  ) %>%
  ungroup()

anomaly_year_mean <- anomaly_year_mean %>%
  group_by(profile_type, num_clusters, year) %>%
  summarise(anomaly_mean = mean(anomaly_mean, na.rm = TRUE)) %>%
  ungroup ()

# Determine profile count by cluster and year
# Count the measurements
cluster_by_year <- anomaly_cluster_all %>% 
  count(profile_type, num_clusters, file_id, cluster, year,
        name = "count_cluster")

# Convert to profiles
cluster_by_year <- cluster_by_year %>% 
  count(profile_type, num_clusters, cluster, year,
        name = "count_cluster")

# total of each type of cluster
cluster_count <- cluster_by_year %>%
  group_by(profile_type, num_clusters, cluster) %>% 
  summarise(count_profiles = sum(count_cluster)) %>%
  ungroup()

anomaly_cluster_mean <- left_join(anomaly_cluster_mean, cluster_count)

```

Base profiles

```{r cluster_analysis_means}

# create figure of cluster mean profiles
anomaly_cluster_mean %>%
  filter (profile_type == "base") %>%
  group_split(profile_type, num_clusters) %>%
  map(
    ~ ggplot(data = .x,) +
      geom_path(aes(x = anomaly_mean,
                    y = depth)) +
      geom_ribbon(
        aes(
          xmax = anomaly_mean + anomaly_sd,
          xmin = anomaly_mean - anomaly_sd,
          y = depth
        ),
        alpha = 0.2
      ) +
      geom_vline(xintercept = 0) +
      scale_y_reverse() +
      facet_wrap(~ paste0(cluster, " (", formatC(count_profiles, big.mark=",") , ")")) +
      coord_cartesian(xlim = opt_xlim) +
      scale_x_continuous(breaks = opt_xbreaks) +
      labs(
        title = paste0(
          'Overall mean anomaly profiles by cluster \n',
          'type = ', unique(.x$profile_type), ', ', 
          'num clusters = ', unique(.x$num_clusters)
        ),
        x = opt_measure_label,
        y = 'depth (m)'
      )
  )

```

Adjusted profiles

```{r cluster_analysis_means_adj}

if (opt_norm_anomaly) {

  # repeat for adjusted profiles profiles
  anomaly_cluster_mean %>%
    filter (profile_type == "adjusted") %>%
    group_split(profile_type, num_clusters) %>%
    map(
      ~ ggplot(data = .x,) +
        geom_path(aes(x = anomaly_mean,
                      y = depth)) +
        geom_ribbon(
          aes(
            xmax = anomaly_mean + anomaly_sd,
            xmin = anomaly_mean - anomaly_sd,
            y = depth
          ),
          alpha = 0.2
        ) +
        geom_vline(xintercept = 0) +
        scale_y_reverse() +
        facet_wrap(~ paste0(cluster, " (", formatC(count_profiles, big.mark=",") , ")")) +
        coord_cartesian(xlim = opt_xlim_adjusted) +
        scale_x_continuous(breaks = opt_xbreaks_adjusted) +
        labs(
          title = paste0(
            'Overall mean anomaly profiles by cluster \n',
            'type = ', unique(.x$profile_type), ', ', 
            'num clusters = ', unique(.x$num_clusters)
          ),
          x = opt_measure_label_adjusted,
          y = 'depth (m)'
        )
    )
}

```

### Cluster mean by year

```{r cluster_mean_year}

# cluster means by year
anomaly_cluster_mean_year %>%
  filter (profile_type == "base") %>%
  mutate(year = as.factor(year)) %>%
  group_split(profile_type, num_clusters) %>%
  map(
    ~ ggplot(data = .x, ) +
      geom_path(aes(
        x = anomaly_mean,
        y = depth,
        col = year
      )) +
      geom_vline(xintercept = 0) +
      scale_y_reverse() +
      facet_wrap(~ cluster) +
      coord_cartesian(xlim = opt_xlim) +
      scale_x_continuous(breaks = opt_xbreaks) +
      scale_color_viridis_d() +
      labs(
        title = paste0(
          'Overall mean anomaly profiles by cluster \n',
          'type = ', unique(.x$profile_type), ', ', 
          'num clusters = ', unique(.x$num_clusters)
        ),
        x = opt_measure_label,
        y = 'depth (m)'
      )
  )

```

Adjusted profiles

```{r cluster_mean_year_adj}

if (opt_norm_anomaly) {
  
  # Repeat for adjusted profiles
  anomaly_cluster_mean_year %>%
    filter (profile_type == "adjusted") %>%
    mutate(year = as.factor(year)) %>%
    group_split(profile_type, num_clusters) %>%
    map(
      ~ ggplot(data = .x, ) +
        geom_path(aes(
          x = anomaly_mean,
          y = depth,
          col = year
        )) +
        geom_vline(xintercept = 0) +
        scale_y_reverse() +
        facet_wrap(~ cluster) +
        coord_cartesian(xlim = opt_xlim_adjusted) +
        scale_x_continuous(breaks = opt_xbreaks_adjusted) +
        scale_color_viridis_d() +
        labs(
          title = paste0(
            'Overall mean anomaly profiles by cluster \n',
            'type = ', unique(.x$profile_type), ', ', 
            'num clusters = ', unique(.x$num_clusters)
          ),
          x = opt_measure_label_adjusted,
          y = 'depth (m)'
        )
    )
  
}
    
```

### Cluster by year

count of each cluster by year

```{r cluster_by_year}

year_min <- min(cluster_by_year$year)
year_max <- max(cluster_by_year$year)

# create figure
cluster_by_year %>%
  filter (profile_type == "base") %>%
  group_split(profile_type, num_clusters) %>%
  map(
    ~ ggplot(data = .x, aes(
        x = year,
        y = count_cluster,
        col = cluster,
        group = cluster
      )) +
      geom_point() +
      geom_line() +
      scale_x_continuous(breaks = seq(year_min, year_max, 2)) +
      scale_color_brewer(palette = 'Dark2') +
      labs(
        title = paste0(
          'Count of profiles by year and cluster \n',
          'type = ', unique(.x$profile_type), ', ', 
          'num clusters = ', unique(.x$num_clusters)
        ),
        x = 'year',
        y = 'number of profiles',
        col = 'cluster'
      )
  )

```

Adjusted profiles

```{r cluster_by_year_adj}

if (opt_norm_anomaly) {

  year_min <- min(cluster_by_year$year)
  year_max <- max(cluster_by_year$year)
  
  # create figure
  cluster_by_year %>%
    filter (profile_type == "adjusted") %>%
    group_split(profile_type, num_clusters) %>%
    map(
      ~ ggplot(data = .x, aes(
          x = year,
          y = count_cluster,
          col = cluster,
          group = cluster
        )) +
        geom_point() +
        geom_line() +
        scale_x_continuous(breaks = seq(year_min, year_max, 2)) +
        scale_color_brewer(palette = 'Dark2') +
        labs(
          title = paste0(
            'Count of profiles by year and cluster \n',
            'type = ', unique(.x$profile_type), ', ', 
            'num clusters = ', unique(.x$num_clusters)
          ),
          x = 'year',
          y = 'number of profiles',
          col = 'cluster'
        )
    )
  
}

```

### Cluster by month

count of each cluster by month of year

```{r cluster_by_month}

# Determine profile count by cluster and year
# Count the measurements
cluster_by_year <- anomaly_cluster_all %>% 
  count(profile_type, num_clusters, file_id, cluster, month,
        name = "count_cluster")

# Convert to profiles
cluster_by_year <- cluster_by_year %>% 
  count(profile_type, num_clusters, cluster, month,
        name = "count_cluster")

# create figure
cluster_by_year %>%
  filter (profile_type == "base") %>%
  group_split(profile_type, num_clusters) %>%
  map(
    ~ ggplot(
      data = .x,
      aes(
        x = month,
        y = count_cluster,
        col = cluster,
        group = cluster
      )
    ) +
      geom_point() +
      geom_line() +
      scale_x_continuous(breaks = seq(1, 12, 2)) +
      scale_color_brewer(palette = 'Dark2') +
      labs(
        title = paste0(
          'Count of profiles by month and cluster \n',
          'type = ', unique(.x$profile_type), ', ', 
          'num clusters = ', unique(.x$num_clusters)
        ),
        x = 'month',
        y = 'number of profiles',
        col = 'cluster'
      )
  )

```

Adjusted profiles

```{r cluster_by_month_adj}

if (opt_norm_anomaly) {

  # create figure
  cluster_by_year %>%
    filter (profile_type == "adjusted") %>%
    group_split(profile_type, num_clusters) %>%
    map(
      ~ ggplot(
        data = .x,
        aes(
          x = month,
          y = count_cluster,
          col = cluster,
          group = cluster
        )
      ) +
        geom_point() +
        geom_line() +
        scale_x_continuous(breaks = seq(1, 12, 2)) +
        scale_color_brewer(palette = 'Dark2') +
        labs(
          title = paste0(
            'Count of profiles by month and cluster \n',
            'type = ', unique(.x$profile_type), ', ', 
            'num clusters = ', unique(.x$num_clusters)
          ),
          x = 'month',
          y = 'number of profiles',
          col = 'cluster'
        )
    )
  
}

```

### Cluster spatial

location of each cluster on map, spatial analysis

```{r spatial_cluster_01}

# create figure
anomaly_cluster_all %>%
  filter (profile_type == "base") %>%
  group_split(profile_type, num_clusters) %>%
  map(
    ~ map +
      geom_tile(data = .x,
                aes(
                  x = lon,
                  y = lat,
                  fill = cluster
                )) +
      lims(y = opt_map_lat_limit) +
      scale_fill_brewer(palette = 'Dark2') +
      labs(
        title = paste0(
          'cluster spatial distribution \n',
          'type = ', unique(.x$profile_type), ', ', 
          'num clusters = ', unique(.x$num_clusters)
        ),
      )
  )

```

Adjusted profiles

```{r spatial_cluster_01_adj}

if (opt_norm_anomaly) {

  # create figure
  anomaly_cluster_all %>%
    filter (profile_type == "adjusted") %>%
    group_split(profile_type, num_clusters) %>%
    map(
      ~ map +
        geom_tile(data = .x,
                  aes(
                    x = lon,
                    y = lat,
                    fill = cluster
                  )) +
        lims(y = opt_map_lat_limit) +
        scale_fill_brewer(palette = 'Dark2') +
        labs(
          title = paste0(
            'cluster spatial distribution \n',
            'type = ', unique(.x$profile_type), ', ', 
            'num clusters = ', unique(.x$num_clusters)
          ),
        )
    )
  
}

```

### Cluster spatial counts

count of measurements for each cluster on separate maps, spatial analysis

```{r spatial_cluster_03}

# Count profiles    
cluster_by_location <- anomaly_cluster_all %>%
  count(profile_type, num_clusters, file_id, lat, lon, cluster,
        name = "count_cluster")

# # Add cluster counts to 
cluster_by_location <- left_join(cluster_by_location, cluster_count)
    
# create figure
cluster_by_location %>%
  filter (profile_type == "base") %>%
  group_split(profile_type, num_clusters) %>%
  map(
    ~ map +
      geom_tile(data = .x %>%
                  count(lat, lon, cluster, count_profiles),
                aes(
                  x = lon,
                  y = lat,
                  fill = n
                )) +
      lims(y = opt_map_lat_limit) +
      scale_fill_gradient(low = "blue",
                          high = "red",
                          trans = "log10") +
      facet_wrap(~ paste0(cluster, " (", formatC(count_profiles, big.mark=",") , ")"), ncol = 2) +
      labs(
        title = paste0(
          'cluster spatial distribution \n',
          'type = ', unique(.x$profile_type), ', ', 
          'num clusters = ', unique(.x$num_clusters)
        )
      )
  )

```

Adjusted profiles

```{r spatial_cluster_03_adj}

if (opt_norm_anomaly) {

  # create figure
  cluster_by_location %>%
    filter (profile_type == "adjusted") %>%
    group_split(profile_type, num_clusters) %>%
    map(
      ~ map +
        geom_tile(data = .x %>%
                    count(lat, lon, cluster, count_profiles),
                  aes(
                    x = lon,
                    y = lat,
                    fill = n
                  )) +
        lims(y = opt_map_lat_limit) +
        scale_fill_gradient(low = "blue",
                            high = "red",
                            trans = "log10") +
        facet_wrap(~ paste0(cluster, " (", formatC(count_profiles, big.mark=",") , ")"), ncol = 2) +
        labs(
          title = paste0(
            'cluster spatial distribution \n',
            'type = ', unique(.x$profile_type), ', ', 
            'num clusters = ', unique(.x$num_clusters)
          )
        )
    )
  
}
```


```{r source_cluster_analysis_overlay_child, include = FALSE}

cluster_analysis_overlay <- c("", "")

if (opt_category == "bgc_ph_ph" | opt_category == "bgc_ph_h_plus") {

  cluster_analysis_overlay <-
  knitr::knit_expand(
    file = here::here("analysis/child/cluster_analysis_overlay.Rmd")
  )
  
}

```

`r knitr::knit(text = unlist(cluster_analysis_overlay))`
