---
title: "BGC-Argo Oxygen Data"
author: "Pasqualina Vonlanthen, David Stappard & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r set_options_global, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

# Task 

Explore BGC-Argo oxygen data through timeseries and climatological maps 

```{r loading_libraries, include=FALSE}
# load in the necessary libraries
library(tidyverse)
library(argodata)
library(ggplot2)
library(lubridate)
library(oce)
```

```{r set_root_directory}

path_argo <- '/nfs/kryo/work/updata/bgc_argo_r_argodata'
path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"

path_argo <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/bgc_argo'
# /nfs/kryo/work/datasets/ungridded/3d/ocean/floats/bgc_argo/preprocessed_bgc_data
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")

```

# Load oxygen data 

Load in delayed-mode, adjusted oxygen data from the BGC-Argo synthetic profile files 

```{r load_preprocessed_oxygen_data}

path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")

oxy_merge <-
  read_rds(file = paste0(path_argo_preprocessed, "/doxy_bgc_observed.rds"))

```

### Southern Ocean surface oxygen

Focus on surface oxygen (top 10 m of the watercolumn) in the Southern Ocean, south of 30ºS 

```{r select_relevant_oxygen_data}
# select only best pH data (with QC flag 1) below 30ºS, for the top 10 m of the watercolumn
oxy_surface <- oxy_merge %>%
  filter(lat <= -30,                   # keep only data at or south of 30ºS
         depth <= 10)                  # keep only data above or at 10 m depth

# check the correct latitudes, QC flags, and depth levels have been filtered
# max(oxy_surface$lat)
# min(oxy_surface$lat)
# table(oxy_surface$doxy_adjusted_qc)
# max(oxy_surface$depth)
# max(oxy_surface$date)
# min(oxy_surface$date)
```

#### Monthly climatological map

Create a map of climatological monthly oxygen values, from January 2013 to August 2021, for the region south of 30ºS 

```{r set_global_theme, include = FALSE}
theme_set(theme_bw())
```


```{r climatological_oxygen_map}

# average oxygen values in the top 10 m for each month in each 2 x 2º longitude/latitude grid 
oxy_mean <- oxy_surface %>%
  group_by(lat, lon, month) %>%
  summarise(oxy_ave_month = mean(doxy_adjusted))

# read in the map from updata
map <-
  read_rds(paste(path_emlr_utilities,
                 "map_landmask_WOA18.rds",
                 sep = ""))

# map a monthly climatology of surface oxygen
map +
  geom_tile(data = oxy_mean,
            aes(lon, lat, fill = oxy_ave_month)) +
  lims(y = c(-85, -25)) +
  scale_fill_gradientn(colors = oceColorsJet(n = oxy_mean$oxy_ave_month)) +
  labs(x = 'lon',
       y = 'lat',
       fill = 'dissolved oxygen \n(µmol kg-1)',
       title = 'Monthly average surface dissolved oxygen values') +
  theme(legend.position = 'bottom')+
  facet_wrap(~month)

```

#### Monthly timeseries

Plot a timeseries of monthly-mean dissolved oxygen for the region south of 30ºS for the upper 10 m of the watercolumn 

```{r plot_monthly_oxygen_timeseries}

# plot a timeseries of monthly values over the whole southern ocean south of 30ºS
oxy_month <- oxy_surface %>%
  group_by(year, month) %>%
  summarise(oxy_ave = mean(doxy_adjusted))

# timeseries of monthly pH values (separate panels for each month)
oxy_month %>%
  ggplot(aes(x = year, y = oxy_ave)) +
  facet_wrap(~month) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(2013, 2021, 2))+
  labs(x = 'year', 
       y = 'dissolved O2 (µmol kg-1)', 
       title = 'monthly mean dissolved oxygen (south of 30ºS)')

```

Monthly average dissolved oxygen, per year, over the whole region south of 30ºS 

```{r monthly_oxygen_per_year}

# timeseries of monthly oxygen values for each year (separate years on the same plot)
oxy_month %>%
  ggplot(aes(x = month, y = oxy_ave, group = year, col = as.character(year)))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 12, 1))+
  labs(x = 'month',
       y = 'dissolved O2 (µmol kg-1)',
       title = 'monthly mean dissolved oxygen (south of 30ºS)',
       col = 'year')

```

### Northeast Pacific surface oxygen 

Focus on surface oxygen (upper 10 m) in the north-east Pacific (10ºN - 70ºN, -190ºE, -140ºE)

```{r select_nepacific_oxygen_data}

# select only best oxygen data (with QC flag 1) between 10 and 70ºN, and 190 and 140ºW, for the top 10 m of the watercolumn
oxy_nepacific <- oxy_merge %>%
  filter(between(lat, 10, 70),
        between(lon, 190, 240),                   # keep only data for the NE Pacific
        depth <= 10)                              # keep only data above or at 10 m depth

# longitudes larger than -180ºE are lon-380 

```

#### Monthly climatological map

Create a map of climatological monthly surface oxygen values, in the north-east Pacific ocean (10ºN - 70ºN, -190ºE, -140ºE)

```{r climatological_oxygen_map_nepacific}

# average oxygen values in the top 10 m for each month in each 2 x 2º longitude/latitude grid 
oxy_mean_nepacific <- oxy_nepacific %>%
  group_by(lat, lon, month) %>%
  summarise(oxy_ave_month = mean(doxy_adjusted))

# map a monthly climatology of surface oxygen (Jan 2013 - August 2021)
map +
  geom_tile(data = oxy_mean_nepacific,
            aes(lon, lat, fill = oxy_ave_month)) +
  lims(y = c(5, 60), 
       x = c(180, 250)) +
  scale_fill_gradientn(colors = oceColorsJet(n = oxy_mean_nepacific$oxy_ave_month)) +
  labs(x = 'lon',
       y = 'lat',
       fill = 'dissolved oxygen \n(µmol kg-1)',
       title = 'Monthly average surface dissolved oxygen') +
  theme(legend.position = 'right')+
  facet_wrap(~month)

```

#### Monthly timeseries

Timeseries of monthly mean oxygen for the northeast Pacific

```{r timeseries_nepacific_oxygen}

# plot a timeseries of monthly values over the whole NE Pacific
oxy_month_nepacific <- oxy_nepacific %>%
  group_by(year, month) %>%
  summarise(oxy_ave = mean(doxy_adjusted))

# timeseries of monthly pH values (separate panels for each month)
oxy_month_nepacific %>%
  ggplot(aes(x = year, y = oxy_ave)) +
  facet_wrap(~month) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(2013, 2021, 2))+
  labs(x = 'year', 
       y = 'dissolved O2 (µmol kg-1)', 
       title = 'monthly mean dissolved oxygen (NE Pacific)')
```

Timeseries of monthly surface oxygen, per year, in the NE Pacific 

```{r timeseries_nepacific_oxygen_per_year}

# timeseries of monthly oxygen values for each year (separate years on the same plot)
oxy_month_nepacific %>%
  ggplot(aes(x = month, y = oxy_ave, group = year, col = as.character(year)))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 12, 1))+
  labs(x = 'month',
       y = 'dissolved O2 (µmol kg-1)',
       title = 'monthly mean dissolved oxygen (NE Pacific)',
       col = 'year')

```

