---
title: "BGC-Argo Temperature Data"
author: "Pasqualina Vonlanthen & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r set_options_global, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

# Task

Explore BGC-Argo temperature data through timeseries and monthly climatological maps


```{r loading_libraries, include=FALSE}
# load in the necessary libraries
library(tidyverse)
library(ggplot2)
library(lubridate)
library(oce)
library(ggOceanMaps)
```


```{r set_global_theme, include=FALSE}

theme_set(theme_bw())
```


```{r set_root_directory}

path_argo <- '/nfs/kryo/work/updata/bgc_argo_r_argodata'
path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"
path_basin_mask <- "/nfs/kryo/work/updata/reccap2/"
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")
```

# Load BGC-SST data 

Using only temperature data from flag A profiles which have an associated flag A pH profile 

```{r load_temp_data}

# flag A temperature data in the top 20 m 
sst <-
  read_rds(file = paste0(path_argo_preprocessed, "/bgc_merge_flag_AB.rds")) %>%
  filter(between(depth, 0, 20)) %>%
  mutate(year = year(date),
         month = month(date)) %>%
  select(
    -c(
      ph_in_situ_total_adjusted,
      ph_in_situ_total_adjusted_error,
      ph_in_situ_total_adjusted_qc,
      profile_ph_in_situ_total_qc
    )
  )

# load in biome separations 

nm_biomes <- read_rds(file = paste0(path_argo_preprocessed, "/nm_biomes.rds"))
```

# Southern Ocean SST 

```{r select_SO_sst}

sst_SO <- sst %>% 
  filter(lat <= -30)

```

## SST offset with depth 

Difference between the in-situ measured sst (20 m) and the profile-mean 20m temperature 

```{r sst_offset_with_depth}

# calculate the mean sst for each surface profile 
mean_profile_sst <- sst_SO %>% 
  group_by(platform_number, cycle_number) %>% 
  mutate(mean_prof_sst = mean(temp_adjusted, na.rm = TRUE), 
         .before = depth) %>% 
  ungroup() %>% 
  mutate(offset = temp_adjusted-mean_prof_sst,   
         .after = mean_prof_sst) # subtract the mean profile sst from the measured in situ sst

mean_profile_sst %>%
  ggplot()+
  geom_point(aes(x = offset, y = depth, col = as.character(year)), size = 0.3, pch = 19) +
  scale_y_reverse()+
  geom_vline(xintercept = 0, col = 'red', size = 0.6)+
  labs(x = 'offset (ºC)',
       y = 'depth (m)',
       col = 'year',
       title = 'in situ sst - mean profile sst')

```

Bin the sst data into 2m-depth intervals and calculate the offset for each sst observation in each depth interval relative to the profile-mean sst

```{r binned_offset}

# bin the sst values into 2m bins and calculate the offset for each 2m bin 

mean_profile_sst_binned <- sst_SO %>% 
  mutate(depth = cut(depth, seq(0, 20, 2), seq(1, 19, 2)),
         depth = as.numeric(as.character(depth))) %>% 
  group_by(platform_number, cycle_number) %>% 
  mutate(mean_prof_sst = mean(temp_adjusted, na.rm = TRUE),
         .before = depth) %>% 
  ungroup() %>% 
  mutate(offset = temp_adjusted-mean_prof_sst, 
         .after = mean_prof_sst) 

# plot the offset of the depth-binned values 
mean_profile_sst_binned %>%
  ggplot()+
  geom_point(aes(x = offset, y = depth, col = as.character(year)), size = 0.3, pch = 19) +
  scale_y_reverse()+
  geom_vline(xintercept = 0, col = 'red', size = 0.6)+
  labs(x = 'offset (ºC)',
       y = 'depth (m)',
       col = 'year',
       title = 'in situ sst - mean profile sst (2m depth bins)')

```

 
Mean binned offset 

```{r mean_binned_sst_offset}

# bin the ph values into 2m bins and calculate the offset for each 2m bin 
profile_sst_binned_ave <- sst_SO %>% 
  mutate(depth = cut(depth, seq(0, 20, 2), seq(1, 19, 2)),
         depth = as.numeric(as.character(depth))) %>% 
  group_by(platform_number, cycle_number) %>% 
  mutate(mean_prof_sst = mean(temp_adjusted, na.rm = TRUE),
         .before = depth) %>% 
  ungroup() %>% 
  mutate(offset = temp_adjusted-mean_prof_sst, 
         .after = mean_prof_sst) %>% 
  group_by(depth) %>% 
  summarise(mean_offset = mean(offset))

# plot the offset of the depth-binned values 
profile_sst_binned_ave %>%
  ggplot()+
  geom_point(aes(x = mean_offset, y = depth), size = 1, pch = 19) +
  geom_line(aes(x = mean_offset, y = depth))+
  scale_y_reverse()+
  geom_vline(xintercept = 0, col = 'red', size = 1)+
  labs(x = 'mean offset (ºC)',
       y = 'depth (m)',
       col = 'year',
       title = 'in situ sst - mean profile sst (2m depth bins)')

```

## Monthly climatological sst 

Map of monthly climatological Argo temperature (BGC floats, flag A pH profiles only), April 2014-December 2021

```{r monthly_clim_sst_maps, fig.asp=2}

# average pH values in the top 20 m for each month in each 2 x 2º longitude/latitude grid 
sst_clim_SO <- sst_SO %>%
  group_by(lat, lon, month) %>%
  summarise(sst_clim_month = mean(temp_adjusted))

# read in the map from updata
map <-
  read_rds(paste(path_emlr_utilities,
                 "map_landmask_WOA18.rds",
                 sep = ""))

# map a monthly climatology of pH (April 2014 - August 2021)
map +
  geom_tile(data = sst_clim_SO,
            aes(lon, lat, fill = sst_clim_month)) +
  lims(y = c(-85, -25)) +
  scale_fill_viridis_c() +
  labs(x = 'lon',
       y = 'lat',
       fill = 'SST',
       title = 'Monthly climatological \nArgo SST (Apr 2014 - Dec 2021)') +
  theme(legend.position = 'right') +
  facet_wrap(~month, ncol = 2)

```


```{r using_ggOceanMaps, fig.asp=2, eval=FALSE}

basemap(limits = -32, data = sst_clim_SO) +   # change to polar projection 
  geom_spatial_tile(data = sst_clim_SO, 
            aes(x = lon,
                y = lat,
                fill = sst_clim_month),
            linejoin = 'mitre',
            col = 'transparent',
            detail = 60)+
  scale_fill_viridis_c()+
  theme(legend.position = 'bottom')+
  labs(x = 'lon',
       y = 'lat',
       fill = 'SST',
       title = 'monthly climatological \nArgo SST (Apr 2014 - Dec 2021)')+
  facet_wrap(~month, ncol = 2)
```

## Monthly timeseries 

Timeseries of monthly SST values, for each Mayot biome 

```{r map_southern_ocean_biomes}

# plot the region separations on a map 

map +
  geom_raster(data = nm_biomes, 
              aes(x = lon, 
                  y = lat, 
                  fill = biome_name)) +
  labs(title = 'Southern Ocean Mayot biomes', 
       fill = 'biome')

```

```{r plot_monthly_sst_timeseries}

# plot a timeseries of monthly values over the whole southern ocean south of 30ºS

sst_SO <- inner_join(sst_SO, nm_biomes)

sst_month_SO <- sst_SO %>%
  group_by(year, month, biome_name) %>%
  summarise(sst_ave = mean(temp_adjusted, na.rm = TRUE))

# timeseries of monthly pH values over 2014-2021 (separate panels for each month)
sst_month_SO %>%
  ggplot(aes(x = year, 
             y = sst_ave, 
             group = biome_name, 
             col = biome_name)) +
  facet_wrap(~month) +
  geom_line() +
  geom_point() +
  labs(x = 'year', 
       y = 'SST (ºC)', 
       title = 'monthly mean Argo SST (Apr 2014-Dec 2021, Southern Ocean)', 
       col = 'region')

```

Monthly average Southern Ocean SST, for each biome 

```{r plot_monthly_sst_per_year}

# timeseries of monthly sst values for each year (separate years on the same plot)
sst_month_SO %>%
  # filter(year != 2014) %>%    # remove the year that is missing data 
  ggplot(aes(x = month, 
             y = sst_ave, 
             group = year,
             col = as.character(year)))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 12, 2))+
  facet_wrap(~biome_name)+
  labs(x = 'month',
       y = 'SST (ºC)',
       title = 'monthly mean Argo SST (Apr 2014-Dec 2021, Southern Ocean regions)',
       col = 'year')

```


```{r yearly_average_sst, eval=FALSE}

# calculate a yearly average SST (one SST value per year, for the whole biome)
sst_year_SO <- sst_SO %>%
  group_by(year, biome_name) %>%
  summarise(sst_ave = mean(temp_adjusted, na.rm = TRUE))

# plot a timeseries of the yearly average SST value (one value per year)
sst_year_SO %>%
  ggplot(aes(x = year, y = sst_ave, group = biome_name, col = biome_name))+
  geom_line()+
  geom_point()+
  labs(x = 'year',
       y = 'SST (ºC)',
       title = 'yearly mean Argo SST (Apr 2014-Dec 2021, south of 30ºS)', 
       col = 'region')

```
