---
title: "Temporal data coverage"
author: "David Stappard & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r set_options_global, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

# Task

Count the number of profiles, and plot their evolution over time.
Categories include core temperature, BGC temperature, ph, disolved oxyge, nitrate, chlorophyll a and combinations of BGC properties.
Counts are profiles by profile_range. The profiles have already been check to ensure they only contain good measurements and that the profiles do not contain significant gaps.

Dependencies
------------
temp_core_va.rds - core preprocessed folder created by temp_core_align_climatology.Rmd
temp_bgc_va.rds - bgc preprocessed folder created by temp_align_climatology.Rmd
pH_bgc_va.rds - bgc preprocessed folder created by pH_align_climatology.Rmd
doxy_bgc_va.rds - bgc preprocessed folder created by doxy_vertical_align.Rmd
nitrate_bgc_va.rds - bgc preprocessed folder created by nitrate_vertical_align.Rmd
chla_bgc_va.rds - bgc preprocessed folder created by chla_vertical_align.Rmd

Outputs
-------
None

```{r loading_libraries, include=FALSE}

library(tidyverse)
library(argodata)
library(lubridate)

```

```{r set_updata_root_directory, include=FALSE}

path_argo <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/bgc_argo'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")

path_argo_core <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/core_argo_r_argodata'
path_argo_core_preprocessed <- paste0(path_argo_core, "/preprocessed_core_data")

```

# Load data

## BGC-Argo data

Read the files created in loading_data.html:

```{r load_bgc_data}

bgc_temp <- read_rds(file = paste0(path_argo_preprocessed, "/temp_bgc_va.rds")) %>%
  filter(!is.na(year))

bgc_ph <- read_rds(file = paste0(path_argo_preprocessed, "/pH_bgc_va.rds")) %>%
  filter(!is.na(year))

bgc_doxy <- read_rds(file = paste0(path_argo_preprocessed, "/doxy_bgc_va.rds")) %>%
  filter(!is.na(year))

bgc_nitrate <- read_rds(file = paste0(path_argo_preprocessed, "/nitrate_bgc_va.rds")) %>%
  filter(!is.na(year))

bgc_chla <- read_rds(file = paste0(path_argo_preprocessed, "/chla_bgc_va.rds")) %>%
  filter(!is.na(year))

```

## Core-Argo data

```{r load_core_temp_data}

core_temp <- read_rds(file = paste0(path_argo_core_preprocessed, "/temp_core_va.rds")) %>%
  filter(!is.na(year))

```


```{r set_ggplot_theme_global, include = FALSE}
theme_set(theme_bw())
```

# Core - temperature

```{r core_by_profile_range}

# Number of measurements
core_temp_count <- core_temp %>%
  group_by(year, month, file_id, profile_range) %>%
  summarise(count_measures = n()) %>%
  ungroup()

# Number of profiles
core_temp_count <- core_temp_count %>%
  group_by(year, month, profile_range) %>%
  summarise(count_profiles = n()) %>%
  ungroup()

# Aggregate profile range
core_temp_count_agg <- core_temp_count %>%
  group_by(year, month) %>%
  summarise(count_profiles = sum(count_profiles)) %>%
  mutate(profile_range = 1) %>%
  ungroup()

core_temp_count_agg <- rbind(
  core_temp_count_agg,
  core_temp_count %>%
    filter (profile_range %in% c(2, 3)) %>%
    group_by(year, month) %>%
    summarise(count_profiles = sum(count_profiles)) %>%
    mutate(profile_range = 2) %>%
    ungroup()
)

core_temp_count_agg <- rbind(
  core_temp_count_agg,
  core_temp_count %>%
    filter (profile_range ==3)
)

# count of temperature profiles by year, month and profile range
core_temp_count_agg %>% 
  ggplot(aes(x = month, y = count_profiles, col = as.character(profile_range))) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year,
             scales = "free_y") +
  scale_x_continuous(breaks = seq(2,12,2)) +
  labs(x = 'month', 
       y = 'number of profiles',
       col = 'profile range',
       title = "Number of profiles",
       subtitle = "Complete core temperature profiles (1 = 600m, 2 = 1,200m, 3 = 1,500m)")

rm (core_temp, core_temp_count, core_temp_count_agg)

```

# BGC - temperature

```{r bgc_temp_by_profile_range}

bgc_temp_count <- bgc_temp %>%
  group_by(year, month, file_id, profile_range) %>%
  summarise(count_measures = n()) %>%
  ungroup()

bgc_temp_count <- bgc_temp_count %>%
  group_by(year, month, profile_range) %>%
  summarise(count_profiles = n()) %>%
  ungroup()

bgc_temp_count_agg <- bgc_temp_count %>%
  group_by(year, month) %>%
  summarise(count_profiles = sum(count_profiles)) %>%
  mutate(profile_range = 1) %>%
  ungroup()

bgc_temp_count_agg <- rbind(
  bgc_temp_count_agg,
  bgc_temp_count %>%
    filter (profile_range %in% c(2, 3)) %>%
    group_by(year, month) %>%
    summarise(count_profiles = sum(count_profiles)) %>%
    mutate(profile_range = 2) %>%
    ungroup()
)

bgc_temp_count_agg <- rbind(
  bgc_temp_count_agg,
  bgc_temp_count %>%
    filter (profile_range ==3)
)

# count of temperature profiles by year, month and profile range
bgc_temp_count_agg %>% 
  ggplot(aes(x = month, y = count_profiles, col = as.character(profile_range))) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year,
             scales = "free_y") +
  scale_x_continuous(breaks = seq(2,12,2)) +
  labs(x = 'month', 
       y = 'number of profiles',
       col = 'profile range',
       title = "Number of profiles",
       subtitle = "Complete bgc temperature profiles (1 = 600m, 2 = 1,200m, 3 = 1,500m)")

rm (bgc_temp_count, bgc_temp_count_agg)

```

# BGC - pH

```{r bgc_ph_by_profile_range}

bgc_ph_count <- bgc_ph %>%
  group_by(year, month, file_id, profile_range) %>%
  summarise(count_measures = n()) %>%
  ungroup()

bgc_ph_count <- bgc_ph_count %>%
  group_by(year, month, profile_range) %>%
  summarise(count_profiles = n()) %>%
  ungroup()

bgc_ph_count_agg <- bgc_ph_count %>%
  group_by(year, month) %>%
  summarise(count_profiles = sum(count_profiles)) %>%
  mutate(profile_range = 1) %>%
  ungroup()

bgc_ph_count_agg <- rbind(
  bgc_ph_count_agg,
  bgc_ph_count %>%
    filter (profile_range %in% c(2, 3)) %>%
    group_by(year, month) %>%
    summarise(count_profiles = sum(count_profiles)) %>%
    mutate(profile_range = 2) %>%
    ungroup()
)

bgc_ph_count_agg <- rbind(
  bgc_ph_count_agg,
  bgc_ph_count %>%
    filter (profile_range ==3)
)

# count of pH profiles by year, month and profile range
bgc_ph_count_agg %>% 
  ggplot(aes(x = month, y = count_profiles, col = as.character(profile_range))) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year,
             scales = "free_y") +
  scale_x_continuous(breaks = seq(2,12,2)) +
  labs(x = 'month', 
       y = 'number of profiles',
       col = 'profile range',
       title = "Number of profiles",
       subtitle = "Complete bgc pH profiles (1 = 614m, 2 = 1,225m, 3 = 1,600m)")

rm (bgc_ph_count, bgc_ph_count_agg)

```

## BGC - temp AND pH

```{r bgc_temp_ph_by_profile_range}

bgc_ph_temp <- inner_join(bgc_ph, bgc_temp %>% distinct(file_id)) 

bgc_ph_temp_count <- bgc_ph_temp %>%
  group_by(year, month, file_id, profile_range) %>%
  summarise(count_measures = n()) %>%
  ungroup()

bgc_ph_temp_count <- bgc_ph_temp_count %>%
  group_by(year, month, profile_range) %>%
  summarise(count_profiles = n()) %>%
  ungroup()

bgc_ph_temp_count_agg <- bgc_ph_temp_count %>%
  group_by(year, month) %>%
  summarise(count_profiles = sum(count_profiles)) %>%
  mutate(profile_range = 1) %>%
  ungroup()

bgc_ph_temp_count_agg <- rbind(
  bgc_ph_temp_count_agg,
  bgc_ph_temp_count %>%
    filter (profile_range %in% c(2, 3)) %>%
    group_by(year, month) %>%
    summarise(count_profiles = sum(count_profiles)) %>%
    mutate(profile_range = 2) %>%
    ungroup()
)

bgc_ph_temp_count_agg <- rbind(
  bgc_ph_temp_count_agg,
  bgc_ph_temp_count %>%
    filter (profile_range ==3)
)

# count of temperature AND pH profiles by year, month and profile range
bgc_ph_temp_count_agg %>% 
  ggplot(aes(x = month, y = count_profiles, col = as.character(profile_range))) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year,
             scales = "free_y") +
  scale_x_continuous(breaks = seq(2,12,2)) +
  labs(x = 'month', 
       y = 'number of profiles',
       col = 'profile range',
       title = "Number of profiles",
       subtitle = "Complete bgc temperature AND pH profiles (1 = 614m, 2 = 1,225m, 3 = 1,600m)")

rm (bgc_ph_temp, bgc_ph_temp_count, bgc_ph_temp_count_agg)

```

# BGC - dissolved oxygen

```{r bgc_doxy_by_profile_range}

bgc_doxy_count <- bgc_doxy %>%
  group_by(year, month, file_id, profile_range) %>%
  summarise(count_measures = n()) %>%
  ungroup()

bgc_doxy_count <- bgc_doxy_count %>%
  group_by(year, month, profile_range) %>%
  summarise(count_profiles = n()) %>%
  ungroup()

bgc_doxy_count_agg <- bgc_doxy_count %>%
  group_by(year, month) %>%
  summarise(count_profiles = sum(count_profiles)) %>%
  mutate(profile_range = 1) %>%
  ungroup()

bgc_doxy_count_agg <- rbind(
  bgc_doxy_count_agg,
  bgc_doxy_count %>%
    filter (profile_range %in% c(2, 3)) %>%
    group_by(year, month) %>%
    summarise(count_profiles = sum(count_profiles)) %>%
    mutate(profile_range = 2) %>%
    ungroup()
)

bgc_doxy_count_agg <- rbind(
  bgc_doxy_count_agg,
  bgc_doxy_count %>%
    filter (profile_range ==3)
)

# count of dissolved oxygen profiles by year, month and profile range
bgc_doxy_count_agg %>% 
  ggplot(aes(x = month, y = count_profiles, col = as.character(profile_range))) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year,
             scales = "free_y") +
  scale_x_continuous(breaks = seq(2,12,2)) +
  labs(x = 'month', 
       y = 'number of profiles',
       col = 'profile range',
       title = "Number of profiles",
       subtitle = "Complete bgc dissolved oxygen profiles (1 = 614m, 2 = 1,225m, 3 = 1,600m)")

rm (bgc_doxy_count, bgc_doxy_count_agg)

```

## BGC - temp AND dissolved oxygen

```{r bgc_temp_doxy_by_profile_range}

bgc_doxy_temp <- inner_join(bgc_doxy, bgc_temp %>% distinct(file_id)) 

bgc_doxy_temp_count <- bgc_doxy_temp %>%
  group_by(year, month, file_id, profile_range) %>%
  summarise(count_measures = n()) %>%
  ungroup()

bgc_doxy_temp_count <- bgc_doxy_temp_count %>%
  group_by(year, month, profile_range) %>%
  summarise(count_profiles = n()) %>%
  ungroup()

bgc_doxy_temp_count_agg <- bgc_doxy_temp_count %>%
  group_by(year, month) %>%
  summarise(count_profiles = sum(count_profiles)) %>%
  mutate(profile_range = 1) %>%
  ungroup()

bgc_doxy_temp_count_agg <- rbind(
  bgc_doxy_temp_count_agg,
  bgc_doxy_temp_count %>%
    filter (profile_range %in% c(2, 3)) %>%
    group_by(year, month) %>%
    summarise(count_profiles = sum(count_profiles)) %>%
    mutate(profile_range = 2) %>%
    ungroup()
)

bgc_doxy_temp_count_agg <- rbind(
  bgc_doxy_temp_count_agg,
  bgc_doxy_temp_count %>%
    filter (profile_range ==3)
)

# count of temperature AND dissolved oxygen profiles by year, month and profile range
bgc_doxy_temp_count_agg %>% 
  ggplot(aes(x = month, y = count_profiles, col = as.character(profile_range))) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year,
             scales = "free_y") +
  scale_x_continuous(breaks = seq(2,12,2)) +
  labs(x = 'month', 
       y = 'number of profiles',
       col = 'profile range',
       title = "Number of profiles",
       subtitle = "Complete bgc temperature AND dissolved oxygen profiles (1 = 614m, 2 = 1,225m, 3 = 1,600m)")

rm (bgc_doxy_temp, bgc_doxy_temp_count, bgc_doxy_temp_count_agg)

```

# BGC - nitrate

```{r bgc_nitrate_by_profile_range}

bgc_nitrate_count <- bgc_nitrate %>%
  group_by(year, month, file_id, profile_range) %>%
  summarise(count_measures = n()) %>%
  ungroup()

bgc_nitrate_count <- bgc_nitrate_count %>%
  group_by(year, month, profile_range) %>%
  summarise(count_profiles = n()) %>%
  ungroup()

bgc_nitrate_count_agg <- bgc_nitrate_count %>%
  group_by(year, month) %>%
  summarise(count_profiles = sum(count_profiles)) %>%
  mutate(profile_range = 1) %>%
  ungroup()

bgc_nitrate_count_agg <- rbind(
  bgc_nitrate_count_agg,
  bgc_nitrate_count %>%
    filter (profile_range %in% c(2, 3)) %>%
    group_by(year, month) %>%
    summarise(count_profiles = sum(count_profiles)) %>%
    mutate(profile_range = 2) %>%
    ungroup()
)

bgc_nitrate_count_agg <- rbind(
  bgc_nitrate_count_agg,
  bgc_nitrate_count %>%
    filter (profile_range ==3)
)

# count of nitrate profiles by year, month and profile range
bgc_nitrate_count_agg %>% 
  ggplot(aes(x = month, y = count_profiles, col = as.character(profile_range))) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year,
             scales = "free_y") +
  scale_x_continuous(breaks = seq(2,12,2)) +
  labs(x = 'month', 
       y = 'number of profiles',
       col = 'profile range',
       title = "Number of profiles",
       subtitle = "Complete bgc nitrate profiles (1 = 614m, 2 = 1,225m, 3 = 1,600m)")

rm (bgc_nitrate_count, bgc_nitrate_count_agg)

```

## BGC - temp AND nitrate

```{r bgc_temp_nitrate_by_profile_range}

bgc_nitrate_temp <- inner_join(bgc_nitrate, bgc_temp %>% distinct(file_id)) 

bgc_nitrate_temp_count <- bgc_nitrate_temp %>%
  group_by(year, month, file_id, profile_range) %>%
  summarise(count_measures = n()) %>%
  ungroup()

bgc_nitrate_temp_count <- bgc_nitrate_temp_count %>%
  group_by(year, month, profile_range) %>%
  summarise(count_profiles = n()) %>%
  ungroup()

bgc_nitrate_temp_count_agg <- bgc_nitrate_temp_count %>%
  group_by(year, month) %>%
  summarise(count_profiles = sum(count_profiles)) %>%
  mutate(profile_range = 1) %>%
  ungroup()

bgc_nitrate_temp_count_agg <- rbind(
  bgc_nitrate_temp_count_agg,
  bgc_nitrate_temp_count %>%
    filter (profile_range %in% c(2, 3)) %>%
    group_by(year, month) %>%
    summarise(count_profiles = sum(count_profiles)) %>%
    mutate(profile_range = 2) %>%
    ungroup()
)

bgc_nitrate_temp_count_agg <- rbind(
  bgc_nitrate_temp_count_agg,
  bgc_nitrate_temp_count %>%
    filter (profile_range ==3)
)

# count of temperature AND nitrate profiles by year, month and profile range
bgc_nitrate_temp_count_agg %>% 
  ggplot(aes(x = month, y = count_profiles, col = as.character(profile_range))) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year,
             scales = "free_y") +
  scale_x_continuous(breaks = seq(2,12,2)) +
  labs(x = 'month', 
       y = 'number of profiles',
       col = 'profile range',
       title = "Number of profiles",
       subtitle = "Complete bgc temperature AND nitrate profiles (1 = 614m, 2 = 1,225m, 3 = 1,600m)")

rm (bgc_nitrate_temp, bgc_nitrate_temp_count, bgc_nitrate_temp_count_agg)

```

# BGC - chlorophyll a

```{r bgc_chla_by_profile_range}

bgc_chla_count <- bgc_chla %>%
  group_by(year, month, file_id, profile_range) %>%
  summarise(count_measures = n()) %>%
  ungroup()

bgc_chla_count <- bgc_chla_count %>%
  group_by(year, month, profile_range) %>%
  summarise(count_profiles = n()) %>%
  ungroup()

bgc_chla_count_agg <- bgc_chla_count %>%
  group_by(year, month) %>%
  summarise(count_profiles = sum(count_profiles)) %>%
  mutate(profile_range = 1) %>%
  ungroup()

bgc_chla_count_agg <- rbind(
  bgc_chla_count_agg,
  bgc_chla_count %>%
    filter (profile_range %in% c(2, 3)) %>%
    group_by(year, month) %>%
    summarise(count_profiles = sum(count_profiles)) %>%
    mutate(profile_range = 2) %>%
    ungroup()
)

bgc_chla_count_agg <- rbind(
  bgc_chla_count_agg,
  bgc_chla_count %>%
    filter (profile_range ==3)
)

# count of chlorophyll a profiles by year, month and profile range
bgc_chla_count_agg %>% 
  ggplot(aes(x = month, y = count_profiles, col = as.character(profile_range))) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year,
             scales = "free_y") +
  scale_x_continuous(breaks = seq(2,12,2)) +
  labs(x = 'month', 
       y = 'number of profiles',
       col = 'profile range',
       title = "Number of profiles",
       subtitle = "Complete bgc chlorophyll a profiles (1 = 614m, 2 = 1,225m, 3 = 1,600m)")

rm (bgc_chla_count, bgc_chla_count_agg)

```

## BGC - temp AND chlorophyll a

```{r bgc_temp_chla_by_profile_range}

bgc_chla_temp <- inner_join(bgc_chla, bgc_temp %>% distinct(file_id)) 

bgc_chla_temp_count <- bgc_chla_temp %>%
  group_by(year, month, file_id, profile_range) %>%
  summarise(count_measures = n()) %>%
  ungroup()

bgc_chla_temp_count <- bgc_chla_temp_count %>%
  group_by(year, month, profile_range) %>%
  summarise(count_profiles = n()) %>%
  ungroup()

bgc_chla_temp_count_agg <- bgc_chla_temp_count %>%
  group_by(year, month) %>%
  summarise(count_profiles = sum(count_profiles)) %>%
  mutate(profile_range = 1) %>%
  ungroup()

bgc_chla_temp_count_agg <- rbind(
  bgc_chla_temp_count_agg,
  bgc_chla_temp_count %>%
    filter (profile_range %in% c(2, 3)) %>%
    group_by(year, month) %>%
    summarise(count_profiles = sum(count_profiles)) %>%
    mutate(profile_range = 2) %>%
    ungroup()
)

bgc_chla_temp_count_agg <- rbind(
  bgc_chla_temp_count_agg,
  bgc_chla_temp_count %>%
    filter (profile_range ==3)
)

# count of temperature AND chlorophyll a profiles by year, month and profile range
bgc_chla_temp_count_agg %>% 
  ggplot(aes(x = month, y = count_profiles, col = as.character(profile_range))) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year,
             scales = "free_y") +
  scale_x_continuous(breaks = seq(2,12,2)) +
  labs(x = 'month', 
       y = 'number of profiles',
       col = 'profile range',
       title = "Number of profiles",
       subtitle = "Complete bgc temperature AND chlorophyll a profiles (1 = 614m, 2 = 1,225m, 3 = 1,600m)")

rm (bgc_chla_temp, bgc_chla_temp_count, bgc_chla_temp_count_agg)

```

# BGC - temp, pH, doxy, nitrate AND chl a

```{r bgc_all_by_profile_range}

bgc_all <- inner_join(bgc_ph, bgc_temp %>% distinct(file_id)) 
bgc_all <- inner_join(bgc_all, bgc_doxy %>% distinct(file_id)) 
bgc_all <- inner_join(bgc_all, bgc_nitrate %>% distinct(file_id)) 
bgc_all <- inner_join(bgc_all, bgc_chla %>% distinct(file_id)) 

bgc_all_count <- bgc_all %>%
  group_by(year, month, file_id, profile_range) %>%
  summarise(count_measures = n()) %>%
  ungroup()

bgc_all_count <- bgc_all_count %>%
  group_by(year, month, profile_range) %>%
  summarise(count_profiles = n()) %>%
  ungroup()

bgc_all_count_agg <- bgc_all_count %>%
  group_by(year, month) %>%
  summarise(count_profiles = sum(count_profiles)) %>%
  mutate(profile_range = 1) %>%
  ungroup()

bgc_all_count_agg <- rbind(
  bgc_all_count_agg,
  bgc_all_count %>%
    filter (profile_range %in% c(2, 3)) %>%
    group_by(year, month) %>%
    summarise(count_profiles = sum(count_profiles)) %>%
    mutate(profile_range = 2) %>%
    ungroup()
)

bgc_all_count_agg <- rbind(
  bgc_all_count_agg,
  bgc_all_count %>%
    filter (profile_range ==3)
)

# count of temp, pH, doxy, nitrate AND chl a profiles by year, month and profile range
bgc_all_count_agg %>% 
  ggplot(aes(x = month, y = count_profiles, col = as.character(profile_range))) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year,
             scales = "free_y") +
  scale_x_continuous(breaks = seq(2,12,2)) +
  labs(x = 'month', 
       y = 'number of profiles',
       col = 'profile range',
       title = "Number of profiles",
       subtitle = "Complete bgc temp, pH, doxy, nitrate AND chl a profiles (1 = 614m, 2 = 1,225m, 3 = 1,600m)")

rm (bgc_all, bgc_all_count, bgc_all_count_agg)

```
