---
title: "Temporal data coverage"
author: "Pasqualina Vonlanthen & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Task

Count the number of bgc-argo profiles, and plot their evolution over time.

```{r loading_libraries, include=FALSE}

library(tidyverse)
library(argodata)
library(lubridate)
```

```{r set_updata_root_directory, include=FALSE}

path_argo <- '/nfs/kryo/work/updata/bgc_argo_r_argodata'
path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"

```

# Load data

Read the files created in loading_data.html:

```{r load_preprpocessed_data}

path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")

bgc_metadata <-
  read_rds(file = paste0(path_argo_preprocessed, "/bgc_metadata.rds"))

```

```{r load_map_and_basin_mask}

basinmask <-
  read_csv(paste(path_emlr_utilities,
                 "basin_mask_WOA18.csv",
                 sep = ""),
           col_types = cols("MLR_basins" = col_character()))

basinmask <- basinmask %>% 
  filter(MLR_basins == unique(basinmask$MLR_basins)[1]) %>% 
  select(lon, lat, basin_AIP)

map <-
  read_rds(paste(path_emlr_utilities,
                 "map_landmask_WOA18.rds",
                 sep = ""))

```

```{r set_ggplot_theme_global, include = FALSE}
theme_set(theme_bw())
```


# QC flags

**QC flags** for values ('`flag`' column) are between 1 and 8, where:

-   1 is 'good' data
-   2 is 'probably good' data,
-   3 is 'probably bad' data,
-   4 is 'bad' data,
-   5 is 'value changed',
-   8 is 'estimated value',
-   9 is 'missing value'.
-   (6 and 7 are not used)

**Profile QC flags** ('`profile_flag`' column) are QC codes attributed to the entire profile, and indicate the number of depth levels (in %) where the value is considered to be good data (QC flags of 1, 2, 5, and 8):

-   'A' means 100% of profile levels contain good data,
-   'B' means 75-\<100% of profile levels contain good data,
-   'C' means 50-75% of profile levels contain good data,
-   'D' means 25-50% of profile levels contain good data,
-   'E' means \>0-50% of profile levels contain good data,
-   'F' means 0% of profile levels contain good data.

# Number of profiles

## Per parameter

```{r number_profiles_per_parameter}
# count the number of profiles per parameter 

bgc_profile_counts <- bgc_metadata %>% 
  select(platform_number, cycle_number, date,
         profile_doxy_qc, profile_ph_in_situ_total_qc, profile_nitrate_qc) %>% 
  pivot_longer(cols = profile_doxy_qc:profile_nitrate_qc,
               names_to = "parameter",
               values_to = "profile_flag",
               names_prefix = "profile_") %>% 
  mutate(year = year(date),
         month = month(date)) %>% 
  count(year, month, parameter, profile_flag) %>%   # count the number of occurrences of unique flags for each parameter, in each month of each year 
  filter(!is.na(profile_flag),
         profile_flag != "")

# the 'parameter' column contains character strings of either 'doxy_qc', 'ph_in_situ_total_qc', or 'nitrate_qc', with the corresponding profile QC flag in the 'profile_flag' column

# count the total number of profiles for each parameter and each flag:
bgc_profile_counts_total <- bgc_metadata %>% 
  select(platform_number, cycle_number, date,
         profile_doxy_qc, profile_ph_in_situ_total_qc, profile_nitrate_qc) %>% 
  pivot_longer(cols = profile_doxy_qc:profile_nitrate_qc,
               names_to = "parameter",
               values_to = "profile_flag",
               names_prefix = "profile_") %>% 
  mutate(year = year(date),
         month = month(date)) %>% 
  count(parameter, profile_flag) %>%   # count the number of occurrences of flags for each parameter
  filter(!is.na(profile_flag),
         profile_flag != "")

```

#### Total number of profiles
```{r count_data}

# count the total number of profiles, regardless of QC flag 
total_data_count <- bgc_metadata %>% 
  select(platform_number, cycle_number, date,
         profile_doxy_qc, profile_ph_in_situ_total_qc, profile_nitrate_qc) %>% 
  pivot_longer(cols = profile_doxy_qc:profile_nitrate_qc,
               names_to = "parameter",
               values_to = "profile_flag",
               names_prefix = "profile_") %>% 
  mutate(year = year(date),
         month = month(date)) %>% 
  count(parameter, profile_flag) %>%   # count the number of occurrences of flags for each parameter
  filter(!is.na(profile_flag),
         profile_flag != "") %>% 
  group_by(parameter) %>% 
  summarise(n = sum(n))

knitr::kable(total_data_count, caption = 'total number of profiles', format = 'markdown')
```


```{r count_usable_data}
# count the number of profiles which have QC flags of A, B, C, D, or E (profiles which contain data that can be used)
usable_data_count <- bgc_metadata %>% 
  select(platform_number, cycle_number, date,
         profile_doxy_qc, profile_ph_in_situ_total_qc, profile_nitrate_qc) %>% 
  pivot_longer(cols = profile_doxy_qc:profile_nitrate_qc,
               names_to = "parameter",
               values_to = "profile_flag",
               names_prefix = "profile_") %>% 
  mutate(year = year(date),
         month = month(date)) %>% 
  count(parameter, profile_flag) %>%   # count the number of occurrences of flags for each parameter
  filter(!is.na(profile_flag),
         profile_flag != "",
         profile_flag != 'F') %>% 
  group_by(parameter) %>% 
  summarise(n = sum(n)) 

knitr::kable(usable_data_count, 
             caption = 'total number of profiles with QC flags A, B, C, D, E', 
             format = 'markdown')
```


```{r count_A_flag_data}
# count the number of profiles with QC flag A 
A_data_count <- bgc_metadata %>% 
  select(platform_number, cycle_number, date,
         profile_doxy_qc, profile_ph_in_situ_total_qc, profile_nitrate_qc) %>% 
  pivot_longer(cols = profile_doxy_qc:profile_nitrate_qc,
               names_to = "parameter",
               values_to = "profile_flag",
               names_prefix = "profile_") %>% 
  mutate(year = year(date),
         month = month(date)) %>% 
  count(parameter, profile_flag) %>%   # count the number of occurrences of flags for each parameter
  filter(!is.na(profile_flag),
         profile_flag != "",
         profile_flag == 'A') %>% 
  group_by(parameter) %>% 
  summarise(n = sum(n)) 

knitr::kable(A_data_count, 
             caption = 'total number of profiles with QC flag A',
             format = 'markdown')
```


```{r count_F_flag_data}
# count the number of profiles with QC Flag F (not usable data)
F_data_count <- bgc_metadata %>% 
  select(platform_number, cycle_number, date,
         profile_doxy_qc, profile_ph_in_situ_total_qc, profile_nitrate_qc) %>% 
  pivot_longer(cols = profile_doxy_qc:profile_nitrate_qc,
               names_to = "parameter",
               values_to = "profile_flag",
               names_prefix = "profile_") %>% 
  mutate(year = year(date),
         month = month(date)) %>% 
  count(parameter, profile_flag) %>%   # count the number of occurrences of flags for each parameter
  filter(!is.na(profile_flag),
         profile_flag != "",
         profile_flag == 'F') %>% 
  group_by(parameter) %>% 
  summarise(n = sum(n)) 

knitr::kable(F_data_count, 
             caption = 'total number of profiles with QC flag F',
             format = 'markdown')

```


Plot the evolution of the number of profiles over time

```{r profiles_timeseries, eval = FALSE}

bgc_profile_counts %>% 
  ggplot(aes(x = month, y = n, col = profile_flag)) +
  geom_line() +
  geom_point() +
  facet_grid(parameter ~ year,
             scales = "free_y") +
  scale_x_continuous(breaks = seq(1,12,4))+
  labs(x = 'month', y = 'number of profiles', title = 'number of profiles per year')

```

```{r profiles_timeseries_per_parameter, fig.asp=0.5}
# draw separate plots for the separate parameters

bgc_profile_counts %>%
  group_split(parameter) %>%   # creates a separate flag count for each parameter 
  map(
    ~ ggplot(data = .x,       # repeats the ggplot for each separate parameter 
             aes(
               x = month, y = n, col = profile_flag
             )) +
      geom_line() +
      geom_point() +
      facet_grid(. ~ year,
                 scales = "free_y") +
      labs(title = paste("Parameter: ", unique(.x$parameter)), 
           x = 'month', y = 'number of profiles') +
      scale_x_continuous(breaks = seq(1,12,4))
  )

ggsave("output/figures/time_series_profiles_per_parameter.png",
width = 7,
height = 4)

```

## All flag A

```{r count_A_flag_profiles}

# count the number of profiles which have a QC flag of A for all three BGC parameters
# the if_all(starts_with()) notation allows to filter over a range of columns simultaneously
# this new approach is identical to your previous solution
# except that it filters also the pres, temp, and sal flags
# (plotted below)

bgc_profile_counts_total_A <- bgc_metadata %>% 
  select(platform_number, cycle_number, date,
         profile_doxy_qc, profile_ph_in_situ_total_qc, profile_nitrate_qc) %>% 
  filter(if_all(starts_with("profile_"), ~. == 'A')) %>% 
  pivot_longer(cols = profile_doxy_qc:profile_nitrate_qc,
               names_to = "parameter",
               values_to = "profile_flag",
               names_prefix = "profile_") %>% 
  mutate(year = year(date),
         month = month(date)) %>% 
  distinct(platform_number, cycle_number, year, month) %>% 
  count(year, month)


```

```{r profiles_timeseries_total_A, fig.asp=0.5}

bgc_profile_counts_total_A %>% 
  ggplot(aes(x = month, y = n)) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year,
             scales = "free_y") +
  scale_x_continuous(breaks = seq(1,12,4)) +
  labs(x = 'month', y = 'number of profiles',
       title = "Number of profiles",
       subtitle = "All three BGC + core parameters (QC flag A)")

ggsave("output/figures/time_series_flag_A_profiles.png",
width = 7,
height = 4)

```

## By region

```{r plot_region_map}

# bgc_metadata <- bgc_metadata %>%
#   mutate(
#     lat = cut(lat, seq(-90, 90, 1), seq(-89.5, 89.5, 1)),
#     lat = as.numeric(as.character(lat)),
#     lon = cut(lon, seq(20, 380, 1), seq(20.5, 379.5, 1)),
#     lon = as.numeric(as.character(lon))
#   )

bgc_grid <- bgc_metadata %>%
  distinct(lat, lon)

bgc_grid <- inner_join(
  basinmask, bgc_grid
)

map +
  geom_raster(data = basinmask,
              aes(lon, lat, fill = basin_AIP)) +
  geom_raster(data = bgc_grid,
              aes(lon, lat)) +
  scale_fill_brewer(palette = "Dark2")

rm(bgc_grid)

```

```{r flag_A_per_region}

bgc_profile_counts_total_A_region <-
  inner_join(bgc_metadata,
             basinmask) %>% 
  select(platform_number, cycle_number, date,
         profile_doxy_qc, profile_ph_in_situ_total_qc, profile_nitrate_qc,
         basin_AIP) %>% 
  filter(if_all(starts_with("profile_"), ~. == 'A')) %>% 
  pivot_longer(cols = profile_doxy_qc:profile_nitrate_qc,
               names_to = "parameter",
               values_to = "profile_flag",
               names_prefix = "profile_") %>% 
  mutate(year = year(date),
         month = month(date)) %>% 
  distinct(platform_number, cycle_number, year, month, basin_AIP) %>% 
  count(year, month, basin_AIP)

```


```{r profiles_timeseries_total_A_per_region, fig.asp=0.5}

bgc_profile_counts_total_A_region %>% 
  ggplot(aes(x = month, y = n, col = basin_AIP)) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year,
             scales = "free_y") +
  scale_x_continuous(breaks = seq(1,12,2)) +
  labs(x = 'month', y = 'number of profiles',
       title = "Number of profiles",
       subtitle = "All three BGC + core parameters (QC flag A)")

ggsave("output/figures/time_series_profiles_per_region.png",
       width = 7,
       height = 4)


```
