---
title: "Combined cluster analysis"
author: "David Stappard & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r set_options_global, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

# Tasks

Carried out cluster analysis having first set options and determined the data sets on which that cluster analysis should be based on.

```{r loading_libraries, include=FALSE}

library(tidyverse)
library(argodata)
library(lubridate)
library(gridExtra)
library(gsw)
library(ggforce)
library(ggpmisc)
library(tidymodels)

```

## Set directories

location of pre-prepared data

```{r set_updata_root_directory, include=FALSE}

path_argo <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/bgc_argo'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")

path_core_argo <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/core_argo_r_argodata'
path_core_preprocessed <- paste0(path_core_argo, "/preprocessed_core_data")

path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"

```

## Category options

What category of data is the cluster analysis related to

```{r set_category_options}

# opt_category
opt_category <- "bgc_doxy"

```

## Analysis options

Define options that are used to determine the nature of the cluster analysis that is carried out.

```{r set_cluster_options}

# Options

# opt_num_clusters
# How many clusters are used in the cluster analysis for each depth 1 (600 m), 2 (1000 m) and 3 (1500 m)
opt_num_clusters_min <- c(8, 8, 4)
opt_num_clusters_max <- c(8, 8, 5)
# Which profile range is used
opt_profile_range <- 3

# options relating to cluster analysis
opt_n_start <- 15
opt_max_iterations <- 500
opt_n_clusters <- 14 # Max number of clusters to try when determining optimal number of clusters

# opt_extreme_determination
# 1 - based on the trend of de-seasonal data - we believe this results in more summer extremes where variation tend to be greater.
# 2 - based on the trend of de-seasonal data by month. grouping is by lat, lon and month.
opt_extreme_determination <- 2

# Options associated with profiles under surface extreme conditions
extreme_type <- c('L', 'N', 'H')
opt_num_clusters_ext_min <- c(4, 4, 4)
opt_num_clusters_ext_max <- c(5, 5, 5)

# Option related to normalising the anomaly profiles.
# TRUE - anomaly profiles are normalised by the surface anomaly. Every depth anomaly is divided by the surface anomaly.
#      - The is only carried out for profiles where the abs(surface temp) > 1.
#      - This analysis is carried out in addition to the analysis on base anomaly profiles.  
# FALSE - The normalisation process is not carried out. 
opt_norm_anomaly <- TRUE

```


```{r set_global_theme, include=TRUE}

theme_set(theme_bw())

map <-
  read_rds(paste(path_emlr_utilities,
                 "map_landmask_WOA18.rds",
                 sep = ""))

```


## Preparation

Prepare data for cluster analysis



```{r prep_bgc_doxy}

if (opt_category == "bgc_doxy") {
  
  # ---------------------------------------------------------------------------------------------
  # spatial restrictions
  # ---------------------------------------------------------------------------------------------
  # global
  opt_lat_min <- -90
  opt_lat_max <- 90
  opt_lon_min <- 20
  opt_lon_max <- 380
  
  # Mapping latitude limits
  opt_map_lat_limit <- c(-85, 85) # global
  
  # ---------------------------------------------------------------------------------------------
  # read data - bgc oxygen must have a ph profile
  # ---------------------------------------------------------------------------------------------
  # read data, applying geographical limits and standardize field names.
  anomaly_va <-
    read_rds(file = paste0(path_argo_preprocessed, "/doxy_anomaly_va.rds")) %>%
    filter (lat >= opt_lat_min &
            lat <= opt_lat_max &
            lon >= opt_lon_min &
            lon <= opt_lon_max) %>%
    select(file_id,
           date,
           year,
           month,
           lat,
           lon,
           profile_range,
           depth, 
           prof_measure = doxy,
           clim_measure = clim_doxy,
           anomaly
           )
  
  # ---------------------------------------------------------------------------------------------
  # Associated data restrictions and formatting
  # ---------------------------------------------------------------------------------------------
  # What is the max depth of each profile_range
  opt_max_depth <- c(600, 1000, 1500)
  
  # opt_measure_label, opt_xlim and opt_xbreaks are associated with formatting
  opt_measure_label <- expression("dissolved oxygen anomaly ( µmol kg"^"-1"~")")
  opt_xlim <- c(-40, 40)
  opt_xbreaks <- c(-40, -20, 0, 20, 40)

  # adjusted to be in scale -1 to 1
  opt_measure_label_adjusted <- "adjusted oxygen anomaly"
  opt_xlim_adjusted <- c(-1, 1)
  opt_xbreaks_adjusted <- c(-1.0, -0.5, 0, 0.5, 1.0)

  # Under extreme analysis
  opt_extreme_analysis <- FALSE
  
}

```


```{r source_cluster_analysis_base_child, include = FALSE}


cluster_analysis_base <-
  knitr::knit_expand(
    file = here::here("analysis/child/cluster_analysis_base.Rmd")
  )


```

`r knitr::knit(text = unlist(cluster_analysis_base))`
