---
title: "Extreme pH Profiles"
author: "Pasqualina Vonlanthen & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r set_options_global, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Task

Compare depth profiles of normal pH and of extreme pH, as identified in the surface OceanSODA pH data product

```{r load_libraries, include=FALSE}
library(tidyverse)
library(lubridate)
library(broom)
library(ggOceanMaps)
library(scico)
library(metR)
library(ggforce)
```

```{r set_global_theme}

theme_set(theme_bw())
HNL_colors <- c("H" = "#b2182b",
                "N" = "#636363",
                "L" = "#2166ac")

```

# Load data

```{r set_root_directories}
path_argo <- '/nfs/kryo/work/updata/bgc_argo_r_argodata'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")
path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"
```

```{r load_data}

# RECCAP2-ocean region mask
region_masks_all_2x2 <- read_rds(file = paste0(path_argo_preprocessed,
                                               "/region_masks_all_2x2.rds"))

region_masks_all_2x2 <- region_masks_all_2x2 %>%
  rename(biome = value) %>% 
  mutate(coast = as.character(coast))

# WOA 18 basin mask

basinmask <-
  read_csv(
    paste(path_emlr_utilities,
          "basin_mask_WOA18.csv",
          sep = ""),
    col_types = cols("MLR_basins" = col_character())
  )

basinmask <- basinmask %>%
  filter(MLR_basins == unique(basinmask$MLR_basins)[1]) %>% 
  select(-c(MLR_basins, basin))

# OceanSODA
OceanSODA <- read_rds(file = paste0(path_argo_preprocessed, "/OceanSODA.rds"))

OceanSODA <- OceanSODA %>%
  mutate(month = month(date))

# load in the full argo data
full_argo <- read_rds(file = paste0(path_argo_preprocessed, "/bgc_merge_pH_qc_1.rds"))

# change the date format for compatibility with OceanSODA pH data
full_argo <- full_argo %>%
  mutate(year = year(date),
         month = month(date)) %>%
  mutate(date = ymd(format(date, "%Y-%m-15")))

```

# Regions

## Biomes

```{r filter_southern_ocean_biomes}

region_masks_all_2x2 <- region_masks_all_2x2 %>%
  filter(region == 'southern',
         biome != 0) %>% 
  select(-region)

```

### Remove coastal data

```{r coast_map}

basemap(limits = -32) +
  geom_spatial_tile(
    data = region_masks_all_2x2,
    aes(x = lon,
        y = lat,
        fill = coast),
    col = 'transparent'
  ) +
  scale_fill_brewer(palette = "Dark2")


```

```{r remove_coastal_data}

region_masks_all_2x2 <- region_masks_all_2x2 %>% 
  filter(coast == "0")

```

### Grid reduction

```{r biomes_map_pre_grid_reduction}

basemap(limits = -32) +
  geom_spatial_tile(
    data = region_masks_all_2x2,
    aes(x = lon,
        y = lat,
        fill = biome),
    col = 'transparent'
  ) +
  scale_fill_brewer(palette = "Dark2")

```

```{r grid_reduction_to_true_2x2_region_mask}

region_masks_all_2x2 <- region_masks_all_2x2 %>%
  count(lon, lat, biome) %>%
  group_by(lon, lat) %>%
  slice_max(n, with_ties = FALSE) %>%
  ungroup()

```

```{r biomes_map_post_grid_reduction}

basemap(limits = -32) +
  geom_spatial_tile(
    data = region_masks_all_2x2,
    aes(x = lon,
        y = lat,
        fill = biome),
    col = 'transparent'
  ) +
  scale_fill_brewer(palette = "Dark2")

```

## Basins

```{r filter_southern_ocean_basins}

basinmask <- basinmask %>%
  filter(lat < -30)

```

### Grid reduction

```{r basin_map_pre_grid_reduction}

basemap(limits = -32) +
  geom_spatial_tile(
    data = basinmask,
    aes(x = lon,
        y = lat,
        fill = basin_AIP),
    col = 'transparent'
  ) +
  scale_fill_brewer(palette = "Dark2")

```

```{r grid_reduction_to_true_2x2_basinmask}

basinmask_2x2 <- basinmask %>%
  mutate(
    lat = cut(lat, seq(-90, 90, 2), seq(-89, 89, 2)), 
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 2), seq(21, 379, 2)), 
    lon = as.numeric(as.character(lon))
   ) # regrid into 2x2º grid

# assign basins from each pixel to to each 2 Lon x Lat pixel, based on the majority of basins in each 2x2 grid  

basinmask_2x2 <- basinmask_2x2 %>%
  count(lon, lat, basin_AIP) %>%
  group_by(lon, lat) %>%
  slice_max(n, with_ties = FALSE) %>%
  ungroup() %>% 
  select(-n)

rm(basinmask)

```

```{r basin_map_post_grid_reduction}

basemap(limits = -32) +
  geom_spatial_tile(
    data = basinmask_2x2 %>% filter(lat < -30),
    aes(x = lon,
        y = lat,
        fill = basin_AIP),
    col = 'transparent'
  ) +
  scale_fill_brewer(palette = "Dark2")

```

# OceanSODA

## Grid reduction

```{r grid_reduction_to_2x2_OceanSODA}

# Note: While reducing lon x lat grid,
# we keep the original number of observations

OceanSODA_2x2 <- OceanSODA %>% 
  mutate(
    lat = cut(lat, seq(-90, 90, 2), seq(-89, 89, 2)), 
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 2), seq(21, 379, 2)), 
    lon = as.numeric(as.character(lon))) # regrid into 2x2º grid 

```


## Apply region masks

```{r apply_region_masks_OceanSODA}

# keep only Southern Ocean data
OceanSODA_2x2_SO <- inner_join(OceanSODA_2x2, region_masks_all_2x2)

# add in basin separations
OceanSODA_2x2_SO <- inner_join(OceanSODA_2x2_SO, basinmask_2x2)
# expected number of rows from -30 to -70º latitude, 360º longitude, for 12 months, 8 years:
# 40 lat x 360 lon x 12 months x 8 years = 1 382 400 rows 

OceanSODA_2x2_SO <- OceanSODA_2x2_SO %>% 
  filter(!is.na(ph_total))

```

# OceanSODA pH anomalies

## Grid level

### Climatological thresholds

### Fit lm models

```{r fit_linear_regression}

# fit a linear regression of OceanSODA pH against time (temporal trend)
# in each lat/lon/month grid

OceanSODA_2x2_SO <- OceanSODA_2x2_SO %>% 
  mutate(year = year(date))


OceanSODA_regression <- OceanSODA_2x2_SO %>% 
  # filter(basin_AIP == "Indian",
  #        biome == "2",
  #        lon < 40) %>%
  nest(data = -c(lon, lat, month)) %>%
  mutate(fit = map(.x = data,
                   .f = ~ lm(ph_total ~ year, data = .x)),
         tidied = map(.x = fit, .f = tidy),
         glanced = map(.x = fit, .f = glance),
         augmented = map(.x = fit, .f = augment))


OceanSODA_regression_tidied <- OceanSODA_regression %>%
  select(-c(data, fit, augmented, glanced)) %>%
  unnest(tidied)

OceanSODA_regression_tidied <- OceanSODA_regression_tidied %>% 
  select(lat:estimate) %>% 
  pivot_wider(names_from = term,
              values_from = estimate) %>% 
  rename(intercept = `(Intercept)`,
         slope = year)

OceanSODA_regression_augmented <- OceanSODA_regression %>%
  select(-c(data, fit, tidied, glanced)) %>%
  unnest(augmented)

OceanSODA_regression_glanced <- OceanSODA_regression %>%
  select(-c(data, fit, tidied, augmented)) %>%
  unnest(glanced)

```

### Slope maps

```{r map_regression_slopes, fig.asp=2}

basemap(limits = -32) +
  geom_spatial_tile(data = OceanSODA_regression_tidied,
                    aes(x = lon,
                        y = lat,
                        fill = slope),
                    col = 'transparent') +
  scale_fill_scico(palette = "vik", midpoint = 0) +
  facet_wrap( ~ month, ncol = 2)

```

### Thresholds

```{r threshold_calculation}

OceanSODA_regression_augmented_stats <- OceanSODA_regression_augmented %>% 
  group_by(lat, lon, month) %>% 
  summarise(residual_sd = sd(.resid)) %>% 
  ungroup()

compare <- full_join(OceanSODA_regression_augmented_stats,
                     OceanSODA_regression_glanced)

compare %>% 
  ggplot(aes(residual_sd -sigma)) +
  geom_histogram()


OceanSODA_2x2_SO_extreme_grid <-
full_join(OceanSODA_regression_augmented %>% 
            select(lat:year, .resid),
          OceanSODA_regression_glanced %>% 
            select(lat:month, sigma))


# # calculate H and L pH thresholds for climatological monthly pH
# OceanSODA_2x2_SO_extreme_grid <- OceanSODA_2x2_SO_extreme_grid %>% 
#   mutate(ph_L = .fitted - 2*(.sigma),
#          ph_H = .fitted + 2*(.sigma))

```

```{r calculate_clim_OceanSODA_pH_grid, eval=FALSE}

# calculate climatological average OceanSODA pH
# and the 95th percentile of the monthly OceanSODA pH
# 
# OceanSODA_2x2_SO_clim_grid <- OceanSODA_2x2_SO %>%
#   group_by(lon, lat, month) %>%
#   summarise(
#     ph_N = mean(ph_total, na.rm = TRUE),
#     ph_H = quantile(ph_total, 0.95, na.rm = TRUE),
#     ph_L = quantile(ph_total, 0.05, na.rm = TRUE)
#   ) %>%
#   ungroup()
# 
# OceanSODA_2x2_SO_extreme_grid <- inner_join(OceanSODA_2x2_SO, OceanSODA_2x2_SO_clim_grid)

```

### Anomaly identification

Calculate OceanSODA pH anomalies: L for abnormally low, H for abnormally high, N for normal pH

```{r calculate_extreme_OceanSODA_pH_grid}

# when the in-situ OceanSODA pH is lower than the 5th percentile (predicted - 2*residual.st.dev), assign 'L' for low extreme
# when the in-situ OceanSODA pH exceeds the 95th percentile (predicted + 2*residual.st.dev), assign 'H' for high extreme
# when the in-situ OceanSODA pH is within 95% of the range, then assign 'N' for normal pH

OceanSODA_2x2_SO_extreme_grid <- OceanSODA_2x2_SO_extreme_grid %>%
  mutate(
    ph_extreme = case_when(
      .resid < -sigma*2 ~ 'L',
      .resid > sigma*2 ~ 'H',
      TRUE ~ 'N'
    )
  ) 

# table(is.na(OceanSODA_2x2_SO_extreme_grid))

OceanSODA_2x2_SO_extreme_grid <- OceanSODA_2x2_SO_extreme_grid %>%
  mutate(ph_extreme = fct_relevel(ph_extreme, "H", "N", "L"))


OceanSODA_2x2_SO_extreme_grid <- 
full_join(OceanSODA_2x2_SO_extreme_grid,
          OceanSODA_regression_tidied)

# pivot_wider two columns (slope and intercept), values_from = estimate, names_from = terms, names.repair = 'unique' 
# gives a slope and intercept column 
# rename date...26 = slope and date...2 = date
# OceanSODA_2x2_SO_extreme_grid <- OceanSODA_2x2_SO_extreme_grid %>% 
#   pivot_wider(names_from = term,
#               values_from = estimate, 
#               names_repair = 'unique') %>% 
#   rename(date = date...2, 
#          regression_slope = date...24,
#          regression_intercept = `(Intercept)`)

# fill in NAs in the slope and intercept columns (values from above for regression_slope and values from below for regression_intercept) (creates duplicate rows) and remove duplicate rows 
# OceanSODA_2x2_SO_extreme_grid <- OceanSODA_2x2_SO_extreme_grid %>%
#   group_by(lon, lat, date, year, month) %>%
#   fill(regression_slope, .direction = 'up') %>%
#   fill(regression_intercept, .direction = 'down') %>%
#   distinct()

```

```{r anomaly_test_plot_grid, fig.asp=0.5}

OceanSODA_2x2_SO_extreme_grid %>%
  group_split(lon, lat, month) %>%
  head(6) %>%
  map(~ ggplot(data = .x) +
        geom_point(aes(x = year,
                       y = ph_total,
                       col = ph_extreme)) +
        geom_abline(data = .x, aes(slope = slope,
                    intercept = intercept)) +
        geom_abline(data = .x, aes(slope = slope,
                    intercept = intercept + 2*sigma),
                    linetype = 2) +
        geom_abline(data = .x, aes(slope = slope,
                    intercept = intercept - 2*sigma),
                    linetype = 2) +
        labs(title = paste(fititle = paste(
          "lon:", unique(.x$lon),
          "| lat:", unique(.x$lat),
          "| month:", unique(.x$month)
          ))) +
        scale_color_manual(values = HNL_colors))

```

### Anomaly maps

Location of OceanSODA pH extremes

```{r map_OceanSODA_ph_extremes_grid, fig.asp=2}

OceanSODA_2x2_SO_extreme_grid %>% 
  group_split(year) %>% 
  head(1) %>%
  map(
    ~ basemap(limits = -32, data = .x)+
      geom_spatial_tile(data = .x,
                        aes(x = lon,
                            y = lat,
                            fill = ph_extreme),
                        linejoin = 'mitre',
                        col = 'transparent',
                        detail = 60
                        ) +
      scale_fill_manual(values = HNL_colors) +
      facet_wrap(~month, ncol = 2)+
      labs(title = paste("Year:", unique(.x$year)),
           fill = 'pH')
  )

```

### Anomaly time series

```{r extreme_pH_timeseries_grid}
# calculate a regional mean pH for each biome, basin, and ph extreme (H/L/N) and plot a timeseries 


OceanSODA_2x2_SO_extreme_grid <- left_join(
  OceanSODA_2x2_SO_extreme_grid,
  basinmask_2x2
)

OceanSODA_2x2_SO_extreme_grid <- left_join(
  OceanSODA_2x2_SO_extreme_grid,
  region_masks_all_2x2
)

OceanSODA_2x2_SO_extreme_grid %>% 
  group_by(year, biome, basin_AIP, ph_extreme) %>% 
  summarise(ph_regional = mean(ph_total, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot(aes(x = year, y = ph_regional, col = ph_extreme))+
  geom_point(size = 0.3)+
  geom_line()+
  scale_color_manual(values = HNL_colors) +
  facet_grid(basin_AIP~biome)+
  theme(legend.position = 'bottom')

```

### Anomaly histogram

```{r extreme_pH_histogram_grid}

OceanSODA_2x2_SO_extreme_grid %>%
  ggplot(aes(ph_total, col = ph_extreme)) +
  geom_density() +
  scale_color_manual(values = HNL_colors) +
  facet_grid(basin_AIP ~ biome) +
  coord_cartesian(xlim = c(8, 8.2)) +
  labs(x = 'value',
       y = 'density',
       col = 'pH anomaly') +
  theme(legend.position = 'bottom')

```

### Threshold histogram

```{r threshold_pH_histogram_grid}

OceanSODA_2x2_SO_extreme_grid %>%
  mutate(ph_extreme = as.double(ph_extreme)) %>%
  pivot_longer(starts_with("ph_"),
               names_to = "level",
               values_to = "value",
               names_prefix = "ph_") %>%
  distinct() %>% 
  ggplot(aes(value, col = level)) +
  geom_density() +
  scale_color_manual(values = HNL_colors, name = "threshold") +
  coord_cartesian(xlim = c(8, 8.2)) +
  lims(y = c(0, 230))+
  theme(legend.position = 'bottom')

```
# Argo

## Grid reduction

```{r grid_reduction_to_2x2_argo}

# Note: While reducing lon x lat grid,
# we keep the original number of observations

full_argo_2x2 <- full_argo %>%
  mutate(
    lat = cut(lat, seq(-90, 90, 2), seq(-89, 89, 2)),
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 2), seq(21, 379, 2)),
    lon = as.numeric(as.character(lon)))  # re-grid to 2x2


```


## Apply region masks

```{r apply_region_masks_argo}

# keep only Southern Ocean argo data
full_argo_2x2_SO <- inner_join(full_argo_2x2, region_masks_all_2x2)

# add in basin separations
full_argo_2x2_SO <- inner_join(full_argo_2x2_SO, basinmask_2x2)

# # remove duplicate rows (keep only distinct rows)
# full_argo_2x2_SO <- full_argo_2x2_SO %>%
#   distinct()

```

## Join OceanSODA

```{r combine_extreme_OceanSODA_to_Argo}

# rename OceanSODA columns
OceanSODA_2x2_SO_extreme <- OceanSODA_2x2_SO_extreme_grid %>%
  rename(OceanSODA_ph = ph_total)

# combine the argo profile data to the surface extreme data
profile_extreme <- inner_join(full_argo_2x2_SO, OceanSODA_2x2_SO_extreme)

```

# Plot profiles

Argo profiles plotted according to the surface OceanSODA pH

L profiles correspond to a surface acidification event (low pH), as recorded in OceanSODA

H profiles correspond to an event of high surface pH, as recorded in OceanSODA

N profiles correspond to normal surface OceanSODA pH

## pH

```{r pH_profiles}

profile_extreme %>%
  group_split(biome, basin_AIP, year) %>% 
  head(1) %>%
  map(
    ~ ggplot(
      data = .x,
      aes(
        x = ph_in_situ_total_adjusted,
        y = depth,
        group = ph_extreme,
        col = ph_extreme
      )
    ) +
      geom_point(pch = 19, size = 0.3) +
      scale_y_reverse() +
      scale_color_manual(values = HNL_colors) +
      facet_wrap(~ month, ncol = 6) +
      labs(
        x = 'Argo pH (total scale)',
        y = 'depth (m)',
        title = paste(
          unique(.x$basin_AIP),
          "|",
          unique(.x$year),
          "| biome:",
          unique(.x$biome)
        ),
        col = 'OceanSODA pH \nanomaly'
      )
  )

```
# Plot monthly profiles

```{r calculate_mean_profiles}
# calculate mean profiles in each basin and biome, for each month between 2014 and 2021 
# cut depth levels at 10, 20, .... etc m
# add seasons 
# Dec, Jan, Feb <- summer 
# Mar, Apr, May <- autumn 
# Jun, Jul, Aug <- winter 
# Sep, Oct, Nov <- spring 

profile_extreme_monthly <- profile_extreme %>%
  mutate(
    depth = Hmisc::cut2(
      depth,
      cuts = c(10, 20, 30, 50, 70, 100, 300, 500, 800, 1000, 1500, 2000, 2500),
      m = 5,
      levels.mean = TRUE
    ),
    depth = as.numeric(as.character(depth))
  ) %>%
  mutate(
    season = case_when(
      between(month, 3, 5) ~ 'autumn',
      between(month, 6, 8) ~ 'winter',
      between(month, 9, 11) ~ 'spring',
      month == 12 | 1 | 2 ~ 'summer'
    ),
    .after = date
  ) %>%
  group_by(season, biome, basin_AIP, ph_extreme, depth) %>%
  summarise(
    ph_mean = mean(ph_in_situ_total_adjusted, na.rm = TRUE),
    temp_mean = mean(temp_adjusted, na.rm = TRUE)
  ) %>%
  ungroup()


```

## pH

#### By season

```{r mean_seasonal_profiles_pH, fig.asp=1}

profile_extreme_monthly %>%
  arrange(depth) %>% 
  group_split(season) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(
               x = ph_mean,
               y = depth,
               group = ph_extreme,
               col = ph_extreme
             )) +
      geom_path() +
      scale_color_manual(values = HNL_colors) +
      labs(title = paste("season:", unique(.x$season)),
           col = 'OceanSODA\npH\nanomaly') +
      scale_y_reverse() +
      facet_grid(basin_AIP ~ biome)
  )

```

#### Averaged over biomes

```{r mean_biome_profiles, fig.asp=1}

profile_extreme_biome <- profile_extreme_monthly %>% 
  group_by(season, biome, ph_extreme, depth) %>% 
  summarise(ph_biome = mean(ph_mean, na.rm = TRUE)) %>% 
  ungroup()
  

profile_extreme_biome %>%
  ggplot(aes(
    x = ph_biome,
    y = depth,
    group = ph_extreme,
    col = ph_extreme
  )) +
  geom_path() +
  scale_color_manual(values = HNL_colors) +
  labs(col = 'OceanSODA\npH\nanomaly') +
  scale_y_continuous(trans = trans_reverser("sqrt"),
                     breaks = c(10, 100, 250, 500, seq(1000, 5000, 500))) +
  facet_grid(season ~ biome)

```

#### Averaged over ocean basins

```{r mean_basin_profiles, fig.asp=1}

profile_extreme_basin <- profile_extreme_monthly %>% 
  group_by(season, basin_AIP, ph_extreme, depth) %>% 
  summarise(ph_basin = mean(ph_mean, na.rm = TRUE)) %>% 
  ungroup()

profile_extreme_basin %>% 
  ggplot(aes(x = ph_basin, 
             y = depth, 
             group = ph_extreme, 
             col = ph_extreme))+
  geom_path()+
  scale_color_manual(values = HNL_colors)+
  labs(col = 'OceanSODA\npH\nanomaly')+
  scale_y_continuous(trans = trans_reverser("sqrt"),
                     breaks = c(10, 100, 250, 500, seq(1000, 5000, 500))) +
  facet_grid(season~basin_AIP)

```
