---
title: "Extreme pH Profiles"
author: "Pasqualina Vonlanthen & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r set_options_global, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

# Task

Compare depth profiles of normal pH and of extreme pH, as identified in the surface OceanSODA pH data product

```{r load_libraries, include=FALSE}
library(tidyverse)
library(lubridate)
library(ggOceanMaps)
```

```{r set_global_theme}
theme_set(theme_bw())
```

# Load data

```{r set_root_directories}
path_argo <- '/nfs/kryo/work/updata/bgc_argo_r_argodata'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")
path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"
```

```{r load_data}
region_masks_all_1x1 <- read_rds(file = paste0(path_argo_preprocessed,
                                               "/region_masks_all_1x1.rds"))

OceanSODA <- read_rds(file = paste0(path_argo_preprocessed, "/OceanSODA.rds"))

# keep only Southern Ocean data (southern ocean RECCAP biomes)
OceanSODA_SO <-
  inner_join(region_masks_all_1x1, OceanSODA) %>%
  filter(region == 'southern',
         value != 0) %>%
  mutate(month = month(date)) %>%
  mutate(date = format_ISO8601(date, precision = 'ym'))

# add in basin separations
basinmask <-
  read_csv(paste(path_emlr_utilities,
                 "basin_mask_WOA18.csv",
                 sep = ""),
           col_types = cols("MLR_basins" = col_character()))

basinmask <- basinmask %>%
  filter(MLR_basins == unique(basinmask$MLR_basins)[1]) %>%
  select(lon, lat, basin_AIP)

OceanSODA_SO <- inner_join(OceanSODA_SO, basinmask)
```

### OceanSODA pH

Climatological monthly OceanSODA pH and the 5th and 95th percentiles, calculated for 2013-2021, with the full spatial OceanSODA data

```{r calculate_clim_OceanSODA_pH}

# calculate climatological average OceanSODA pH, and the 95th percentile of the monthly OceanSODA pH
OceanSODA_SO_clim <- OceanSODA_SO %>%
  group_by(lon, lat, month) %>%
  summarise(
    clim_OceanSODA_ph = mean(ph_total, na.rm = TRUE),
    threshold_high = quantile(ph_total, 0.95, na.rm = TRUE),
    threshold_low = quantile(ph_total, 0.05, na.rm = TRUE)
  ) %>%
  ungroup()

OceanSODA_SO_extreme <- inner_join(OceanSODA_SO, OceanSODA_SO_clim,
                                     by = c('lon', 'lat', 'month'))

```

Calculate extreme OceanSODA pH, L for abnormally low, H for abnormally high, N for normal pH

```{r calculate_extreme_OceanSODA_pH}

# when the in-situ OceanSODA pH is lower than the 5th percentile, assign 'L' for low extreme
# when the in-situ OceanSODA pH exceeds the 95th percentile, assign 'H' for high extreme
# when the in-situ OceanSODA pH is within 95% of the range, then assign 'N' for normal pH
OceanSODA_SO_extreme <- OceanSODA_SO_extreme %>%
  mutate(extreme = case_when(ph_total < threshold_low ~ 'L',
                             ph_total > threshold_high ~ 'H',
                             TRUE ~ 'N')) %>%
  drop_na()

```

Location of OceanSODA pH extremes

```{r map_OceanSODA_extremes}

OceanSODA_SO_extreme %>% 
  group_split(year) %>% 
  map(
    ~ basemap(limits = -32, data = .x)+
      geom_spatial_tile(data = .x, 
                        aes(x = lon, 
                            y = lat, 
                            fill = extreme), 
                        linejoin = 'mitre',
                        col = 'transparent',
                        detail = 60) +
      facet_wrap(~month, ncol = 2)+
      labs(title = paste('OceanSODA pH (High/Low/Normal)', unique(.x$year)),
           fill = 'pH')
  )

# OceanSODA_SO_extreme %>% 
#   filter(year == 2013) %>% 
#   basemap(limits = -32, data = filter(OceanSODA_SO_extreme, year == 2013))+
#   geom_spatial_tile(data = filter(OceanSODA_SO_extreme, year == 2013), 
#                     aes(x = lon,
#                         y = lat, 
#                         fill = extreme),
#                     linejoin = 'mitre',
#                     col = 'transparent',
#                     detail = 60)+
#   facet_wrap(~month, ncol = 2)+
#   labs(title = 'OceanSODA pH (High/Low/Normal) 2013',
#        fill = 'pH')

```

OceanSODA pH timeseries

```{r extreme_pH_timeseries}
# calculate a regional mean pH for each biome, basin, and ph extreme (H/L/N) and plot a timeseries 

OceanSODA_SO_extreme %>% 
  group_by(date, value, basin_AIP, extreme) %>% 
  summarise(ph_regional = mean(ph_total)) %>% 
  ggplot(aes(x = date, y = ph_regional, group = extreme, col = extreme))+
  geom_point()+
  geom_line()+
  scale_x_discrete(breaks = c('2013-01', '2014-01', '2015-01', '2016-01', '2017-01', '2018-01', '2019-01', '2020-01'))+
  facet_grid(basin_AIP~value)+
  labs(title = 'OceanSODA pH (High/Low/Normal)',
       x = 'date',
       y = 'regional mean pH',
       col = 'High / Low / Normal pH') +
  theme(legend.position = 'bottom')

```

### Argo data

Load in full Argo data

```{r load_Argo_data}
# load in the full argo data
full_argo <- read_rds(file = paste0(path_argo_preprocessed, "/bgc_merge.rds"))

# change the date format for compatibility with OceanSODA pH data
full_argo <- full_argo %>%
  mutate(year = year(date),
         month = month(date)) %>%
  mutate(date = format_ISO8601(date, precision = 'ym')) %>% 
  select(depth:platform_number,
         cycle_number,
         float_serial_no,
         date,
         lat,
         lon,
         profile_pres_qc:profile_ph_in_situ_total_qc) %>% 
  filter(ph_in_situ_total_adjusted_qc == '1')

# add in RECCAP biomes, and keep only Southern Ocean argo data
full_argo <- inner_join(full_argo, region_masks_all_1x1) 

full_argo_SO <- full_argo %>%
  filter(region == 'southern',
         value != 0)

# add in basin separations to full profile argo data
full_argo_SO <- inner_join(full_argo_SO, basinmask)

# rename OceanSODA columns
OceanSODA_SO_extreme <- OceanSODA_SO_extreme %>%
  rename(OceanSODA_ph_uncert = ph_total_uncert,
         OceanSODA_ph = ph_total)
```

```{r combine_extreme_OceanSODA_to_Argo}

# combine the argo profile data to the surface extreme data
profile_extreme <- inner_join(full_argo_SO, OceanSODA_SO_extreme)

```

Calculate mean monthly profiles

```{r calculate_mean_profiles}
# calculate mean profiles in each basin and biome, for each month between 2014 and 2021 

monthly_profile_extreme <- profile_extreme %>% 
  mutate(depth = round(depth, digits = 0)) %>%
  group_by(year, month, value, basin_AIP, extreme, depth) %>% 
  summarise(mean_ph = mean(ph_in_situ_total_adjusted, na.rm = TRUE),
            mean_temp = mean(temp_adjusted, na.rm = TRUE))

```

# Argo profiles

Argo profiles plotted according to the surface OceanSODA pH

L profiles correspond to a surface acidification event (low pH), as recorded in OceanSODA

H profiles correspond to an event of high surface pH, as recorded in OceanSODA

N profiles correspond to normal surface OceanSODA pH

## Atlantic basin

### 2019

### pH

```{r Atlantic_pH_profiles}

profile_extreme %>%
  filter(basin_AIP == 'Atlantic',
         year == '2019') %>%
  group_split(value) %>%
  map(
    ~ ggplot(data = .x,
         aes(x = ph_in_situ_total_adjusted,
             y = depth,
             group = extreme,
             col = extreme))+
      geom_point(pch = 19, size = 0.5)+
      scale_y_reverse()+
      theme_bw()+
      facet_wrap(~month, ncol = 6)+
      labs(x = 'Argo pH (total scale)',
       y = 'depth (m)',
       title = paste('Atlantic basin, 2019, biome', unique(.x$value)),
       col = 'OceanSODA pH \n(high/low/normal)')
  )
```

### Temperature

```{r Atlantic_temp_profiles}

# plot temperature profiles for the Atlantic
profile_extreme %>%
  filter(basin_AIP == 'Atlantic',
         year == '2019') %>%
  group_split(value) %>%
  map(
    ~ ggplot(data = .x,
             aes(x = temp_adjusted,
                 y = depth,
                 group = extreme,
                 col = extreme))+
      geom_point(pch = 19, size = 0.5)+
      scale_y_reverse()+
      theme_bw()+
      facet_wrap(~month, ncol = 6)+
      labs(x = 'Argo temperature (ºC)',
           y = 'depth (m)',
           title = paste('Atlantic basin, 2019, biome', unique(.x$value)),
           col = 'OceanSODA pH \n(high/low/normal)')
  )
```

## Pacific basin

### 2019

### pH

```{r Pacific_pH_profiles}
profile_extreme %>%
  filter(basin_AIP == 'Pacific',
         year == '2019') %>%
  group_split(value) %>%
  map(
    ~ ggplot(data = .x,
             aes(x = ph_in_situ_total_adjusted,
                 y = depth,
                 group = extreme,
                 col = extreme))+
      geom_point(pch = 19, size = 0.5)+
      scale_y_reverse()+
      theme_bw()+
      facet_wrap(~month, ncol = 6)+
      labs(x = 'Argo pH (total scale)',
           y = 'depth (m)',
           title = paste('Pacific basin, 2019, biome', unique(.x$value)),
           col = 'OceanSODA pH \n(high/low/normal)')
  )
```

### Temperature

```{r Pacific_temp_profiles}

profile_extreme %>%
  filter(basin_AIP == 'Pacific',
         year == '2019') %>%
  group_split(value) %>%
  map(
    ~ ggplot(data = .x,
             aes(x = temp_adjusted,
                 y = depth,
                 group = extreme,
                 col = extreme))+
      geom_point(pch = 19, size = 0.5)+
      scale_y_reverse()+
      theme_bw()+
      facet_wrap(~month, ncol = 6)+
      labs(x = 'Argo temperature (ºC)',
           y = 'depth (m)',
           title = paste('Pacific basin, 2019, biome', unique(.x$value)),
           col = 'OceanSODA pH \n(high/low/normal)')
  )

```

## Indian basin

### 2019

### pH

```{r Indian_pH_profiles}
profile_extreme %>%
  filter(basin_AIP == 'Indian',
         year == '2019') %>%
  group_split(value) %>%
  map(
    ~ ggplot(data = .x,
             aes(x = ph_in_situ_total_adjusted,
                 y = depth,
                 group = extreme,
                 col = extreme))+
      geom_point(pch = 19, size = 0.5)+
      scale_y_reverse()+
      theme_bw()+
      facet_wrap(~month, ncol = 6)+
      labs(x = 'Argo pH (total scale)',
           y = 'depth (m)',
           title = paste('Indian basin, 2019, biome', unique(.x$value)),
           col = 'OceanSODA pH \n(high/low/normal)')
  )

```

### Temperature

```{r Indian_temp_profiles}

profile_extreme %>%
  filter(basin_AIP == 'Indian',
         year == '2019') %>%
  group_split(value) %>%
  map(
    ~ ggplot(data = .x,
             aes(x = temp_adjusted,
                 y = depth,
                 group = extreme,
                 col = extreme))+
      geom_point(pch = 19, size = 0.5)+
      scale_y_reverse()+
      theme_bw()+
      facet_wrap(~month, ncol = 6)+
      labs(x = 'Argo temperature (ºC))',
           y = 'depth (m)',
           title = paste('Indian basin, 2019, biome', unique(profile_extreme$value)),
           col = 'OceanSODA pH \n(high/low/normal)')
  )

```

Mean monthly profiles

```{r plot_mean_monthly_profiles}

# monthly_profile_extreme %>%
#   group_split(year, basin_AIP, value) %>%
#   map(
#     ~ggplot(data = .x,
#              aes(x = mean_ph,
#                  y = depth,
#                  group = extreme,
#                  col = extreme))+
#       geom_point(size = 0.3)+
#       scale_y_reverse()+
#       facet_wrap(~month)
#   )

monthly_profile_extreme %>% 
  ggplot(aes(x = mean_ph,
             y = depth, 
             group = extreme, 
             col = extreme))+
  geom_point(size = 0.3)+
  scale_y_reverse()+
  facet_grid(basin_AIP~value)


```
