---
title: "Extreme pH Profiles"
author: "Pasqualina Vonlanthen & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r set_options_global, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Task

Compare depth profiles of normal pH and of extreme pH, as identified in the surface OceanSODA pH data product

```{r load_libraries, include=FALSE}
library(tidyverse)
library(lubridate)
library(broom)
library(ggOceanMaps)
library(scico)
library(metR)
library(ggforce)
```

```{r set_global_theme}

theme_set(theme_bw())
HNL_colors <- c("H" = "#b2182b",
                "N" = "#636363",
                "L" = "#2166ac")

```

# Load data

```{r set_root_directories}
path_argo <- '/nfs/kryo/work/updata/bgc_argo_r_argodata'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")
path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"
```

```{r load_data}

# RECCAP2-ocean region mask

region_masks_all_2x2 <- read_rds(file = paste0(path_argo_preprocessed,
                                               "/region_masks_all_2x2.rds"))

region_masks_all_2x2 <- region_masks_all_2x2 %>%
  rename(biome = value) %>% 
  mutate(coast = as.character(coast))

# WOA 18 basin mask

basinmask <-
  read_csv(
    paste(path_emlr_utilities,
          "basin_mask_WOA18.csv",
          sep = ""),
    col_types = cols("MLR_basins" = col_character())
  )

basinmask <- basinmask %>%
  filter(MLR_basins == unique(basinmask$MLR_basins)[1]) %>% 
  select(-c(MLR_basins, basin))

# OceanSODA
OceanSODA <- read_rds(file = paste0(path_argo_preprocessed, "/OceanSODA.rds"))

OceanSODA <- OceanSODA %>%
  mutate(year = year(date),
         month = month(date))

# full argo data
full_argo <- read_rds(file = paste0(path_argo_preprocessed, "/bgc_merge_pH_qc_1.rds"))

# change the date format for compatibility with OceanSODA pH data
full_argo <- full_argo %>%
  mutate(year = year(date),
         month = month(date)) %>%
  mutate(date = ymd(format(date, "%Y-%m-15")))

map <-
  read_rds(paste(path_emlr_utilities,
                 "map_landmask_WOA18.rds",
                 sep = ""))
```

# Regions

## Biomes

```{r filter_southern_ocean_biomes}

region_masks_all_2x2 <- region_masks_all_2x2 %>%
  filter(region == 'southern',
         biome != 0) %>% 
  select(-region)

```

### Remove coastal data

```{r coast_map, eval=FALSE}

basemap(limits = -32) +
  geom_spatial_tile(
    data = region_masks_all_2x2,
    aes(x = lon,
        y = lat,
        fill = coast),
    col = 'transparent'
  ) +
  scale_fill_brewer(palette = "Dark2")


```

```{r coast_flat_map, fig.asp=1}
map + 
  geom_tile(data = region_masks_all_2x2, 
            aes(x = lon, 
                y = lat, 
                fill = coast))+
  lims(y = c(-85, -30))+
  scale_fill_brewer(palette = 'Dark2')


```

```{r remove_coastal_data}

region_masks_all_2x2 <- region_masks_all_2x2 %>% 
  filter(coast == "0")

```

### Grid reduction

```{r biomes_map_pre_grid_reduction, eval=FALSE}

basemap(limits = -32) +
  geom_spatial_tile(
    data = region_masks_all_2x2,
    aes(x = lon,
        y = lat,
        fill = biome),
    col = 'transparent'
  ) +
  scale_fill_brewer(palette = "Dark2")

```

```{r biomes_flat_map_pre_grid_reduction, fig.asp=1}

map +
  geom_tile(data = region_masks_all_2x2, 
            aes(x = lon, 
                y = lat, 
                fill = biome))+
  lims(y = c(-85, -30))+
  scale_fill_brewer(palette = 'Dark2')

```

```{r grid_reduction_to_true_2x2_region_mask}

region_masks_all_2x2 <- region_masks_all_2x2 %>%
  count(lon, lat, biome) %>%
  group_by(lon, lat) %>%
  slice_max(n, with_ties = FALSE) %>%
  ungroup()

```

```{r biomes_map_post_grid_reduction, eval=FALSE}

basemap(limits = -32) +
  geom_spatial_tile(
    data = region_masks_all_2x2,
    aes(x = lon,
        y = lat,
        fill = biome),
    col = 'transparent'
  ) +
  scale_fill_brewer(palette = "Dark2")

```

```{r biomes_flat_map_post_grid_reduction, fig.asp=1}

map+
  geom_tile(data = region_masks_all_2x2,
            aes(x = lon,
                y = lat, 
                fill = biome))+
  lims(y = c(-85, -30))+
  scale_fill_brewer(palette = 'Dark2')

```

## Basins

```{r filter_southern_ocean_basins}

basinmask <- basinmask %>%
  filter(lat < -30)

```

### Grid reduction

```{r basin_map_pre_grid_reduction, eval=FALSE}

basemap(limits = -32) +
  geom_spatial_tile(
    data = basinmask,
    aes(x = lon,
        y = lat,
        fill = basin_AIP),
    col = 'transparent'
  ) +
  scale_fill_brewer(palette = "Dark2")

```

```{r basin_flat_map_pre_grid_reduction, fig.asp=1}
map +
  geom_tile(data = basinmask, 
            aes(x = lon, 
                y = lat, 
                fill = basin_AIP))+
  lims(y = c(-85, -30))+
  scale_fill_brewer(palette = 'Dark2')

```

```{r grid_reduction_to_true_2x2_basinmask}

basinmask_2x2 <- basinmask %>%
  mutate(
    lat = cut(lat, seq(-90, 90, 2), seq(-89, 89, 2)), 
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 2), seq(21, 379, 2)), 
    lon = as.numeric(as.character(lon))
   ) # regrid into 2x2º grid

# assign basins from each pixel to to each 2 Lon x Lat pixel, based on the majority of basins in each 2x2 grid  

basinmask_2x2 <- basinmask_2x2 %>%
  count(lon, lat, basin_AIP) %>%
  group_by(lon, lat) %>%
  slice_max(n, with_ties = FALSE) %>%
  ungroup() %>% 
  select(-n)

rm(basinmask)

```

```{r basin_map_post_grid_reduction, eval=FALSE}

basemap(limits = -32) +
  geom_spatial_tile(
    data = basinmask_2x2 %>% filter(lat < -30),
    aes(x = lon,
        y = lat,
        fill = basin_AIP),
    col = 'transparent'
  ) +
  scale_fill_brewer(palette = "Dark2")

```

```{r basin_flat_map_post_grid_reduction, fig.asp=1}

map+
  geom_tile(data = basinmask_2x2 %>% filter(lat < -30),
            aes(x = lon,
                y = lat, 
                fill = basin_AIP))+
  lims(y = c(-85, -30))+
  scale_fill_brewer(palette = 'Dark2')

```

# OceanSODA

## Grid reduction

```{r grid_reduction_to_2x2_OceanSODA}

# Note: While reducing lon x lat grid,
# we keep the original number of observations

OceanSODA_2x2 <- OceanSODA %>% 
  mutate(
    lat_raw = lat,
    lon_raw = lon,
    lat = cut(lat, seq(-90, 90, 2), seq(-89, 89, 2)), 
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 2), seq(21, 379, 2)), 
    lon = as.numeric(as.character(lon))) # regrid into 2x2º grid 

```

## Apply region masks

```{r apply_region_masks_OceanSODA}

# keep only Southern Ocean data
OceanSODA_2x2_SO <- inner_join(OceanSODA_2x2, region_masks_all_2x2)

# add in basin separations
OceanSODA_2x2_SO <- inner_join(OceanSODA_2x2_SO, basinmask_2x2)
# expected number of rows from -30 to -70º latitude, 360º longitude, for 12 months, 8 years:
# 40 lat x 360 lon x 12 months x 8 years = 1 382 400 rows 

OceanSODA_2x2_SO <- OceanSODA_2x2_SO %>% 
  filter(!is.na(ph_total))

```

# OceanSODA pH anomalies

## Grid level

### Climatological thresholds

### Fit lm models

```{r fit_linear_regression}

# fit a linear regression of OceanSODA pH against time (temporal trend)
# in each lat/lon/month grid

OceanSODA_regression <- OceanSODA_2x2_SO %>% 
  # filter(basin_AIP == "Indian",
  #        biome == "2",
  #        lon < 40) %>%
  nest(data = -c(lon, lat, month)) %>%
  mutate(fit = map(.x = data,
                   .f = ~ lm(ph_total ~ year, data = .x)),
         tidied = map(.x = fit, .f = tidy),
         glanced = map(.x = fit, .f = glance),
         augmented = map(.x = fit, .f = augment))


OceanSODA_regression_tidied <- OceanSODA_regression %>%
  select(-c(data, fit, augmented, glanced)) %>%
  unnest(tidied)

OceanSODA_regression_tidied <- OceanSODA_regression_tidied %>% 
  select(lat:estimate) %>% 
  pivot_wider(names_from = term,
              values_from = estimate) %>% 
  rename(intercept = `(Intercept)`,
         slope = year)

OceanSODA_regression_augmented <- OceanSODA_regression %>%
  select(-c(fit, tidied, glanced, data)) %>%
  unnest(augmented) %>% 
  select(lat:year, .resid)

OceanSODA_regression_augmented <- bind_cols(
  OceanSODA_regression_augmented,
  OceanSODA_2x2_SO %>% select(
    lon_raw, lat_raw, basin_AIP, biome)
)

OceanSODA_regression_glanced <- OceanSODA_regression %>%
  select(-c(data, fit, tidied, augmented)) %>%
  unnest(glanced)

```

### Slope maps

```{r map_regression_slopes, fig.asp=2, eval=FALSE}

basemap(limits = -32) +
  geom_spatial_tile(data = OceanSODA_regression_tidied,
                    aes(x = lon,
                        y = lat,
                        fill = slope),
                    col = 'transparent') +
  scale_fill_scico(palette = "vik", midpoint = 0) +
  facet_wrap( ~ month, ncol = 2)

```

```{r map_regression_slopes_flat, fig.asp=1}

map+
  geom_tile(data = OceanSODA_regression_tidied,
            aes(x = lon,
                y = lat, 
                fill = slope))+
  scale_fill_scico(palette = 'vik', midpoint = 0)+
  lims(y = c(-85, -30))+
  facet_wrap(~month, ncol = 2)

```

### Residual st. dev. maps

```{r map_sigma, fig.asp=2, eval=FALSE}

basemap(limits = -32)+
  geom_spatial_tile(data = OceanSODA_regression_glanced,
                    aes(x = lon, 
                        y = lat, 
                        fill = sigma),
                    col = 'transparent')+
  scale_fill_viridis_c()+
  facet_wrap(~month, ncol = 2)+
  labs(fill = '1 residual \nst. dev.')

```

```{r flat_map_sigma, fig.asp=1}

map+
  geom_tile(data = OceanSODA_regression_glanced,
            aes(x = lon,
                y = lat, 
                fill = sigma))+
  scale_fill_viridis_c()+
  lims(y = c(-85, -30))+
  facet_wrap(~month, ncol = 2)+
  labs(fill = '1 residual \nst. dev.')

```

### Anomaly identification

Calculate OceanSODA pH anomalies: L for abnormally low, H for abnormally high, N for normal pH

```{r calculate_extreme_OceanSODA_pH_grid}

# when the in-situ OceanSODA pH is lower than the 5th percentile (predicted - 2*residual.st.dev), assign 'L' for low extreme
# when the in-situ OceanSODA pH exceeds the 95th percentile (predicted + 2*residual.st.dev), assign 'H' for high extreme
# when the in-situ OceanSODA pH is within 95% of the range, then assign 'N' for normal pH

# combine observations and regression statistics

OceanSODA_2x2_SO_extreme_grid <-
  full_join(
    OceanSODA_regression_augmented,
    OceanSODA_regression_glanced %>%
      select(lat:month, sigma)
  )

# identify observations in anomaly classes

OceanSODA_2x2_SO_extreme_grid <- OceanSODA_2x2_SO_extreme_grid %>%
  mutate(
    ph_extreme = case_when(
      .resid < -sigma*2 ~ 'L',
      .resid > sigma*2 ~ 'H',
      TRUE ~ 'N'
    )
  ) 

OceanSODA_2x2_SO_extreme_grid <- OceanSODA_2x2_SO_extreme_grid %>%
  mutate(ph_extreme = fct_relevel(ph_extreme, "H", "N", "L"))


# combine with regression coefficients

OceanSODA_2x2_SO_extreme_grid <-
  full_join(OceanSODA_2x2_SO_extreme_grid,
            OceanSODA_regression_tidied)


```

```{r anomaly_test_plot_grid, fig.asp=0.5}

OceanSODA_2x2_SO_extreme_grid %>%
  group_split(lon, lat, month) %>%
  head(6) %>%
  map(~ ggplot(data = .x) +
        geom_point(aes(x = year,
                       y = ph_total,
                       col = ph_extreme)) +
        geom_abline(data = .x, aes(slope = slope,
                    intercept = intercept)) +
        geom_abline(data = .x, aes(slope = slope,
                    intercept = intercept + 2*sigma),
                    linetype = 2) +
        geom_abline(data = .x, aes(slope = slope,
                    intercept = intercept - 2*sigma),
                    linetype = 2) +
        labs(title = paste(fititle = paste(
          "lon:", unique(.x$lon),
          "| lat:", unique(.x$lat),
          "| month:", unique(.x$month)
          ))) +
        scale_color_manual(values = HNL_colors))

```

### Anomaly maps

Location of OceanSODA pH extremes

```{r map_OceanSODA_ph_extremes_grid, fig.asp=2, eval=FALSE}

OceanSODA_2x2_SO_extreme_grid %>% 
  group_split(year) %>% 
  # head(2) %>%
  map(
    ~ basemap(limits = -32, data = .x)+
      geom_spatial_tile(data = .x,
                        aes(x = lon,
                            y = lat,
                            fill = ph_extreme),
                        linejoin = 'mitre',
                        col = 'transparent',
                        detail = 60
                        ) +
      scale_fill_manual(values = HNL_colors) +
      facet_wrap(~month, ncol = 2)+
      labs(title = paste("Year:", unique(.x$year)),
           fill = 'pH')
  )

```

```{r flat_map_OceanSODA_ph_extremes_grid, fig.asp=2}

OceanSODA_2x2_SO_extreme_grid %>% 
  group_split(year) %>% 
  map(
    ~map +
      geom_tile(data = .x,
                aes(x = lon,
                    y = lat, 
                    fill = ph_extreme))+
      scale_fill_manual(values = HNL_colors)+
      facet_wrap(~month, ncol = 2)+
      lims(y = c(-85, -30))+
      labs(title = paste('Year:', unique(.x$year)),
           fill = 'pH')
  )

```

### Anomaly time series

```{r extreme_pH_timeseries_grid}
# calculate a regional mean pH for each biome, basin, and ph extreme (H/L/N) and plot a timeseries 

OceanSODA_2x2_SO_extreme_grid %>% 
  group_by(year, biome, basin_AIP, ph_extreme) %>% 
  summarise(ph_regional = mean(ph_total, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot(aes(x = year, y = ph_regional, col = ph_extreme))+
  geom_point(size = 0.3)+
  geom_line()+
  scale_color_manual(values = HNL_colors) +
  facet_grid(basin_AIP~biome)+
  theme(legend.position = 'bottom')

```

### Anomaly histogram

```{r extreme_pH_histogram_grid}

OceanSODA_2x2_SO_extreme_grid %>%
  ggplot(aes(ph_total, col = ph_extreme)) +
  geom_density() +
  scale_color_manual(values = HNL_colors) +
  facet_grid(basin_AIP ~ biome) +
  coord_cartesian(xlim = c(8, 8.2)) +
  labs(x = 'value',
       y = 'density',
       col = 'pH anomaly') +
  theme(legend.position = 'bottom')

```

# Argo

## Grid reduction

```{r grid_reduction_to_2x2_argo, eval=FALSE}

# Note: While reducing lon x lat grid,
# we keep the original number of observations

full_argo_2x2 <- full_argo %>%
  mutate(
    lat_raw = lat,
    lon_raw = lon,
    lat = cut(lat, seq(-90, 90, 2), seq(-89, 89, 2)),
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 2), seq(21, 379, 2)),
    lon = as.numeric(as.character(lon)))  # re-grid to 2x2


```

## Apply region masks

```{r apply_region_masks_argo, eval=FALSE}

# keep only Southern Ocean argo data
full_argo_2x2_SO <- inner_join(full_argo_2x2, region_masks_all_2x2)

# add in basin separations
full_argo_2x2_SO <- inner_join(full_argo_2x2_SO, basinmask_2x2)

```

## Join OceanSODA

```{r combine_extreme_OceanSODA_to_Argo}

# rename OceanSODA columns
OceanSODA_2x2_SO_extreme_grid <- OceanSODA_2x2_SO_extreme_grid %>%
  select(-c(lon, lat)) %>% 
  rename(OceanSODA_ph = ph_total,
         lon = lon_raw,
         lat = lat_raw)

# combine the argo profile data to the surface extreme data
profile_extreme <- inner_join(
  full_argo %>% 
    select(year, month, date, lon, lat, depth,
           ph_in_situ_total_adjusted,
           platform_number,
           cycle_number),
  OceanSODA_2x2_SO_extreme_grid %>% 
    select(year, month, lon, lat,
           OceanSODA_ph, ph_extreme,
           biome, basin_AIP))

```

# Plot profiles

Argo profiles plotted according to the surface OceanSODA pH

L profiles correspond to a surface acidification event (low pH), as recorded in OceanSODA

H profiles correspond to an event of high surface pH, as recorded in OceanSODA

N profiles correspond to normal surface OceanSODA pH

## Raw

```{r profiles_raw}

profile_extreme %>%
  group_split(biome, basin_AIP, year) %>% 
  #head(1) %>%
  map(
    ~ ggplot(
      data = .x,
      aes(
        x = ph_in_situ_total_adjusted,
        y = depth,
        group = ph_extreme,
        col = ph_extreme
      )
    ) +
      geom_point(pch = 19, size = 0.3) +
      scale_y_reverse() +
      scale_color_manual(values = HNL_colors) +
      facet_wrap(~ month, ncol = 6) +
      labs(
        x = 'Argo pH (total scale)',
        y = 'depth (m)',
        title = paste(
          unique(.x$basin_AIP),
          "|",
          unique(.x$year),
          "| biome:",
          unique(.x$biome)
        ),
        col = 'OceanSODA pH \nanomaly'
      )
  )

```

## Averaged profiles

```{r calculate_mean_profiles}
# calculate mean profiles in each basin and biome, for each month between 2014 and 2021 
# cut depth levels at 10, 20, .... etc m
# add seasons 
# Dec, Jan, Feb <- summer 
# Mar, Apr, May <- autumn 
# Jun, Jul, Aug <- winter 
# Sep, Oct, Nov <- spring 

profile_extreme <- profile_extreme %>%
  mutate(
    depth = Hmisc::cut2(
      depth,
      cuts = c(10, 20, 30, 50, 70, 100, 300, 500, 800, 1000, 1500, 2000, 2500),
      m = 5,
      levels.mean = TRUE
    ),
    depth = as.numeric(as.character(depth))
  ) %>%
  mutate(
    season = case_when(
      between(month, 3, 5) ~ 'autumn',
      between(month, 6, 8) ~ 'winter',
      between(month, 9, 11) ~ 'spring',
      month == 12 | 1 | 2 ~ 'summer'
    ),
    .after = date
  ) 

```

### Overall mean

```{r mean_profiles, fig.asp=1}

profile_extreme_mean <- profile_extreme %>%
  group_by(ph_extreme, depth) %>%
  summarise(ph_mean = mean(ph_in_situ_total_adjusted, na.rm = TRUE)) %>%
  ungroup()

profile_extreme_mean %>%
  arrange(depth) %>%
  ggplot(aes(
    x = ph_mean,
    y = depth,
    group = ph_extreme,
    col = ph_extreme
  )) +
  geom_path() +
  scale_color_manual(values = HNL_colors) +
  labs(title = "Overall mean",
       col = 'OceanSODA\npH\nanomaly',
       y = 'log(depth)') +
  scale_y_continuous(trans = trans_reverser("sqrt"),
                     breaks = c(10, 100, 250, 500, seq(1000, 5000, 500)))


```

Number of profiles

```{r mean_profile_count, fig.asp=1}

profile_count_mean <- profile_extreme %>% 
  select(ph_extreme, platform_number, cycle_number) %>% 
  distinct() %>%       # keep only unique combinations of ph_extreme, float number, and cycle number 
  group_by(ph_extreme) %>% 
  count()

profile_count_mean %>% 
  ggplot(aes(x = ph_extreme, y = n, fill = ph_extreme))+
  geom_col(width = 0.5)+
  labs(y = 'number of profiles',
       title = 'Number of profiles')

```

Surface Argo pH vs surface OceanSODA pH (30 m)

```{r argo_vs_OceanSODA_ph_mean, fig.asp=1}

profile_extreme %>% 
  filter(depth < 30) %>% 
  group_split(ph_extreme) %>% 
  map(
  ~ggplot(data = .x, aes(x = OceanSODA_ph, 
             y = ph_in_situ_total_adjusted))+
  geom_bin2d(data = .x, aes(x = OceanSODA_ph, 
                 y = ph_in_situ_total_adjusted)) +
  geom_abline(slope = 1, intercept = 0)+
  coord_fixed(ratio = 1, 
              xlim = c(7.95, 8.2),
              ylim = c(7.95, 8.2))+
  facet_grid(basin_AIP ~ biome) +
    labs(title = paste('pH extreme:', unique(.x$ph_extreme)),
         x = 'OceanSODA pH',
         y = 'Argo pH')
  )

```

### Season x biome

```{r mean_profiles_season_biome, fig.asp=1}

profile_extreme_biome <- profile_extreme %>% 
  group_by(season, biome, ph_extreme, depth) %>% 
  summarise(ph_biome = mean(ph_in_situ_total_adjusted, na.rm = TRUE)) %>% 
  ungroup()
  

profile_extreme_biome %>%
  ggplot(aes(
    x = ph_biome,
    y = depth,
    group = ph_extreme,
    col = ph_extreme
  )) +
  geom_path() +
  scale_color_manual(values = HNL_colors) +
  labs(col = 'OceanSODA\npH\nanomaly',
       y = 'log(depth)') +
  scale_y_continuous(trans = trans_reverser("sqrt"),
                     breaks = c(10, 100, 250, 500, seq(1000, 5000, 500))) +
  facet_grid(season ~ biome)

```

Number of profiles season x biome

```{r count_profiles_season_biome, fig.asp=1}

profile_count_biome <- profile_extreme %>% 
  select(season, biome, ph_extreme, platform_number, cycle_number) %>% 
  distinct() %>% 
  group_by(season, biome, ph_extreme) %>% 
  count()

profile_count_biome %>% 
  ggplot(aes(x = ph_extreme, y = n, fill = ph_extreme))+
  geom_col(width = 0.5)+
  facet_grid(season ~ biome)+
  labs(y = 'number of profiles',
       title = 'Number of profiles season x biome')
```

Surface Argo vs surface OceanSODA pH (30 m) season x biome

```{r argo_vs_OceanSODA_biome, fig.asp=1}

profile_extreme %>% 
  filter(depth < 30) %>% 
  group_split(ph_extreme) %>% 
  map(
  ~ggplot(data = .x, aes(x = OceanSODA_ph, 
             y = ph_in_situ_total_adjusted))+
  geom_bin2d(data = .x, aes(x = OceanSODA_ph, 
                 y = ph_in_situ_total_adjusted)) +
  geom_abline(slope = 1, intercept = 0)+
  coord_fixed(ratio = 1, 
              xlim = c(7.95, 8.2),
              ylim = c(7.95, 8.2))+
  facet_grid(season~biome) +
    labs(title = paste( 'pH extreme:', unique(.x$ph_extreme)),
         x = 'OceanSODA pH',
         y = 'Argo pH')
  )

```

### Season x basin

```{r mean_basin_profiles, fig.asp=1}

profile_extreme_basin <- profile_extreme %>% 
  group_by(season, basin_AIP, ph_extreme, depth) %>% 
  summarise(ph_basin = mean(ph_in_situ_total_adjusted, na.rm = TRUE)) %>% 
  ungroup()

profile_extreme_basin %>% 
  ggplot(aes(x = ph_basin, 
             y = depth, 
             group = ph_extreme, 
             col = ph_extreme))+
  geom_path()+
  scale_color_manual(values = HNL_colors)+
  labs(col = 'OceanSODA\npH\nanomaly',
       y = 'log(depth)')+
  scale_y_continuous(trans = trans_reverser("sqrt"),
                     breaks = c(10, 100, 250, 500, seq(1000, 5000, 500))) +
  facet_grid(season~basin_AIP)

```

Number of profiles season x basin

```{r count_basin_profiles, fig.asp=1}

profile_count_basin <- profile_extreme %>% 
  select(season, basin_AIP, ph_extreme, platform_number, cycle_number) %>% 
  distinct() %>% 
  group_by(season, basin_AIP, ph_extreme) %>% 
  count()

profile_count_basin %>% 
  ggplot(aes(x = ph_extreme, y = n, fill = ph_extreme))+
  geom_col(width = 0.5)+
  facet_grid(season~basin_AIP)+
  labs(y = 'number of profiles',
       title = 'Number of profiles season x basin')
```

Surface Argo vs surface OceanSODA pH (30 m) season x basin

```{r argo_vs_OceanSODA_pH_basin}

profile_extreme %>% 
  filter(depth < 30) %>% 
  group_split(ph_extreme) %>% 
  map(
  ~ggplot(data = .x, aes(x = OceanSODA_ph, 
             y = ph_in_situ_total_adjusted))+
  geom_bin2d(data = .x, aes(x = OceanSODA_ph, 
                 y = ph_in_situ_total_adjusted)) +
  geom_abline(slope = 1, intercept = 0)+
  coord_fixed(ratio = 1, 
              xlim = c(7.95, 8.2),
              ylim = c(7.95, 8.2))+
  facet_grid(season~basin_AIP) +
    labs(title = paste('pH extreme:', unique(.x$ph_extreme)),
         x = 'OceanSODA pH',
         y = 'Argo pH')
  )

```

### Season x biome x basin

```{r mean_profiles_season_biome_basin, fig.asp=1}

profile_extreme_season <- profile_extreme %>%
  group_by(season, biome, basin_AIP, ph_extreme, depth) %>%
  summarise(ph_mean = mean(ph_in_situ_total_adjusted, na.rm = TRUE)) %>%
  ungroup()

profile_extreme_season %>%
  arrange(depth) %>%
  group_split(season) %>%
  # head(1) %>%
  map(
    ~ ggplot(
      data = .x,
      aes(
        x = ph_mean,
        y = depth,
        group = ph_extreme,
        col = ph_extreme
      )
    ) +
      geom_path() +
      scale_color_manual(values = HNL_colors) +
      labs(title = paste("season:", unique(.x$season)),
           col = 'OceanSODA\npH\nanomaly',
           y = 'log(depth)') +
      scale_y_continuous(
        trans = trans_reverser("sqrt"),
        breaks = c(10, 100, 250, 500, seq(1000, 5000, 500))
      ) +
      facet_grid(basin_AIP ~ biome)
  )

```

Number of profiles per season x biome x basin x pH extreme

```{r count_number_season_biome_basin_profiles}

profile_count_season <- profile_extreme %>% 
  select(season, 
         biome, 
         basin_AIP, 
         ph_extreme, 
         platform_number, 
         cycle_number) %>% 
  distinct() %>% 
  group_by(season, biome, basin_AIP, ph_extreme) %>% 
  count()

profile_count_season %>% 
  group_by(season) %>% 
  group_split(season) %>% 
  map(
    ~ggplot()+
      geom_col(data =.x, 
               aes(x = ph_extreme,
                   y = n,
                   fill = ph_extreme),
               width = 0.5)+
      facet_grid(basin_AIP ~ biome)+
      labs(y = 'number of profiles',
           title = paste('season:', unique(.x$season)))
  )


```

Surface OceanSODA pH vs surface Argo pH (30 m)

```{r surface_OceanSODA_vs_argo_ph, fig.asp=1}

profile_extreme %>% 
  filter(depth < 30) %>% 
  group_split(season, ph_extreme) %>% 
  map(
  ~ggplot(data = .x, aes(x = OceanSODA_ph, 
             y = ph_in_situ_total_adjusted))+
  geom_bin2d(data = .x, aes(x = OceanSODA_ph, 
                 y = ph_in_situ_total_adjusted)) +
  geom_abline(slope = 1, intercept = 0)+
  coord_fixed(ratio = 1, 
              xlim = c(7.95, 8.2),
              ylim = c(7.95, 8.2))+
  facet_grid(basin_AIP ~ biome) +
    labs(title = paste('season:', unique(.x$season), 
                        '| pH extreme:', unique(.x$ph_extreme)),
         x = 'OceanSODA pH',
         y = 'Argo pH')
  )

```
