---
title: "Extreme Temperature Profiles"
author: "Pasqualina Vonlanthen & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r set_options_global, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Task

Compare depth profiles of normal temperature and of extreme temperature, as identified in the surface OceanSODA data product

```{r load_libraries, include=FALSE}
library(tidyverse)
library(lubridate)
library(broom)
library(ggOceanMaps)
library(scico)
library(metR)
library(ggforce)
library(oce)

```

```{r set_global_theme}

theme_set(theme_bw())
HNL_colors <- c("H" = "#b2182b",
                "N" = "#636363",
                "L" = "#2166ac")

HNL_colors_map <- c('H' = 'red3',
                    'N' = 'gray90',
                    'L' = 'blue3')
```

# Load data

```{r set_root_directories}
path_argo <- '/nfs/kryo/work/updata/bgc_argo_r_argodata'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")
path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"
path_updata <- '/nfs/kryo/work/updata'
path_argo_clim_temp <- paste0(path_updata, "/argo_climatology/temperature")
```

```{r load_data}

# RECCAP2-ocean region mask

region_masks_all_2x2 <- read_rds(file = paste0(path_argo_preprocessed,
                                               "/region_masks_all_2x2.rds"))
  
region_masks_all_2x2 <- region_masks_all_2x2 %>%
  rename(biome = value) %>% 
  mutate(coast = as.character(coast))

# WOA 18 basin mask

basinmask <-
  read_csv(
    paste(path_emlr_utilities,
          "basin_mask_WOA18.csv",
          sep = ""),
    col_types = cols("MLR_basins" = col_character())
  )

basinmask <- basinmask %>%
  filter(MLR_basins == unique(basinmask$MLR_basins)[1]) %>% 
  select(-c(MLR_basins, basin))

# OceanSODA temperature

OceanSODA_temp <- read_rds(file = paste0(path_argo_preprocessed, "/OceanSODA_temp.rds"))

OceanSODA_temp <- OceanSODA_temp %>%
  mutate(month = month(date))

# Argo flag A profiles

full_argo <- read_rds(file = paste0(path_argo_preprocessed, "/bgc_merge_flag_A.rds"))

full_argo <- full_argo %>%
  select(-c(ph_in_situ_total_adjusted:ph_in_situ_total_adjusted_error,
            profile_ph_in_situ_total_qc))

# change the date format for compatibility with OceanSODA data
full_argo <- full_argo %>%
  mutate(year = year(date),
         month = month(date)) %>%
  mutate(date = ymd(format(date, "%Y-%m-15")))

# base map for plotting
map <-
  read_rds(paste(path_emlr_utilities,
                 "map_landmask_WOA18.rds",
                 sep = ""))
# restrict base map to Southern Ocean
map <- map +
  lims(y = c(-85, -30))

```

# Regions

## Biomes

```{r filter_southern_ocean_biomes}

region_masks_all_2x2 <- region_masks_all_2x2 %>%
  filter(region == 'southern',
         biome != 0) %>% 
  select(-region)
```

## Remove coastal data

```{r coast_map, fig.asp=0.4}
map + 
  geom_tile(data = region_masks_all_2x2, 
            aes(x = lon, 
                y = lat, 
                fill = coast))+
  scale_fill_brewer(palette = 'Dark2')

```

```{r remove_coastal_data}

region_masks_all_2x2 <- region_masks_all_2x2 %>% 
  filter(coast == "0")

```

## Grid reduction

```{r biomes_map_pre_grid_reduction, fig.asp=0.4}

map +
  geom_tile(data = region_masks_all_2x2, 
            aes(x = lon, 
                y = lat, 
                fill = biome))+
  scale_fill_brewer(palette = 'Dark2')

```

```{r grid_reduction_to_true_2x2_region_mask}

region_masks_all_2x2 <- region_masks_all_2x2 %>%
  count(lon, lat, biome) %>%
  group_by(lon, lat) %>%
  slice_max(n, with_ties = FALSE) %>%
  ungroup()

 
```

```{r biomes_map_post_grid_reduction, fig.asp=0.4}

map+
  geom_tile(data = region_masks_all_2x2,
            aes(x = lon,
                y = lat,
                fill = biome))+
  scale_fill_brewer(palette = 'Dark2')

```

```{r polar_projection_biome_map}

basemap(limits = -32)+
  geom_spatial_tile(data = region_masks_all_2x2,
                    aes(x = lon,
                        y = lat,
                        fill = biome),
                    col = NA)+
  scale_fill_brewer(palette = 'Dark2')

```

## Basins

```{r filter_southern_ocean_basins}

basinmask <- basinmask %>%
  filter(lat < -30)
```

## Grid reduction

```{r basin_map_pre_grid_reduction, fig.asp=0.4}
map +
  geom_tile(data = basinmask, 
            aes(x = lon, 
                y = lat, 
                fill = basin_AIP))+
  scale_fill_brewer(palette = 'Dark2')
```

```{r grid_reduction_to_true_2x2_basinmask}

basinmask_2x2 <- basinmask %>%
  mutate(
    lat = cut(lat, seq(-90, 90, 2), seq(-89, 89, 2)),
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 2), seq(21, 379, 2)),
    lon = as.numeric(as.character(lon))
   )

# assign basins from each pixel to to each 2 Lon x Lat pixel, based on the majority of basins in each 2x2 grid

basinmask_2x2 <- basinmask_2x2 %>%
  count(lon, lat, basin_AIP) %>%
  group_by(lon, lat) %>%
  slice_max(n, with_ties = FALSE) %>%
  ungroup() %>%
  select(-n)

rm(basinmask)

```

```{r basin_map_post_grid_reduction, fig.asp=0.4}

map+
  geom_tile(data = basinmask_2x2 %>% filter(lat < -30),
            aes(x = lon,
                y = lat,
                fill = basin_AIP))+
  scale_fill_brewer(palette = 'Dark2')

```

```{r polar_projection_basin_map}

basemap(limits = -32)+
  geom_spatial_tile(data = basinmask_2x2 %>% filter(lat < -32),
                    aes(x = lon,
                        y = lat,
                        fill = basin_AIP),
                    col = NA)+
  scale_fill_brewer(palette = 'Dark2')

```

# OceanSODA

### Monthly clim

```{r calculate_climatological_temperature}

OceanSODA_temp <- OceanSODA_temp %>% 
  group_by(lon, lat, month) %>% 
  mutate(clim_temp = mean(temperature, na.rm = TRUE),
         clim_diff = temperature - clim_temp,
         .after = temperature) %>% 
  ungroup()

```

## Grid reduction

```{r grid_reduction_to_2x2_OceanSODA_temp}

# Note: While reducing lon x lat grid,
# we keep the original number of observations

OceanSODA_temp_2x2 <- OceanSODA_temp %>%
  mutate(
    lat_raw = lat,
    lon_raw = lon,
    lat = cut(lat, seq(-90, 90, 2), seq(-89, 89, 2)),
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 2), seq(21, 379, 2)),
    lon = as.numeric(as.character(lon))) # regrid into 2x2º grid

rm(OceanSODA_temp)

```

## Apply region masks

```{r apply_region_masks_OceanSODA_temp, fig.asp=0.4}

# keep only Southern Ocean data
OceanSODA_temp_2x2_SO <- inner_join(OceanSODA_temp_2x2, region_masks_all_2x2)
rm(OceanSODA_temp_2x2)

# add in basin separations
OceanSODA_temp_2x2_SO <- inner_join(OceanSODA_temp_2x2_SO, basinmask_2x2)
# expected number of rows from -30 to -70º latitude, 360º longitude, for 12 months, 8 years:
# 40 lat x 360 lon x 12 months x 8 years = 1 382 400 rows 
# actual number of rows: 925 260 (in line with expectations)
# original 1x1: 

OceanSODA_temp_2x2_SO <- OceanSODA_temp_2x2_SO %>% 
  filter(!is.na(temperature))
# results in 925 056 observations 
# no NA clim_diff values 

```

## Maps

```{r map_OceanSODA_clim_diff, fig.asp=1.5}

OceanSODA_temp_2x2_SO %>% 
  filter(year == 2020) %>% 
  ggplot(aes(lon_raw, lat_raw, fill = clim_temp)) +
  geom_tile() +
  scale_fill_viridis_c() +
  facet_wrap(~ month, ncol = 2) +
  coord_quickmap(expand = 0)


```

```{r monthly_clim_OceanSODA_temp_map, fig.asp=0.4}

OceanSODA_temp_2x2_SO %>% 
  group_split(month) %>% 
  #head(1) %>% 
  map(
    ~map +
      geom_tile(data = .x,
                aes(x = lon_raw,
                    y = lat_raw,
                    fill = clim_temp))+
      scale_fill_viridis_c()+
      labs(title = paste0('month:', unique(.x$month)))+
      theme(legend.position = 'right')
  )
  
```

```{r OceanSODA_temp_anomaly_map, fig.asp=1.5}
  
OceanSODA_temp_2x2_SO %>% 
  group_split(month) %>% 
  #head(1) %>% 
  map(
    ~map +
      geom_tile(data = .x,
                aes(x = lon_raw,
                    y = lat_raw,
                    fill = clim_diff))+
      scale_fill_divergent(mid = 'grey80')+
      facet_wrap(~year, ncol = 2)+
      labs(title = paste0('month:', unique(.x$month)))+
      theme(legend.position = 'right')
  )

```

# SST anomalies

Through setting only one of the following two code chunks to "eval=FALSE" you can choose between a lm- and a mean-based anomaly detection.

## Fit lm models

```{r fit_temp_linear_regression, eval=FALSE}

# fit a linear regression of OceanSODA pH against time (temporal trend)
# in each lat/lon/month grid

OceanSODA_temp_regression <- OceanSODA_temp_2x2_SO %>% 
  # filter(basin_AIP == "Indian",
  #        biome == "2",
  #        lon < 40) %>%
  nest(data = -c(lon, lat, month)) %>%
  mutate(fit = map(.x = data,
                   .f = ~ lm(clim_diff ~ year, data = .x)),
         tidied = map(.x = fit, .f = tidy),
         glanced = map(.x = fit, .f = glance),
         augmented = map(.x = fit, .f = augment)) 


OceanSODA_temp_regression_tidied <- OceanSODA_temp_regression %>%
  select(-c(data, fit, augmented, glanced)) %>%
  unnest(tidied) 

OceanSODA_temp_regression_tidied <- OceanSODA_temp_regression_tidied %>% 
  select(lon:estimate) %>% 
  pivot_wider(names_from = term,
              values_from = estimate) %>% 
  rename(intercept = `(Intercept)`,
         slope = year)

OceanSODA_temp_regression_augmented <- OceanSODA_temp_regression %>%
  select(-c(fit, tidied, glanced, data)) %>%
  unnest(augmented) %>% 
  select(lon:year, .resid) 

OceanSODA_temp_regression_data <- OceanSODA_temp_regression %>% 
  select(-c(fit, tidied, glanced, augmented)) %>% 
  unnest(data) 

OceanSODA_temp_regression_augmented <- bind_cols(
  OceanSODA_temp_regression_augmented,
  OceanSODA_temp_regression_data %>% 
    select(date,
           biome, basin_AIP, 
           temperature, clim_temp, 
           lat_raw, lon_raw))

OceanSODA_temp_regression_glanced <- OceanSODA_temp_regression %>%
  select(-c(data, fit, tidied, augmented)) %>%
  unnest(glanced) 
```

## Fit mean

```{r fit_temp_mean}

# identify the mean value
# in each lat/lon/month grid

OceanSODA_temp_regression_tidied <- OceanSODA_temp_2x2_SO %>% 
  # filter(basin_AIP == "Indian",
  #        biome == "2",
  #        lon < 40) %>%
  group_by(lon, lat, month) %>%
  summarise(slope = 0,
            intercept = mean(clim_diff, na.rm = TRUE)) %>% 
  ungroup()

OceanSODA_temp_regression_glanced <- OceanSODA_temp_2x2_SO %>% 
  # filter(basin_AIP == "Indian",
  #        biome == "2",
  #        lon < 40) %>%
  group_by(lon, lat, month) %>%
  summarise(sigma = sd(clim_diff, na.rm = TRUE)) %>% 
  ungroup()

OceanSODA_temp_regression_augmented <- OceanSODA_temp_2x2_SO %>% 
  # filter(basin_AIP == "Indian",
  #        biome == "2",
  #        lon < 40) %>%
  mutate(.resid = clim_diff)

```

## Slope maps

```{r map_temp_regression_slopes_flat, fig.asp=1}

map+
  geom_tile(data = OceanSODA_temp_regression_tidied,
            aes(x = lon,
                y = lat, 
                fill = slope))+
  scale_fill_scico(palette = 'vik', midpoint = 0)+
  facet_wrap(~month, ncol = 2)

```

## Sigma maps

```{r temp_map_sigma, fig.asp=1}

map+
  geom_tile(data = OceanSODA_temp_regression_glanced,
            aes(x = lon,
                y = lat, 
                fill = sigma))+
  scale_fill_viridis_c()+
  facet_wrap(~month, ncol = 2)+
  labs(fill = '1 residual \nst. dev.')

```

## Anomaly identification

Calculate OceanSODA surface temperature anomalies; L for abnormally low, H for abnormally high, and N for normal

```{r calculate_extreme_OceanSODA_temp_grid}

# when the in-situ OceanSODA temperature is lower than the 5th percentile (predicted - 2*residual.st.dev), assign 'L' for low extreme
# when the in-situ OceanSODA temperature exceeds the 95th percentile (predicted + 2*residual.st.dev), assign 'H' for high extreme
# when the in-situ OceanSODA temperature is within 95% of the range, then assign 'N' for normal pH

# combine observations and regression statistics

OceanSODA_temp_SO_extreme_grid <-
  full_join(
    OceanSODA_temp_regression_augmented,
    OceanSODA_temp_regression_glanced %>%
      select(lon:month, sigma)
  )
# results in 925 056 rows 

# identify observations in anomaly classes

OceanSODA_temp_SO_extreme_grid <- OceanSODA_temp_SO_extreme_grid %>%
  mutate(
    temp_extreme = case_when(
      .resid < -sigma*2 ~ 'L',
      .resid > sigma*2 ~ 'H',
      TRUE ~ 'N'
    )
  ) 

OceanSODA_temp_SO_extreme_grid <- OceanSODA_temp_SO_extreme_grid %>%
  mutate(temp_extreme = fct_relevel(temp_extreme, "H", "N", "L"))


# combine with regression coefficients

OceanSODA_temp_SO_extreme_grid <-
  full_join(OceanSODA_temp_SO_extreme_grid,
            OceanSODA_temp_regression_tidied) 
# 925 056 rows, in line with expectations for 40 lat x 360 lon x 12 months x 8 years (1 382 400 obs minus NA values)
```

```{r temp_anomaly_test_plot_grid, fig.asp=0.4}

OceanSODA_temp_SO_extreme_grid %>%
  group_split(lon, lat, month) %>%
  head(6) %>%
  map(~ ggplot(data = .x) +
        geom_point(aes(x = year,
                       y = clim_diff,
                       col = temp_extreme)) +
        geom_abline(data = .x, aes(slope = slope,
                    intercept = intercept)) +
        geom_abline(data = .x, aes(slope = slope,
                    intercept = intercept + 2*sigma),
                    linetype = 2) +
        geom_abline(data = .x, aes(slope = slope,
                    intercept = intercept - 2*sigma),
                    linetype = 2) +
        labs(title = paste(fititle = paste(
          "lon:", unique(.x$lon),
          "| lat:", unique(.x$lat),
          "| month:", unique(.x$month)
          ))) +
        scale_color_manual(values = HNL_colors))
```

## Anomaly maps

```{r map_temp_extremes_1x1_plot, fig.asp=1.5}
# anomaly maps on a 1x1 grid 

# OceanSODA_temp_SO_extreme_grid %>%
#   group_split(year) %>%
#   # head(1) %>%
#   map(
#     ~map +
#       geom_tile(data = .x,
#                 aes(x = lon_raw,
#                     y = lat_raw,
#                     fill = temp_extreme),
#                 width = 1,
#                 height = 1)+
#       scale_fill_manual(values = HNL_colors_map)+
#       facet_wrap(~month, ncol = 2)+
#       labs(title = paste('Year:', unique(.x$year)),
#            fill = 'temperature')
#   )

OceanSODA_temp_SO_extreme_grid %>% 
  group_split(month) %>% 
  #head(1) %>% 
  map(
    ~map +
      geom_tile(data = .x,
                aes(x = lon_raw,
                    y = lat_raw,
                    fill = temp_extreme),
                width = 1,
                height = 1)+
      scale_fill_manual(values = HNL_colors_map)+
      facet_wrap(~year, ncol = 2)+
      labs(title = paste('month:', unique(.x$month)),
           fill = 'temperature')
  )
```

## Anomaly timeseries

```{r extreme_temp_timeseries_grid}
# calculate a regional mean temperature for each biome, basin, and ph extreme (H/L/N) and plot a timeseries 

OceanSODA_temp_SO_extreme_grid %>% 
  group_by(year, biome, basin_AIP, temp_extreme) %>% 
  summarise(temp_regional = mean(temperature, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot(aes(x = year, y = temp_regional, col = temp_extreme))+
  geom_point(size = 0.3)+
  geom_line()+
  scale_color_manual(values = HNL_colors) +
  facet_grid(basin_AIP~biome)+
  theme(legend.position = 'bottom')
```

## Anomaly histogram

```{r extreme_temp_histogram_grid}

OceanSODA_temp_SO_extreme_grid %>%
  ggplot(aes(temperature, col = temp_extreme)) +
  geom_density() +
  scale_color_manual(values = HNL_colors) +
  facet_grid(basin_AIP ~ biome) +
  coord_cartesian(xlim = c(-2, 28)) +
  labs(x = 'value',
       y = 'density',
       col = 'temp anomaly') +
  theme(legend.position = 'bottom')
```

# Argo

## Grid reduction

```{r grid_reduction_to_2x2_argo}

# Note: While reducing lon x lat grid,
# we keep the original number of observations

full_argo_2x2 <- full_argo %>%
  mutate(
    lat_raw = lat,
    lon_raw = lon,
    lat = cut(lat, seq(-90, 90, 2), seq(-89, 89, 2)),
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 2), seq(21, 379, 2)),
    lon = as.numeric(as.character(lon)))  # re-grid to 2x2


```

## Apply region masks

```{r apply_region_masks_argo, eval=FALSE}

# keep only Southern Ocean argo data
full_argo_2x2_SO <- inner_join(full_argo_2x2, region_masks_all_2x2)

# add in basin separations
full_argo_2x2_SO <- inner_join(full_argo_2x2_SO, basinmask_2x2)

```

# Join OceanSODA

```{r combine_extreme_temp_OceanSODA_to_Argo}

# revert OceanSODA to regular 1x1 grid
OceanSODA_temp_SO_extreme_grid <- OceanSODA_temp_SO_extreme_grid %>%
  select(-c(lon, lat)) %>%
  rename(OceanSODA_temp = temperature,
         lon = lon_raw,
         lat = lat_raw)
# 925 056 obs 

# combine the argo profile data to the surface extreme data
profile_temp_extreme <- inner_join(
  full_argo %>% 
    select(c(year, month, date, lon, lat, depth,
           temp_adjusted,
           platform_number,
           cycle_number)),                 # 567 327 obs 
  OceanSODA_temp_SO_extreme_grid %>% 
    select(c(year, month, date, lon, lat,
           OceanSODA_temp, temp_extreme,
           clim_temp, clim_diff,
           biome, basin_AIP)))

profile_temp_extreme <- profile_temp_extreme %>% 
  unite('platform_cycle', platform_number:cycle_number, sep = '_', remove = FALSE)
  

```

# Plot profiles

Argo profiles plotted according to the surface OceanSODA temperature

L profiles correspond to a low surface temperature event, as recorded in OceanSODA

H profiles correspond to an event of high surface temperature, as recorded in OceanSODA

N profiles correspond to normal surface OceanSODA temperature

## Raw

```{r temp_profiles_raw}

profile_temp_extreme %>%
  group_split(biome, basin_AIP, year) %>% 
  # head(1) %>%
  map(
    ~ ggplot() +
      geom_path(data = .x %>% filter(temp_extreme == 'N'),
                aes(x = temp_adjusted, 
                    y = depth,
                    group = platform_cycle,
                    col = temp_extreme),
                size = 0.3) +
      geom_path(data = .x %>% filter(temp_extreme == 'H' | temp_extreme == 'L'),
                aes(x = temp_adjusted,
                    y = depth,
                    group = platform_cycle,
                    col = temp_extreme),
                size = 0.5)+
      scale_y_reverse() +
      scale_color_manual(values = HNL_colors) +
      facet_wrap(~ month, ncol = 6) +
      labs(
        x = 'Argo temperature (ºC)',
        y = 'depth (m)',
        title = paste(
          unique(.x$basin_AIP),
          "|",
          unique(.x$year),
          "| biome:",
          unique(.x$biome)
        ),
        col = 'OceanSODA temp \nanomaly'
      )
  )

```

### Atl, b1, Oct17

```{r select_specific_temp_profiles_2017}

# Temperature extreme: 
# Atlantic biome 1, 2018, months 2 and 3 

OceanSODA_temp_SO_extreme_grid_2017 <- OceanSODA_temp_SO_extreme_grid %>% 
  filter(date == '2017-10-15')  

map+
  geom_tile(data = OceanSODA_temp_SO_extreme_grid_2017,
            aes(x = lon,
                y = lat,
                fill = temp_extreme))+
  scale_fill_manual(values = HNL_colors_map)+
  labs(title = 'October 2017',
       fill = 'OceanSODA SST \nextreme')
  
profile_temp_Atl_2017 <- profile_temp_extreme %>% 
  filter(date == '2017-10-15',
         basin_AIP == 'Atlantic',
         biome == '1') 

profile_temp_Atl_2017 %>% 
  ggplot(aes(x = temp_adjusted,
             y = depth,
             group = platform_cycle,
             col = temp_extreme))+
  geom_path(data = profile_temp_Atl_2017 %>% filter(temp_extreme == 'N'),
            aes(x = temp_adjusted,
                y = depth,
                group = platform_cycle,
                col = temp_extreme),
            size = 0.3)+
  geom_path(data = profile_temp_Atl_2017 %>% filter(temp_extreme == 'H'| temp_extreme == 'L'),
            aes(x = temp_adjusted,
                y = depth,
                group = platform_cycle,
                col = temp_extreme),
            size = 0.5)+
  scale_y_reverse()+
  scale_color_manual(values = HNL_colors)+
  labs(title = 'Atlantic, biome 1, October 2017',
       col = 'OceanSODA SST\nextreme',
       x = 'Argo temperature (ºC)')

rm(profile_temp_Atl_2017, OceanSODA_temp_SO_extreme_grid_2017)

```

### Atl, b2, Jul16

```{r select_specific_temp_profiles_2015, eval=FALSE}

# Atlantic biome 2, 2016 month 7 

OceanSODA_temp_SO_extreme_grid_2016 <- OceanSODA_temp_SO_extreme_grid %>% 
  filter(date == '2016-07-15')  

map+
  geom_tile(data = OceanSODA_temp_SO_extreme_grid_2016,
            aes(x = lon,
                y = lat,
                fill = temp_extreme))+
  scale_fill_manual(values = HNL_colors_map)+
  labs(title = 'July 2016',
       fill = 'OceanSODA SST \nextreme')

profile_temp_Atl_2016 <- profile_temp_extreme %>% 
  filter(date == '2016-07-15',
         basin_AIP == 'Atlantic',
         biome == '2') 

profile_temp_Atl_2016 %>% 
  ggplot(aes(x = temp_adjusted,
             y = depth,
             group = platform_cycle,
             col = temp_extreme))+
  geom_path(data = profile_temp_Atl_2016 %>% filter(temp_extreme == 'N'),
            aes(x = temp_adjusted,
                y = depth, 
                group = platform_cycle,
                col = temp_extreme),
            size = 0.3)+
  geom_path(data = profile_temp_Atl_2016 %>% filter(temp_extreme == 'H'|temp_extreme == 'L'),
            aes(x = temp_adjusted,
                y = depth, 
                group = platform_cycle,
                col = temp_extreme),
            size = 0.5)+
  scale_y_reverse()+
  scale_color_manual(values = HNL_colors)+
  labs(title = 'Atlantic, biome 2, July 2016',
       col = 'OceanSODA SST\nextreme',
       x = 'Argo temperature (ºC)')

rm(profile_temp_Atl_2016, OceanSODA_temp_SO_extreme_grid_2016)

```

## Averaged profiles

```{r calculate_mean_temp_profiles}
 
# cut depth levels at 10, 20, .... etc m
# add seasons 
# Dec, Jan, Feb <- summer 
# Mar, Apr, May <- autumn 
# Jun, Jul, Aug <- winter 
# Sep, Oct, Nov <- spring 

profile_temp_extreme <- profile_temp_extreme %>%
  mutate(
    depth = Hmisc::cut2(
      depth,
      cuts = c(10, 20, 30, 50, 70, 100, 300, 500, 800, 1000, 1500, 2000, 2500),
      levels.mean = TRUE,
      digits = 3
    ),
    depth = as.numeric(as.character(depth))
  ) %>%
  mutate(
    season = case_when(
      between(month, 3, 5) ~ 'autumn',
      between(month, 6, 8) ~ 'winter',
      between(month, 9, 11) ~ 'spring',
      month == 12 | 1 | 2 ~ 'summer'
    ),
    .after = date
  )

```

### Overall mean

```{r mean_temp_profiles, fig.asp=1}

profile_temp_extreme_mean <- profile_temp_extreme %>%
  group_by(temp_extreme, depth) %>%
  summarise(temp_mean = mean(temp_adjusted, na.rm = TRUE),
            temp_std = sd(temp_adjusted, na.rm = TRUE)) %>%
  ungroup()

profile_temp_extreme_mean %>%
  arrange(depth) %>%
  ggplot(aes(y = depth)) +
  geom_ribbon(aes(xmin = temp_mean - temp_std,
                  xmax = temp_mean + temp_std,
                  fill = temp_extreme), 
              alpha = 0.2)+
  geom_path(aes(x = temp_mean,
                col = temp_extreme))+
  scale_color_manual(values = HNL_colors) +
  scale_fill_manual(values = HNL_colors)+
  labs(title = "Overall mean",
       col = 'OceanSODA\ntemp anomaly \n(mean ± st dev)',
       fill = 'OceanSODA\ntemp anomaly \n(mean ± st dev)',
       y = 'sqrt(depth)',
       x = 'mean Argo temperature (ºC)') +
  scale_y_continuous(trans = trans_reverser("sqrt"),
                     breaks = c(10, 100, 250, 500, seq(1000, 5000, 500)))

rm(profile_temp_extreme_mean)

```

Number of profiles

```{r mean_temp_profile_count, fig.asp=1}

profile_temp_count_mean <- profile_temp_extreme %>% 
  distinct(temp_extreme, platform_number, cycle_number) %>% 
  count(temp_extreme)

profile_temp_count_mean %>% 
  ggplot(aes(x = temp_extreme, y = n, fill = temp_extreme))+
  geom_col(width = 0.5)+
  scale_y_continuous(trans = 'log10')+
  labs(y = 'log(number of profiles)',
       title = 'Number of profiles')

rm(profile_temp_count_mean)
```

Surface Argo temperature vs surface OceanSODA temperature (20 m)

```{r argo_vs_OceanSODA_temp_mean, fig.asp=1}
# calculate surface-mean argo SST, for each profile 
surface_temp_mean <- profile_temp_extreme %>% 
  filter(depth <= 20) %>% 
  group_by(temp_extreme, platform_number, cycle_number) %>% 
  summarise(argo_surf_temp = mean(temp_adjusted, na.rm = TRUE),
            OceanSODA_surf_temp = mean(OceanSODA_temp, na.rm = TRUE))

surface_temp_mean %>% 
  group_by(temp_extreme) %>%
  group_split() %>% 
  # head(1) %>%
  map(
  ~ggplot(data = .x, aes(x = OceanSODA_surf_temp, 
             y = argo_surf_temp))+
  geom_bin2d(data = .x, aes(x = OceanSODA_surf_temp, 
                 y = argo_surf_temp), size = 0.3, bins = 60) +
  scale_fill_viridis_c()+
  geom_abline(slope = 1, intercept = 0)+
  coord_fixed(ratio = 1,
              xlim = c(-3, 28),
              ylim = c(-3, 28))+
    labs(title = paste('temp extreme:', unique(.x$temp_extreme)),
         x = 'OceanSODA temp',
         y = 'Argo temp')
  )

rm(surface_temp_mean)

```

### January overall mean profiles

```{r jan_overall_mean_temp_profiles, fig.asp=1}

profile_temp_extreme_mean_jan <- profile_temp_extreme %>%
  filter(month == 1) %>% 
  group_by(temp_extreme, depth) %>%
  summarise(temp_mean = mean(temp_adjusted, na.rm = TRUE),
            temp_std = sd(temp_adjusted, na.rm = TRUE)) %>%
  ungroup()

profile_temp_extreme_mean_jan %>%
  arrange(depth) %>%
  ggplot(aes(y = depth)) +
  geom_ribbon(aes(xmin = temp_mean - temp_std,
                  xmax = temp_mean + temp_std,
                  fill = temp_extreme), 
              alpha = 0.2)+
  geom_path(aes(x = temp_mean,
                col = temp_extreme))+
  scale_color_manual(values = HNL_colors) +
  scale_fill_manual(values = HNL_colors)+
  labs(title = "January mean",
       col = 'OceanSODA\ntemp anomaly \n(mean ± st dev)',
       fill = 'OceanSODA\ntemp anomaly \n(mean ± st dev)',
       y = 'sqrt(depth)',
       x = 'mean Argo temperature (ºC)') +
  scale_y_continuous(trans = trans_reverser("sqrt"),
                     breaks = c(10, 100, 250, 500, seq(1000, 5000, 500)))

rm(profile_temp_extreme_mean_jan)


```

### Season x biome

```{r mean_temp_profiles_season_biome, fig.asp=1}

profile_temp_extreme_biome <- profile_temp_extreme %>% 
  group_by(season, biome, temp_extreme, depth) %>% 
  summarise(temp_biome = mean(temp_adjusted, na.rm = TRUE),
            temp_std_biome = sd(temp_adjusted, na.rm = TRUE)) %>% 
  ungroup()
  

profile_temp_extreme_biome %>%
  ggplot(aes(
    x = temp_biome,
    y = depth,
    group = temp_extreme,
    col = temp_extreme
  )) +
  geom_ribbon(aes(xmax = temp_biome + temp_std_biome,
                  xmin = temp_biome - temp_std_biome,
                  group = temp_extreme,
                  fill = temp_extreme),
              col = NA, 
              alpha = 0.2)+
  geom_path() +
  scale_color_manual(values = HNL_colors) +
  scale_fill_manual(values = HNL_colors)+
  labs(col = 'OceanSODA\ntemp anomaly \n(mean ± st dev)',
       fill = 'OceanSODA\ntemp anomaly \n(mean ± st dev)',
       y = 'sqrt(depth)',
       x = 'biome mean Argo temperature (ºC)') +
  scale_y_continuous(trans = trans_reverser("sqrt"),
                     breaks = c(10, 100, 250, 500, seq(1000, 5000, 500))) +
  lims(x = c(-3, 18))+
  facet_grid(season ~ biome)

rm(profile_temp_extreme_biome)

```

Number of profiles season x biome

```{r count_temp_profiles_season_biome, fig.asp=1}

profile_temp_count_biome <- profile_temp_extreme %>% 
  distinct(season, biome, temp_extreme, platform_number, cycle_number) %>% 
  group_by(season, biome, temp_extreme) %>% 
  count(temp_extreme)

profile_temp_count_biome %>% 
  ggplot(aes(x = temp_extreme, y = n, fill = temp_extreme))+
  geom_col(width = 0.5)+
  facet_grid(season ~ biome)+
  scale_y_continuous(trans = 'log10')+
  labs(y = 'log(number of profiles)',
       title = 'Number of profiles season x biome')

rm(profile_temp_count_biome)
```

Surface Argo temp vs surface OceanSODA temp season x biome (20 m)

```{r argo_vs_OceanSODA_temp_biome, fig.asp=1}

surface_temp_biome <- profile_temp_extreme %>% 
  filter(depth <= 20) %>% 
  group_by(season, biome, temp_extreme, platform_number, cycle_number) %>% 
  summarise(argo_surf_temp = mean(temp_adjusted, na.rm=TRUE),
            OceanSODA_surf_temp = mean(OceanSODA_temp, na.rm = TRUE))

surface_temp_biome %>% 
  group_by(temp_extreme) %>% 
  group_split(temp_extreme) %>% 
  map(
  ~ggplot(data = .x, aes(x = OceanSODA_surf_temp, 
             y = argo_surf_temp))+
  geom_bin2d(data = .x, aes(x = OceanSODA_surf_temp, 
                 y = argo_surf_temp)) +
  scale_fill_viridis_c()+
  geom_abline(slope = 1, intercept = 0)+
  coord_fixed(ratio = 1, 
              xlim = c(-3, 25),
              ylim = c(-3, 25))+
  facet_grid(season~biome) +
    labs(title = paste( 'Temp extreme:', unique(.x$temp_extreme)),
         x = 'OceanSODA temp',
         y = 'Argo temp')
  )

rm(surface_temp_biome)

```

### Season x basin

```{r mean_basin_temp_profiles, fig.asp=1}

profile_temp_extreme_basin <- profile_temp_extreme %>% 
  group_by(season, basin_AIP, temp_extreme, depth) %>% 
  summarise(temp_basin = mean(temp_adjusted, na.rm = TRUE),
            temp_basin_std = sd(temp_adjusted, na.rm = TRUE)) %>% 
  ungroup()

profile_temp_extreme_basin %>% 
  ggplot(aes(x = temp_basin, 
             y = depth, 
             group = temp_extreme, 
             col = temp_extreme))+
  geom_ribbon(aes(xmin = temp_basin - temp_basin_std,
                  xmax = temp_basin + temp_basin_std,
                  group = temp_extreme, 
                  fill = temp_extreme),
              col = NA, 
              alpha = 0.2)+
  geom_path()+
  scale_color_manual(values = HNL_colors)+
  scale_fill_manual(values = HNL_colors)+
  labs(col = 'OceanSODA\ntemp anomaly\n(mean ± st dev)',
       fill = 'OceanSODA\ntemp anomaly\n(mean ± st dev)',
       y = 'sqrt(depth)',
       x = 'basin-mean Argo temperature (ªC)')+
  scale_y_continuous(trans = trans_reverser("sqrt"),
                     breaks = c(10, 100, 250, 500, seq(1000, 5000, 500))) +
  facet_grid(season~basin_AIP)

rm(profile_temp_extreme_basin)
```

Number of profiles season x basin

```{r count_basin_temp_profiles, fig.asp=1}

profile_temp_count_basin <- profile_temp_extreme %>% 
  distinct(season, basin_AIP, temp_extreme, platform_number, cycle_number) %>% 
  group_by(season, basin_AIP, temp_extreme) %>% 
  count(temp_extreme)

profile_temp_count_basin %>% 
  ggplot(aes(x = temp_extreme, y = n, fill = temp_extreme))+
  geom_col(width = 0.5)+
  facet_grid(season~basin_AIP)+
  scale_y_continuous(trans = 'log10')+
  labs(y = 'log(number of profiles)',
       title = 'Number of profiles season x basin')

rm(profile_temp_count_basin)

```

Surface Argo temperature vs surface OceanSODA temperature (20 m) season x basin

```{r argo_vs_OceanSODA_temp_basin, fig.asp=1}
# calculate surface-mean argo pH to compare against OceanSODA surface pH (one value)
surface_temp_basin <- profile_temp_extreme %>% 
  filter(depth <= 20) %>% 
  group_by(season, basin_AIP, temp_extreme, platform_number, cycle_number) %>% 
  summarise(surf_argo_temp = mean(temp_adjusted, na.rm=TRUE),
            surf_OceanSODA_temp = mean(OceanSODA_temp, na.rm = TRUE)) 

surface_temp_basin %>% 
  group_by(temp_extreme) %>% 
  group_split(temp_extreme) %>% 
  map(
  ~ggplot(data = .x, aes(x = surf_OceanSODA_temp, 
             y = surf_argo_temp))+
  geom_bin2d(data = .x, aes(x = surf_OceanSODA_temp, 
                 y = surf_argo_temp)) +
    scale_fill_viridis_c()+
  geom_abline(slope = 1, intercept = 0)+
  coord_fixed(ratio = 1, 
              xlim = c(-3, 25),
              ylim = c(-3, 25))+
  facet_grid(season~basin_AIP) +
    labs(title = paste('Temp extreme:', unique(.x$temp_extreme)),
         x = 'OceanSODA temp',
         y = 'Argo temp')
  )

rm(surface_temp_basin)

```

### Season x biome x basin

```{r mean_temp_profiles_season_biome_basin, fig.asp=1}

profile_temp_extreme_season <- profile_temp_extreme %>%
  group_by(season, biome, basin_AIP, temp_extreme, depth) %>%
  summarise(temp_mean = mean(temp_adjusted, na.rm = TRUE),
            temp_std = sd(temp_adjusted, na.rm = TRUE)) %>%
  ungroup()

profile_temp_extreme_season %>%
  arrange(depth) %>%
  group_split(season) %>%
  # head(1) %>%
  map(
    ~ ggplot(
      data = .x,
      aes(x = temp_mean,
          y = depth,
          group = temp_extreme,
          col = temp_extreme)) +
      geom_ribbon(aes(xmax = temp_mean + temp_std,
                      xmin = temp_mean - temp_std,
                      group = temp_extreme,
                      fill = temp_extreme),
                  col = NA,
                  alpha = 0.2)+
      geom_path() +
      scale_color_manual(values = HNL_colors) +
      scale_fill_manual(values = HNL_colors) +
      labs(title = paste("season:", unique(.x$season)),
           col = 'OceanSODA\ntemp anomaly \n(mean ± st dev)',
           fill = 'OceanSODA\ntemp anomaly \n(mean ± st dev)',
           y = 'sqrt(depth)',
           x = 'mean Argo temperature (ºC)') +
      scale_y_continuous(
        trans = trans_reverser("sqrt"),
        breaks = c(10, 100, 250, 500, seq(1000, 5000, 500))
      ) +
      facet_grid(basin_AIP ~ biome)
  )

```

Number of profiles season x biome x basin

```{r number_season_biome_basin_temp_profiles, fig.asp=1}

profile_temp_count_season <- profile_temp_extreme %>% 
  distinct(season, biome, basin_AIP,
           temp_extreme, platform_number, cycle_number) %>% 
  group_by(season, biome, basin_AIP, temp_extreme) %>% 
  count(temp_extreme)


profile_temp_count_season %>% 
  group_by(season) %>% 
  group_split(season) %>% 
  map(
    ~ggplot()+
      geom_col(data =.x, 
               aes(x = temp_extreme,
                   y = n,
                   fill = temp_extreme),
               width = 0.5)+
      facet_grid(basin_AIP ~ biome)+
      scale_y_continuous(trans = 'log10')+
      labs(y = 'log(number of profiles)',
           title = paste('season:', unique(.x$season)))
  )

rm(profile_temp_count_season)

```

Surface Argo temperature vs surface OceanSODA temperature (20 m) season x biome x basin

```{r surface_OceanSODA_vs_argo_temp_season, fig.asp=1}

# calculate surface-mean argo pH, for each season x biome x basin x ph extreme 
surface_temp_season <- profile_temp_extreme %>% 
  filter(depth <= 20) %>% 
  group_by(season, 
           basin_AIP, 
           biome, 
           temp_extreme,  
           platform_number, 
           cycle_number) %>%  
  summarise(surf_argo_temp = mean(temp_adjusted, na.rm=TRUE),
            surf_OceanSODA_temp = mean(OceanSODA_temp, na.rm = TRUE)) 

surface_temp_season %>% 
  group_by(season, temp_extreme) %>% 
  group_split(season, temp_extreme) %>% 
  map(
  ~ggplot(data = .x, aes(x = surf_OceanSODA_temp, 
             y = surf_argo_temp))+
  geom_bin2d(data = .x, aes(x = surf_OceanSODA_temp, 
                 y = surf_argo_temp)) +
  scale_fill_viridis_c()+
  geom_abline(slope = 1, intercept = 0)+
  coord_fixed(ratio = 1, 
              xlim = c(-3, 25),
              ylim = c(-3, 25))+
  facet_grid(basin_AIP ~ biome) +
    labs(title = paste('season:', unique(.x$season), 
                        '| temp extreme:', unique(.x$temp_extreme)),
         x = 'OceanSODA temp',
         y = 'Argo temp')
  )

rm(surface_temp_season)

```

##### Atl, b2, winter

```{r temp_profile_season_biome_Atlantic, fig.asp=1}

profile_temp_extreme_season %>%
  filter(basin_AIP == 'Atlantic',
         biome == '2',
         season == 'winter') %>% 
  arrange(depth) %>%
  ggplot(aes(x = temp_mean,
             y = depth,
             group = temp_extreme,
             col = temp_extreme)) +
  geom_ribbon(aes(xmax = temp_mean + temp_std,
                  xmin = temp_mean - temp_std,
                  group = temp_extreme,
                  fill = temp_extreme),
              col = NA,
              alpha = 0.2)+
  geom_path() +
  scale_color_manual(values = HNL_colors) +
  scale_fill_manual(values = HNL_colors) +
  labs(title = 'Atlantic basin, biome 2, winter',
       col = 'OceanSODA\ntemp anomaly \n(mean ± st dev)',
       fill = 'OceanSODA\ntemp anomaly \n(mean ± st dev)',
       y = 'sqrt(depth)',
       x = 'mean Argo temperature (ºC)') +
  scale_y_continuous(
  trans = trans_reverser("sqrt"),
  breaks = c(10, 100, 250, 500, seq(1000, 5000, 500))) 
```

##### Atl, b1, spring

```{r Atlantic_biome_1_spring_mean_profiles, fig.asp=1}

profile_temp_extreme_season %>%
  filter(basin_AIP == 'Atlantic',
         biome == '1',
         season == 'spring') %>% 
  arrange(depth) %>%
  ggplot(aes(x = temp_mean,
             y = depth,
             group = temp_extreme,
             col = temp_extreme)) +
  geom_ribbon(aes(xmax = temp_mean + temp_std,
                  xmin = temp_mean - temp_std,
                  group = temp_extreme,
                  fill = temp_extreme),
              col = NA,
              alpha = 0.2)+
  geom_path() +
  scale_color_manual(values = HNL_colors) +
  scale_fill_manual(values = HNL_colors) +
  labs(title = 'Atlantic basin, biome 1, spring',
       col = 'OceanSODA\ntemp anomaly \n(mean ± st dev)',
       fill = 'OceanSODA\ntemp anomaly \n(mean ± st dev)',
       y = 'sqrt(depth)',
       x = 'mean Argo temperature (ºC)') +
  scale_y_continuous(
  trans = trans_reverser("sqrt"),
  breaks = c(10, 100, 250, 500, seq(1000, 5000, 500))) 

rm(profile_temp_extreme_season)

```

### Biome-basin January mean profiles

```{r biome_basin_january_mean_profiles, fig.asp=1}

profile_temp_extreme_biome_basin_jan <- profile_temp_extreme %>%
  filter(month == 1) %>% 
  group_by(biome, basin_AIP, temp_extreme, depth) %>%
  summarise(temp_mean = mean(temp_adjusted, na.rm = TRUE),
            temp_std = sd(temp_adjusted, na.rm = TRUE)) %>%
  ungroup()

profile_temp_extreme_biome_basin_jan %>%
  arrange(depth) %>%
  ggplot(aes(x = temp_mean,
             y = depth)) +
  geom_ribbon(aes(xmin = temp_mean - temp_std,
                  xmax = temp_mean + temp_std,
                  fill = temp_extreme), 
              alpha = 0.2)+
  geom_path(aes(x = temp_mean,
                col = temp_extreme))+
  facet_grid(biome~basin_AIP)+
  scale_color_manual(values = HNL_colors) +
  scale_fill_manual(values = HNL_colors)+
  labs(title = "Basin-biome-mean January profiles",
       col = 'OceanSODA\ntemp anomaly \n(mean ± st dev)',
       fill = 'OceanSODA\ntemp anomaly \n(mean ± st dev)',
       y = 'sqrt(depth)',
       x = 'mean Argo temperature (ºC)') +
  scale_y_continuous(trans = trans_reverser("sqrt"),
                     breaks = c(10, 100, 250, 500, seq(1000, 5000, 500)))

```

# Remove climatology

Plot the H/L/N profiles as anomalies relative to the CSIO-MNR Argo temperature climatology

## Argo profiles

```{r filter_january_profiles}

# # keep only january profiles 
profile_temp_extreme_jan <- profile_temp_extreme %>%
  filter(month == 1)

```

```{r average_depth_levels}

profile_temp_extreme_jan_binned <- profile_temp_extreme_jan %>%
  group_by(lon, lat, year, month, platform_cycle,
           biome, basin_AIP, temp_extreme,
           depth) %>%
  summarize(temp_adjusted_binned = mean(temp_adjusted, na.rm = TRUE)) %>%
  ungroup()

```

## ARGO climatology

```{r load_jan_temp_climatology}

boa_clim_argo_temp_jan <- read_rds(file = paste0(path_argo_preprocessed, '/boa_clim_argo_temp_jan.rds'))

# compatibility with profile_temp_extreme_jan
boa_clim_argo_temp_jan_SO <- boa_clim_argo_temp_jan %>% 
  filter(lat < -30) %>% 
  mutate(lon = if_else(lon < 20, lon +360, lon),
         depth_boa = swDepth(pressure = pres, latitude = lat),
         month = 1) %>% 
  rename(clim_temp_jan = temp) %>% 
  select(-c(salt, time, pres))

# grid average climatological temp into the argo depth bins 
boa_clim_argo_temp_jan_SO <- boa_clim_argo_temp_jan_SO %>%
  mutate(
    depth = cut(
      depth_boa,
      breaks = c(0, 10, 20, 30, 50, 70, 100, 300, 500, 800, 1000, 1500, 2000),
      include.lowest = TRUE,
      labels = as.factor(unique(profile_temp_extreme_jan$depth))
    ),
    depth = as.numeric(as.character(depth))
  )


# calculate mean climatological pH per depth bin
boa_clim_argo_temp_jan_SO_binned <- boa_clim_argo_temp_jan_SO %>% 
  group_by(lon, lat, depth, month) %>% 
  summarise(clim_temp_jan_binned = mean(clim_temp_jan, na.rm = TRUE)) %>%
  ungroup()


# join climatology and ARGO profiles

remove_clim_jan <- inner_join(profile_temp_extreme_jan_binned,
                              boa_clim_argo_temp_jan_SO_binned)

```

## Profiles - Jan

Points are the January climatological temperature, lines are the depth-binned Argo profiles colored by H/N/L classification

### Absolute

```{r plot_raw_profiles_with_clim}

remove_clim_jan %>%
  group_split(biome, basin_AIP, year) %>%
  # head(1) %>%
  map(
    ~ ggplot() +
      geom_path(
        data = .x %>%
          filter(temp_extreme == 'N'),
        aes(
          x = temp_adjusted_binned,
          y = depth,
          group = platform_cycle,
          col = temp_extreme
        ),
        size = 0.3
      ) +
      geom_path(
        data = .x %>%
          filter(temp_extreme == 'H' | temp_extreme == 'L'),
        aes(
          x = temp_adjusted_binned,
          y = depth,
          group = platform_cycle,
          col = temp_extreme
        ),
        size = 0.5
      ) +
      geom_point(
        data = .x,
        aes(x = clim_temp_jan_binned,
            y = depth,
            col = temp_extreme),
        size = 0.5
      ) +
      scale_y_reverse() +
      scale_color_manual(values = HNL_colors) +
      labs(
        x = 'Argo temperature (ºC)',
        y = 'depth (m)',
        title = paste(
          "Biome:",
          unique(.x$biome),
          "| basin:",
          unique(.x$basin_AIP),
          " | Jan",
          unique(.x$year)
        ),
        col = 'OceanSODA temp \nanomaly'
      )
  )

```

```{r remove_climatology_from_profiles}

# calculate the difference between the binned climatological argo and in-situ argo for each depth level and grid cell 

remove_clim_jan <- remove_clim_jan %>% 
  mutate(argo_temp_anomaly = temp_adjusted_binned - clim_temp_jan_binned)

```

### Anomaly

```{r plot_january_anomaly_profiles, fig.asp=1}

remove_clim_jan %>% 
  group_split(year) %>% 
  # head(1) %>% 
  map(
    ~ggplot()+
      geom_path(data = .x %>% filter(temp_extreme == 'N'),
                aes(x = argo_temp_anomaly,
                    y = depth,
                    group = platform_cycle,
                    col = temp_extreme),
                size = 0.2)+
      geom_path(data = .x %>% filter(temp_extreme == 'H'| temp_extreme == 'L'),
                 aes(x = argo_temp_anomaly,
                     y = depth,
                     group = platform_cycle,
                     col = temp_extreme),
                 size = 0.3)+
      geom_vline(xintercept = 0)+
      scale_y_continuous(trans = trans_reverser("sqrt"),
                         breaks = c(10, 100, 250, 500, seq(1000, 5000, 500)))+
      scale_color_manual(values = HNL_colors)+
      scale_fill_manual(values = HNL_colors)+
      facet_grid(basin_AIP~biome)+
      labs(title = unique(.x$year))
  )

```

### Overall mean anomaly

```{r plot_overall_mean_clim_removed_profiles, fig.asp=1}

remove_clim_jan_overall_mean <- remove_clim_jan %>% 
  group_by(temp_extreme, depth) %>% 
  summarise(temp_anomaly_mean = mean(argo_temp_anomaly, na.rm = TRUE),
            temp_anomaly_sd = sd(argo_temp_anomaly, na.rm = TRUE))

remove_clim_jan_overall_mean %>% 
  ggplot()+
  geom_path(aes(x = temp_anomaly_mean,
                y = depth,
                group = temp_extreme,
                col = temp_extreme))+
  geom_ribbon(aes(xmax = temp_anomaly_mean + temp_anomaly_sd,
                  xmin = temp_anomaly_mean - temp_anomaly_sd,
                  y = depth,
                  group = temp_extreme,
                  fill = temp_extreme),
              col = NA,
              alpha = 0.2)+
  geom_vline(xintercept = 0)+
  scale_y_continuous(trans = trans_reverser("sqrt"),
                     breaks = c(10, 100, 250, 500, seq(1000, 5000, 500)))+
  scale_color_manual(values = HNL_colors)+
  scale_fill_manual(values = HNL_colors)+
  labs(title = 'Overall mean profiles')

rm(remove_clim_jan_overall_mean)

```

### Biome-mean anomaly

```{r plot_biome_mean_clim_removed_profiles, fig.asp=1}

remove_clim_jan_biome_mean <- remove_clim_jan %>% 
  group_by(temp_extreme, depth, biome) %>% 
  summarise(temp_anomaly_mean = mean(argo_temp_anomaly, na.rm = TRUE),
            temp_anomaly_sd = sd(argo_temp_anomaly, na.rm = TRUE))

remove_clim_jan_biome_mean %>% 
  ggplot(aes(x = temp_anomaly_mean,
                y = depth,
                group = temp_extreme,
                col = temp_extreme))+
  geom_path()+
  geom_ribbon(aes(xmax = temp_anomaly_mean + temp_anomaly_sd,
                  xmin = temp_anomaly_mean - temp_anomaly_sd,
                  group = temp_extreme,
                  fill = temp_extreme),
              col = NA,
              alpha = 0.2)+
  geom_vline(xintercept = 0)+
  scale_y_continuous(trans = trans_reverser("sqrt"),
                     breaks = c(10, 100, 250, 500, seq(1000, 5000, 500)))+
  scale_fill_manual(values = HNL_colors)+
  scale_color_manual(values = HNL_colors)+
  labs(title = 'Biome-mean profiles')+
  facet_wrap(~biome)

rm(remove_clim_jan_biome_mean)

```

### Basin-mean anomaly

```{r plot_basin_mean_anomaly_profiles, fig.asp=1}

remove_clim_jan_basin_mean <- remove_clim_jan %>% 
  group_by(basin_AIP, temp_extreme, depth) %>% 
  summarise(temp_anomaly_mean = mean(argo_temp_anomaly, na.rm = TRUE),
            temp_anomaly_sd = sd(argo_temp_anomaly, na.rm = TRUE))

remove_clim_jan_basin_mean %>% 
  ggplot(aes(x = temp_anomaly_mean,
             y = depth, 
             group = temp_extreme,
             col = temp_extreme))+
  geom_path()+
  geom_ribbon(aes(xmax = temp_anomaly_mean + temp_anomaly_sd,
                  xmin = temp_anomaly_mean - temp_anomaly_sd,
                  group = temp_extreme,
                  fill = temp_extreme),
              col = NA,
              alpha = 0.2)+
  geom_vline(xintercept = 0)+
  facet_wrap(~basin_AIP)+
  scale_y_continuous(trans = trans_reverser("sqrt"),
                     breaks = c(10, 100, 250, 500, seq(1000, 5000, 500)))+
  scale_color_manual(values = HNL_colors)+
  scale_fill_manual(values = HNL_colors)+
  labs(title = 'Basin-mean profiles')

rm(remove_clim_jan_basin_mean)
  
```

### Basin-biome-mean anomaly

```{r plot_basin_biome_mean_anomaly_profiles, fig.asp=1}

remove_clim_jan_basin_biome_mean <- remove_clim_jan %>% 
  group_by(basin_AIP, biome, temp_extreme, depth) %>% 
  summarise(temp_anomaly_mean = mean(argo_temp_anomaly, na.rm = TRUE),
            temp_anomaly_sd = sd(argo_temp_anomaly, na.rm = TRUE))

remove_clim_jan_basin_biome_mean %>% 
  ggplot(aes(x = temp_anomaly_mean,
             y = depth, 
             group = temp_extreme,
             col = temp_extreme))+
  geom_path()+
  geom_ribbon(aes(xmax = temp_anomaly_mean + temp_anomaly_sd,
                  xmin = temp_anomaly_mean - temp_anomaly_sd,
                  group = temp_extreme,
                  fill = temp_extreme),
              col = NA,
              alpha = 0.2)+
  geom_vline(xintercept = 0)+
  facet_grid(biome~basin_AIP)+
  scale_y_continuous(trans = trans_reverser("sqrt"),
                     breaks = c(10, 100, 250, 500, seq(1000, 5000, 500)))+
  scale_color_manual(values = HNL_colors)+
  scale_fill_manual(values = HNL_colors)+
  labs(title = 'Basin-biome-mean profiles')

rm(remove_clim_jan_basin_biome_mean)
  
```
