---
title: "Load Biome Information"
author: "Pasqualina Vonlanthen & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r set_options_global, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

# Task 

Load in biome information and regional separations.

Dependencies
------------
RECCAP2_region_masks_all_v20210412.nc
map_landmask_WOA18.rds

Outputs (in BGC preprocessed folder)
-------

region_masks_all_seamask_1x1.rds
region_masks_all_seamask_2x2.rds
region_masks_all_1x1.rds
region_masks_all_2x2.rds
ph_surface_1x1.rds
ph_surface_2x2.rds
nm_biomes.rds

```{r load_libraries, include=FALSE}

library(tidyverse)
library(ggspatial)
library(ggOceanMaps)
library(lubridate)

```


```{r set_updata_root_directory, include=FALSE}

path_argo <- '/nfs/kryo/work/updata/bgc_argo_r_argodata'
path_updata <- '/nfs/kryo/work/updata'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")
path_basin_mask <- "/nfs/kryo/work/updata/reccap2/"
path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"
```

# RECCAP-2 Biome Separations 

```{r load_in_biome_separations}
# load in the RECCAP biome separations 
region_masks_all <-
  stars::read_ncdf(paste(
    path_basin_mask, "RECCAP2_region_masks_all_v20210412.nc", sep = "")) %>%
  as_tibble() %>% 
  mutate(seamask = as.factor(seamask))
```

## Harmonise RECCAP biomes

```{r harmonise_reccap_biomes}

# harmonise the latitude longitude bands of the biomes to the pH data (2x2 grid)
region_masks_all_seamask_2x2 <- region_masks_all %>% 
  select(lat, lon, seamask) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon)) %>% 
  mutate(
    lat = cut(lat, seq(-90, 90, 2), seq(-89, 89, 2)),
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 2), seq(21, 379, 2)),
    lon = as.numeric(as.character(lon))
  )

region_masks_all_seamask_1x1 <- region_masks_all %>% 
  select(lat, lon, seamask) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon)) %>% 
  mutate(
    lat = cut(lat, seq(-90, 90, 1), seq(-89.5, 89.5, 1)),
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 1), seq(20.5, 379.5, 1)),
    lon = as.numeric(as.character(lon))
  )

region_masks_all <- region_masks_all %>% 
  select(-seamask) %>% 
  pivot_longer(open_ocean:southern, 
               names_to = 'region',
               values_to = 'value') %>% 
  mutate(value = as.factor(value))

# harmonise the lat/lon of the regional separations to our pH data 
region_masks_all_1x1 <- region_masks_all %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon)) %>% 
  mutate(
    lat = cut(lat, seq(-90, 90, 1), seq(-89.5, 89.5, 1)), 
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 1), seq(20.5, 379.5, 1)), 
    lon = as.numeric(as.character(lon))
)

region_masks_all_2x2 <- region_masks_all %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon)) %>% 
  mutate(
    lat = cut(lat, seq(-90, 90, 2), seq(-89, 89, 2)),
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 2), seq(21, 379, 2)),
    lon = as.numeric(as.character(lon))
  )

# add the region names to the surface pH dataframes

ph_surface_1x1 <- read_rds(file = paste0(path_argo_preprocessed, "/ph_surface_1x1.rds"))
ph_surface_2x2 <- read_rds(file = paste0(path_argo_preprocessed, "/ph_surface_2x2.rds"))

ph_surface_2x2 <- inner_join(ph_surface_2x2, region_masks_all_2x2)
ph_surface_1x1 <- inner_join(ph_surface_1x1, region_masks_all_1x1)
```

## Maps of Southern Ocean RECCAP biomes 

```{r map_coastal_regions}

map <-
  read_rds(paste(path_emlr_utilities,
                 "map_landmask_WOA18.rds",
                 sep = ""))
# restrict base map to Southern Ocean
map <- map +
  lims(y = c(-85, -30))

region_masks_all_1x1 <- region_masks_all_1x1 %>%
  filter(region == 'southern',
         value != 0) %>%
  mutate(coast = as.character(coast))

```

### Coastal regions 
```{r RECCAP_coast_map, fig.asp=0.4}

map +
  geom_tile(data = region_masks_all_1x1,
            aes(x = lon,
                y = lat,
                fill = coast))+
  scale_fill_brewer(palette = 'Dark2')

```

### Biomes 
```{r RECCAP_biome_maps, fig.asp=0.4}

map+
  geom_tile(data = region_masks_all_1x1,
            aes(x = lon,
                y = lat,
                fill = value))+
  scale_fill_brewer(palette = 'Dark2')+
  labs(title = 'RECCAP biomes')

```

```{r RECCAP_biome_polar_projection}

basemap(limits = -30)+
  geom_spatial_tile(data = region_masks_all_1x1,
                    aes(x = lon,
                        y = lat,
                        fill = value),
                    col = NA)+
  scale_fill_brewer(palette = 'Dark2')+
  labs(title = 'RECCAP biomes')

```


## Write RECCAP biomes to file
```{r write_reccap_biomes_to_file}

region_masks_all_seamask_1x1 %>% 
  write_rds(file = paste0(path_argo_preprocessed, "/region_masks_all_seamask_1x1.rds"))

region_masks_all_seamask_2x2 %>% 
  write_rds(file = paste0(path_argo_preprocessed, "/region_masks_all_seamask_2x2.rds"))

region_masks_all_1x1 %>% 
  write_rds(file = paste0(path_argo_preprocessed, "/region_masks_all_1x1.rds"))

region_masks_all_2x2 %>% 
  write_rds(file = paste0(path_argo_preprocessed, "/region_masks_all_2x2.rds"))

# joined RECCAP-biomes to surface pH data 
ph_surface_1x1 %>% 
  write_rds(file = paste0(path_argo_preprocessed, "/ph_surface_1x1.rds"))

ph_surface_2x2 %>% 
  write_rds(file = paste0(path_argo_preprocessed, "/ph_surface_2x2.rds"))

```

# Mayot biomes 

```{r load_Mayot_biomes}

nm_biomes <- tidync::hyper_tibble(paste0(path_argo, "/SouthernOcean_mask_NM.nc"))
# 1 degree lon/lat grid

# table(nm_regions$LATITUDE) # 1 degree intervals
# table((nm_regions$LONGITUDE)) # 1 degree longitude intervals
```

## Harmonise Mayot biomes
```{r harmonise_Mayot_biomes}
nm_biomes <- nm_biomes %>% 
  rename(lon = LONGITUDE,
         lat = LATITUDE) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

nm_biomes <- nm_biomes %>% 
  filter(ICE == 1 | STSS == 1 | SPSS == 1)

nm_biomes <- nm_biomes %>% 
  pivot_longer(cols = c(STSS, SPSS, ICE),
               values_to = 'biome_mask',
               names_to = 'biome_name')

nm_biomes <- nm_biomes %>% 
  filter(biome_mask==1,
         lat <= -30)
```

## Maps of Southern Ocean Mayot biomes

```{r map_Mayot_biomes, fig.asp=0.4}
map+
  geom_tile(data = nm_biomes,
            aes(x = lon,
                y = lat,
                fill = biome_name))+
  scale_fill_brewer(palette = 'Dark2')+
  labs(title = 'Mayot biomes')
```

```{r Mayot_biomes_polar_projection}

basemap(limits = -30)+
  geom_spatial_tile(data = nm_biomes,
                    aes(x = lon, 
                        y = lat, 
                        fill = biome_name),
                    col = NA)+
  scale_fill_brewer(palette = 'Dark2')+
  labs(title = 'Mayot biomes')

```


## Write Mayot biomes to file
```{r write_Mayot_biomes_to_file}
# write data to file 

nm_biomes %>% 
  write_rds(file = paste0(path_argo_preprocessed, "/nm_biomes.rds"))
```


