---
title: "UCSD Argo pH Climatology"
author: "Pasqualina Vonlanthen, David Stappard & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
chunk_output_type: console
---

# Task 

This script loads the pH climatology as described in Mazloff et al. (2023). The climatology netCDF has previously been downloaded. The lat and lon fields are harmonised to our requirements, i.e -75.5 ≥ lat ≤ -30.5 and 20.5 ≥ lon ≤ 379.5.

Mazloff, M. R., A. Verdy, S. T. Gille, K. S. Johnson, B. D. Cornuelle, and J. Sarmiento (2023), Southern Ocean Acidification Revealed by Biogeochemical-Argo Floats, Journal of Geophysical Research: Oceans, 128(5), e2022JC019530, doi:https://doi.org/10.1029/2022JC019530.

## Dependencies

pH climatology - /nfs/kryo/work/datasets/gridded/ocean/interior/observation/ph/mazloff_2023/PH-QCv3-v10r1.nc

## Outputs (in BGC preprocessed folder)

ucsd_ph_clim.rds – the pH climatology for the Southern Ocean (30.5° S and south) by 1°x1°, depths (2.1, 6.7…. 1800) and month.

```{r load_libraries}
library(tidyverse)
#library(ggOceanMaps)
library(oce)
#library(ncdf4)
library(stars)
```



```{r set_paths}
path_argo <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/bgc_argo'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")

path_mazloff_ph <-"/nfs/kryo/work/datasets/gridded/ocean/interior/observation/ph/mazloff_2023"
fn_mazloff_ph <- "PH-QCv3-v10r1.nc"
fn_mazloff_ph <- paste0(path_mazloff_ph, "/", fn_mazloff_ph)

theme_set(theme_bw())
```

# Read data 

```{r load_climatological_pH_data}

# # read pH, position and depth data
# nc_pH <- read_ncdf(fn_mazloff_ph, var = c("pH"))
# nc_pH <- as_tibble(nc_pH)
# 
# nc_lon <- read_ncdf(fn_mazloff_ph, var = c("longitude"))
# nc_lon <- as_tibble(nc_lon)
# 
# nc_lat <- read_ncdf(fn_mazloff_ph, var = c("latitude"))
# nc_lat <- as_tibble(nc_lat)
# 
# nc_depth <- read_ncdf(fn_mazloff_ph, var = c("depth"))
# nc_depth <- as_tibble(nc_depth)
# 
# nc_pH <- nc_pH %>%
#    mutate(ny = ny - 0.5,
#           nx = nx - 0.5,
#           t = t + 0.5)
# 
# # Join each attribute in turn to pH data
# clim_argo_ph <- full_join(nc_pH, nc_lat)
# clim_argo_ph <- full_join(clim_argo_ph, nc_lon)
# clim_argo_ph <- full_join(clim_argo_ph, nc_depth)
# 
# clim_argo_ph <- clim_argo_ph %>%
#    select(-c(starts_with("n")))
# 
# clim_argo_ph <- clim_argo_ph %>%
#    filter(pH != 0)
# 
# # harmonise data
# clim_argo_ph <- clim_argo_ph %>%
#   rename(lat = latitude,
#          lon = longitude,
#          month = t,
#          clim_pH = pH) %>%
#   mutate(lat = lat -0.5,
#          lon = if_else(lon < 20, lon + 360, lon))
# 
# read pH, position and depth data
nc_pH <- read_stars(fn_mazloff_ph) %>% 
  as_tibble()

nc_lat <- read_ncdf(fn_mazloff_ph, var = c("latitude")) %>% as_tibble()

nc_pH <- full_join(nc_pH %>% rename(ny = y),
                   nc_lat)

# harmonise data 
clim_argo_ph <- nc_pH %>% 
  select(-ny) %>% 
  rename(lat = latitude,
         lon = x,
         depth = nz,
         month = t,
         clim_pH = "PH-QCv3-v10r1.nc") %>% 
  mutate(depth = round(depth, 2),
         lon = if_else(lon < 20, lon + 360, lon),
         lat = lat - 0.5)
```

# Maps
```{r plots}
clim_argo_ph %>%
   filter(depth < 30) %>%
   ggplot() +
   geom_tile(aes(lon, lat, fill = clim_pH)) +
   facet_wrap(~depth) +
   scale_fill_viridis_c() +
   coord_quickmap()

clim_argo_ph %>%
   ggplot(aes(clim_pH)) +
   geom_histogram() +
   facet_wrap(~depth) +
   scale_y_log10() +
   geom_vline(xintercept = 7.5)

```


# Write data to file

```{r write_data_to_file}

clim_argo_ph %>% 
  drop_na() %>% 
  write_rds(file = paste0(path_argo_preprocessed, "/ucsd_ph_clim.rds"))

```

