---
title: "WOA O2 Climatology"
author: "David Stappard & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
chunk_output_type: console
---

# Task 

This script loads the O2 climatology as described in Garcia et al. (2018). The climatology netCDF has previously been downloaded. The lat and lon fields are harmonised to our requirements, i.e -89.5 ≥ lat ≤ 89.5 and 20.5 ≥ lon ≤ 379.5.

Garcia, H. E., K. Weathers, C. R. Paver, I. Smolyar, T. P. Boyer, R. A. Locarnini, M. M. Zweng, A. V. Mishonov, O. K. Baranova, D. Seidov, and J. R. Reagan, 2018. World Ocean Atlas 2018, Volume 3: Dissolved Oxygen, Apparent Oxygen Utilization, and Oxygen Saturation. A. Mishonov Technical Ed.; NOAA Atlas NESDIS 83, 38pp.
https://www.ncei.noaa.gov/access/world-ocean-atlas-2018/

Dependencies
------------

WOA O2 climatology - /nfs/kryo/work/datasets/gridded/ocean/interior/observation/woa/2018/oxygen/all/1.00/woa18_all_o01_01.nc

Outputs (in BGC preprocessed folder)
-------

woa_doxy_clim.rds


```{r load_libraries}
library(tidyverse)
#library(ggOceanMaps)
library(oce)
#library(ncdf4)
library(stars)
```



```{r set_paths}
path_argo <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/bgc_argo'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")

path_WOAO2 <-"/nfs/kryo/work/datasets/gridded/ocean/interior/observation/woa/2018/oxygen/all/1.00"
# monthly files of the form woa18_all_oMM_01.nc where MM = 01,....12

theme_set(theme_bw())
```

# Read data 

```{r load_climatological_data}

for (i_month in 1:12) {

  fn_WOAO2 <- paste0(path_WOAO2, "/woa18_all_o", formatC(i_month, width=2, flag="0"), "_01.nc")
  
  nc_doxy_mean <- read_ncdf(fn_WOAO2, var = c("o_an")) %>% 
    as_tibble() %>%
    select(-time)
  
  # harmonise data 
  if (!exists("clim_argo_doxy")) {
    clim_argo_doxy <- nc_doxy_mean %>%
      rename(clim_doxy = "o_an") %>%
      mutate(month = i_month,
             lon = if_else(lon < 20, lon + 360, lon))
  } else {
    clim_argo_doxy <- rbind(
      clim_argo_doxy,
      nc_doxy_mean %>%
        rename(clim_doxy = "o_an") %>%
        mutate(month = i_month,
               lon = if_else(lon < 20, lon + 360, lon))
    )
  }

}

```

# Maps
```{r plots}

clim_argo_doxy %>%
   filter(depth < 30) %>%
   ggplot() +
   geom_tile(aes(lon, lat, fill = clim_doxy)) +
   facet_wrap(~depth) +
   scale_fill_viridis_c() +
   coord_quickmap()

clim_argo_doxy %>%
   ggplot(aes(clim_doxy)) +
   geom_histogram(binwidth = 25) +
   facet_wrap(~depth) +
   scale_y_log10()

```


# Write data to file

```{r write_data_to_file}

clim_argo_doxy %>% 
  drop_na() %>% 
  write_rds(file = paste0(path_argo_preprocessed, "/woa_doxy_clim.rds"))

```

