---
title: "Map BGC Argo observations"
author: "Pasqualina Vonlanthen & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
Map the location of oxygen, pH, and nitrate observations recorded by BGC-Argo floats 

```{r loading_packages}

library(tidyverse)
# remotes::install_github("ArgoCanada/argodata")
library(argodata)
library(argoFloats)
library(ggplot2)
library(lubridate)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
# load in coastline data (uses sf and rnaturalearthdata packages)
world = ne_coastline(scale = 'medium', returnclass = 'sf')

```

Load in the data (the process is the same as in loading_data.html)

```{r loading_data}
# Set the cache directory using argo_set_cache_dir():
argo_set_cache_dir('/nfs/kryo/work/updata/bgc_argo_r_argodata')
argo_update_global(max_global_cache_age = Inf)
argo_update_data(max_data_cache_age = Inf)

bgc_subset = argo_global_synthetic_prof() %>%
  argo_filter_data_mode(data_mode = 'delayed') %>%
  argo_filter_date(date_min = '2013-01-01',
                   date_max = '2015-12-31') # download the indexes of synthetic files of delayed-mode data (BGC and core argo data)

# check the dates 
# max(bgc_subset$date, na.rm = TRUE)
# min(bgc_subset$date, na.rm = TRUE)


bgc_data = argo_prof_levels(bgc_subset, vars = c('PRES_ADJUSTED','PRES_ADJUSTED_QC',
                                                 'PSAL_ADJUSTED', 'PSAL_ADJUSTED_QC',
                                                 'TEMP_ADJUSTED','TEMP_ADJUSTED_QC',
                                                 'DOXY_ADJUSTED', 'DOXY_ADJUSTED_QC',
                                                 'NITRATE_ADJUSTED', 'NITRATE_ADJUSTED_QC',
                                                 'PH_IN_SITU_TOTAL_ADJUSTED', 'PH_IN_SITU_TOTAL_ADJUSTED_QC'), quiet = TRUE)
# read in the profiles (takes a while)

bgc_metadata = argo_prof_prof(bgc_subset)  # load in the corresponding metadata

full_data = left_join(bgc_data, bgc_metadata, by = c('file', 'n_prof'))
# joins the metadata to the data and creates one data frame
```

1. MAP OBSERVATIONS  

Create separate dataframes for each variable, with longitude, latitude, date, value, cycle number, and float ID 

1.1 Oxygen 

```{r create_oxygen_dataframe}

oxy = data.frame(full_data$longitude, full_data$latitude, full_data$date,
                 full_data$doxy_adjusted) # add longitude, latitude, date and oxygen to the dataframe 
colnames(oxy) = c('longitude','latitude','date', 'doxy_adjusted') # rename columns 

oxy = oxy %>%
  mutate(date.simple = as.Date(date),    # create one date and one time column
         time = format(date, '%H:%M:%S'),
         cycle = full_data$cycle_number, # add cycle number and float ID 
         float_ID = full_data$float_serial_no)

oxy.no.na = oxy %>%
  filter(!is.na(doxy_adjusted)) # remove the NA values

location_obs_oxy = oxy.no.na %>%
  group_by(round(longitude, digits = 0), round(latitude, digits = 0)) %>% 
  summarise(Count = n()) # count the number of data points per longitude/latitude pair, rounded to the nearest integer 

colnames(location_obs_oxy) = c('longitude', 'latitude', 'Count') # rename columns
```

Map the location of the oxygen observations 

```{r map_oxygen_observations}

ggplot() +
  geom_tile(data = location_obs_oxy, aes(x = longitude, y = latitude, fill = Count))+
  geom_sf(data = world, fill = 'grey')+
  scale_fill_gradientn(colors = rev(rainbow(10)))+
  theme_bw()+
  labs(x = 'longitude', y = 'latitude', fill = 'number of obs',
       title = 'location of delayed-mode adjusted oxygen observations')

```

1.2 pH 

```{r create_pH_dataframe}

ph = data.frame(full_data$longitude, full_data$latitude, full_data$date,
                full_data$ph_in_situ_total_adjusted) # add longitude, latitude, date, and pH to a dataframe 

colnames(ph) = c('longitude','latitude','date', 'ph_in_situ_total') #rename columns

ph = ph %>%
  mutate(date.simple = as.Date(date),  # create one date and one time columns
         time = format(date, '%H:%M:%S'),
         cycle = full_data$cycle_number,  # add cycle number and float ID
         float_ID = full_data$float_serial_no)

ph.no.na = ph %>%
  filter(!is.na(ph_in_situ_total)) # remove NA values 

location_obs_ph = ph.no.na %>%
  group_by(round(longitude, digits = 0), round(latitude, digits = 0)) %>%
  summarise(Count = n()) # count the number of pH observations for each longitude/latitude pair rounded to the nearest integer 

colnames(location_obs_ph) = c('longitude', 'latitude', 'Count') # rename columns 
```

Map the location of pH observations 

```{r map_pH_observations}

ggplot() +
  geom_sf(data = world, fill = 'grey')+
  geom_tile(data = location_obs_ph, aes(x = longitude, y = latitude, fill = Count))+
  scale_fill_gradientn(colors = rev(rainbow(10)))+
  theme_bw()+
  labs(x = 'longitude', y = 'latitude', fill = 'number of obs',
       title = 'location of delayed-mode adjusted pH observations')

```

1.3 Nitrate 

```{r create_nitrate_dataframe}

nitrate = data.frame(full_data$longitude, full_data$latitude, full_data$date,
                     full_data$nitrate_adjusted) # create a dataframe with longitude, latitude, date, and nitrate values 

colnames(nitrate) = c('longitude','latitude', 'date', 'nitrate_adjusted') # rename columns

nitrate = nitrate %>%
  mutate(date.simple = as.Date(date),  # separate date and time into two columns 
         time = format(date, '%H:%M:%S'),
         cycle = full_data$cycle_number,  # add cycle number and float ID 
         float_ID = full_data$float_serial_no)

nitrate.no.na = nitrate %>%
  filter(!is.na(nitrate_adjusted)) # remove NA values 

location_obs_nitrate = nitrate.no.na %>%
  group_by(round(longitude, digits = 0), round(latitude, digits = 0)) %>%
  summarise(Count = n()) # count the number of nitrate observations for each longitude/latitude pair rounded to the nearest integer 

colnames(location_obs_nitrate) = c('longitude', 'latitude', 'Count') # rename columns 

```

Map the location of nitrate observations 

```{r map_nitrate_observations}

ggplot() +
  geom_tile(data = location_obs_nitrate, aes(x = longitude, y = latitude, fill = Count))+
  geom_sf(data = world, fill = 'grey')+
  scale_fill_gradientn(colors = rev(rainbow(10)))+
  theme_bw()+
  labs(x = 'longitude', y = 'latitude', fill = 'number of obs',
       title = 'location of delayed-mode adjusted nitrate observations')

```

1.4 All three BGC variables 

Map the location of observations for which a measurement of all three variables exists 

```{r create_all_bgc_dataframe}
# create a dataframe which contains longitude, latitude, date, cycle number, float ID, and all three bgc variables 
bgc_co_located = data.frame(full_data$longitude, full_data$latitude,
                            full_data$date,
                            full_data$cycle_number,
                            full_data$float_serial_no,
                            full_data$doxy_adjusted,
                            full_data$ph_in_situ_total_adjusted,
                            full_data$nitrate_adjusted)

# rename columns:
colnames(bgc_co_located) = c('longitude', 'latitude',
                             'date',
                             'cycle',
                             'float_ID',
                             'doxy_adjusted',
                             'ph_in_situ_total_adjusted',
                             'nitrate_adjusted')

bgc_co_located = bgc_co_located %>%    # change the date and time format
  mutate(date.simple = as.Date(date),
         time = format(date, '%H:%M:%S'))

bgc_co_located.no.na = bgc_co_located %>%  # remove NA values for each variable
  filter(!is.na(doxy_adjusted)) %>%
  filter(!is.na(ph_in_situ_total_adjusted)) %>%
  filter(!is.na(nitrate_adjusted))

location_obs_bgc = bgc_co_located.no.na %>% 
  group_by(round(longitude, digits = 0), round(latitude, digits = 0)) %>%
  summarise(Count = n()) # count the number of observations for each longitude/latitude pair, rounded to the nearest integer

colnames(location_obs_bgc) = c('longitude', 'latitude', 'Count') # rename columns 
```

Map the locations of BGC observations 

```{r map_bgc_observations}

ggplot() +
  geom_tile(data = location_obs_bgc, aes(x = longitude, y = latitude, fill = Count))+
  geom_sf(data = world, fill = 'grey')+
  scale_fill_gradientn(colors = rev(rainbow(10)))+
  theme_bw()+
  labs(x = 'longitude', y = 'latitude', fill = 'number of obs',
       title = 'location of delayed-mode adjusted BGC observations (oxygen, pH, and nitrate)')

```

2. MAP PROFILE LOCATIONS

2.1 Oxygen 

```{r count_oxygen_profiles}

# count the number of oxygen profiles
prof_oxy = oxy.no.na %>%
  group_by(float_ID, cycle, round(longitude, digits = 0), round(latitude, digits = 0)) %>%    
  summarise(num_obs = n())  # count the number of oxygen observations by float and cycle

colnames(prof_oxy) = c('float_ID', 'cycle', 'longitude', 'latitude', 'num_obs') # rename columns

prof_oxy = prof_oxy %>%
  mutate(prof = rep(1, length(cycle))) # repeat a vector of 1s for each individual cycle 

location_prof_oxy = prof_oxy %>%
  group_by(round(longitude, digits = 0), round(latitude, digits = 0)) %>%
  summarise(count_prof = n())       # count the number of 1s (profiles) for each longitude/latitude pair, rounded to the nearest integer 

colnames(location_prof_oxy) = c('longitude', 'latitude', 'Count')

```

Map the location of oxygen profiles 

```{r map_oxygen_profiles}

ggplot() +
  geom_tile(data = location_prof_oxy, aes(x = longitude, y = latitude, fill = Count))+
  geom_sf(data = world, fill = 'grey')+
  scale_fill_gradientn(colors = rev(rainbow(10)))+
  theme_bw()+
  labs(x = 'longitude', y = 'latitude', fill = 'number of profiles',
       title = 'location of delayed-mode adjusted oxygen profiles')

```

2.2 pH 

```{r count_pH_profiles}

# count the number of pH profiles
prof_ph = ph.no.na %>%
  group_by(float_ID, cycle, round(longitude, digits = 0), round(latitude, digits = 0)) %>%    
  summarise(num_obs = n())  # count the number of ph observations by float and cycle

colnames(prof_ph) = c('float_ID', 'cycle', 'longitude', 'latitude', 'num_obs') # rename columns

prof_ph = prof_ph %>%
  mutate(prof = rep(1, length(cycle))) # repeat a vector of 1s for each individual cycle 

location_prof_ph = prof_ph %>%
  group_by(round(longitude, digits = 0), round(latitude, digits = 0)) %>%
  summarise(count_prof = n())       # count the number of 1s (profiles) for each longitude/latitude pair, rounded to the nearest integer 

colnames(location_prof_ph) = c('longitude', 'latitude', 'Count') # rename columns 

```

Map the location of pH profiles 

```{r map_pH_profiles}

ggplot() +
  geom_tile(data = location_prof_ph, aes(x = longitude, y = latitude, fill = Count))+
  geom_sf(data = world, fill = 'grey')+
  scale_fill_gradientn(colors = rev(rainbow(10)))+
  theme_bw()+
  labs(x = 'longitude', y = 'latitude', fill = 'number of profiles',
       title = 'location of delayed-mode adjusted pH profiles')

```


2.3 Nitrate 

```{r count_nitrate_profiles}

# count the number of nitrate profiles
prof_nitrate = nitrate.no.na %>%
  group_by(float_ID, cycle, round(longitude, digits = 0), round(latitude, digits = 0)) %>%    
  summarise(num_obs = n())  # count the number of nitrate observations by float and cycle

colnames(prof_nitrate) = c('float_ID', 'cycle', 'longitude', 'latitude', 'num_obs') # rename columns

prof_nitrate = prof_nitrate %>%
  mutate(prof = rep(1, length(cycle))) # repeat a vector of 1s for each individual cycle 

location_prof_nitrate = prof_nitrate %>%
  group_by(round(longitude, digits = 0), round(latitude, digits = 0)) %>%
  summarise(count_prof = n())       # count the number of 1s (profiles) for each longitude/latitude pair, rounded to the nearest integer 

colnames(location_prof_nitrate) = c('longitude', 'latitude', 'Count') # rename columns

```

Map the location of nitrate profiles 

```{r map_nitrate_profiles}

ggplot() +
  geom_tile(data = location_prof_nitrate, aes(x = longitude, y = latitude, fill = Count))+
  geom_sf(data = world, fill = 'grey')+
  scale_fill_gradientn(colors = rev(rainbow(10)))+
  theme_bw()+
  labs(x = 'longitude', y = 'latitude', fill = 'number of profiles',
       title = 'location of delayed-mode adjusted nitrate profiles')

```

2.4 Profiles which contain all three BGC variables 

```{r count_bgc_profiles}
# count the number of bgc profiles
prof_bgc = bgc_co_located.no.na %>%
  group_by(float_ID, cycle, round(longitude, digits = 0), round(latitude, digits = 0)) %>%    
  summarise(num_obs = n())  # count the number of nitrate observations by float and cycle

colnames(prof_bgc) = c('float_ID', 'cycle', 'longitude', 'latitude', 'num_obs') # rename columns

prof_bgc = prof_bgc %>%
  mutate(prof = rep(1, length(cycle))) # repeat a vector of 1s for each individual cycle 

location_prof_bgc = prof_bgc %>%
  group_by(round(longitude, digits = 0), round(latitude, digits = 0)) %>%
  summarise(count_prof = n())       # count the number of 1s (profiles) for each longitude/latitude pair, rounded to the nearest integer 

colnames(location_prof_bgc) = c('longitude', 'latitude', 'Count') # rename columns
```

Map the location of profiles containing all three BGC variables 

```{r map_bgc_profiles}

ggplot() +
  geom_tile(data = location_prof_bgc, aes(x = longitude, y = latitude, fill = Count))+
  geom_sf(data = world, fill = 'grey')+
  scale_fill_gradientn(colors = rev(rainbow(10)))+
  theme_bw()+
  labs(x = 'longitude', y = 'latitude', fill = 'number of profiles',
       title = 'location of delayed-mode adjusted BGC profiles')

```

