---
title: "OceanSODA-Argo SST"
author: "Pasqualina Vonlanthen & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Task 

Compare Argo surface temperature to OceanSODA surface temperature 

```{r set_options_global, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r load_libraries, include=FALSE}
# load in the necessary libraries
library(tidyverse)
library(lubridate)
library(ggOceanMaps)
library(metR)
```

```{r set_global_theme}
theme_set(theme_bw())
```

# Load data 

Load in surface Argo temperature, and OceanSODA temperature, on a 1ºx1º grid 

```{r set_root_directories}

path_argo <- '/nfs/kryo/work/updata/bgc_argo_r_argodata'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")
path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"
```

```{r load_argo_OceanSODA_temp_data}
# Load in surface Argo and OceanSODA temperature data 

OceanSODA_temp <- read_rds(file = paste0(path_argo_preprocessed, "/OceanSODA_temp.rds"))

argo_surf_temp <- read_rds(file = paste0(path_argo_preprocessed, "/bgc_merge_temp_qc_1.rds")) %>% 
  filter(between(depth, 0, 20)) %>% 
  mutate(year = year(date),
         month = month(date))

# for plotting later, load in region and coastline information  
region_masks_all_seamask_2x2 <- read_rds(file = paste0(
  path_argo_preprocessed, "/region_masks_all_seamask_2x2.rds"))

region_masks_all_2x2 <- read_rds(file = paste0(path_argo_preprocessed, "/region_masks_all_2x2.rds"))

region_masks_all_1x1 <- read_rds(file = paste0(path_argo_preprocessed, "/region_masks_all_1x1.rds"))
```

```{r read_map}
# read in the map from updata
map <-
  read_rds(paste(path_emlr_utilities,
                 "map_landmask_WOA18.rds",
                 sep = ""))
```

# Harmonise the two datasets 

Calculate monthly-mean Argo temperature for each lat/lon grid and each month 

```{r create_monthly_argo_data}

argo_temp_monthly <- argo_surf_temp %>%
  mutate(year_month = format_ISO8601(date, precision = "ym"), .after = 'date') %>%
  group_by(year, month, year_month, date, lat, lon) %>%
  summarise(argo_temp_month = mean(temp_adjusted, na.rm = TRUE)) %>%
  ungroup() %>%
  select(
    date,
    year_month,
    year,
    month,
    lon,
    lat,
    argo_temp_month
  )
```

Join Argo and OceanSODA 

```{r join_oceanSODA_argo}
OceanSODA_temp <- OceanSODA_temp %>% 
  mutate(year_month = format_ISO8601(date, precision = "ym")) %>% 
  rename(date_OceanSODA = date)# change date format in OceanSODA to match argo date (yyyy-mm)

argo_OceanSODA_temp <- left_join(argo_temp_monthly, OceanSODA_temp) %>%
  rename(OceanSODA_temp = temperature)
```

# Southern Ocean SST 

Focus on the Southern Ocean, south of 30ºS, as defined in the RECCAP biome regions 

```{r extract_southern_ocean_data}

region_masks_all_1x1_SO <- region_masks_all_1x1 %>%
  filter(region == 'southern',
         value != 0)

# keep only Southern Ocean data 
argo_OceanSODA_temp_SO <- 
  inner_join(region_masks_all_1x1_SO, argo_OceanSODA_temp)
```

## Monthly climatological OceanSODA SST

Map monthly mean SST from the OceanSODA data product

Climatological OceanSODA SST

```{r climatological_map, fig.asp=2}

# calculate average monthly pH between April 2013 and August 2021, and the difference between the two (offset)
argo_OceanSODA_temp_SO_clim <- argo_OceanSODA_temp_SO %>%
  group_by(lon, lat, month) %>%
  summarise(
    clim_OceanSODA_temp = mean(OceanSODA_temp, na.rm = TRUE),
    clim_argo_temp = mean(argo_temp_month, na.rm = TRUE),
    offset_clim = clim_OceanSODA_temp - clim_argo_temp
  ) %>%
  ungroup()

# regrid to a 2x2 grid for mapping 
argo_OceanSODA_temp_SO_clim_2x2 <- argo_OceanSODA_temp_SO_clim %>%
  mutate(
    lat = cut(lat, seq(-90, 90, 2), seq(-89, 89, 2)),
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 2), seq(21, 379, 2)),
    lon = as.numeric(as.character(lon))
  ) %>%
  group_by(lon, lat, month) %>%
  summarise(
    clim_OceanSODA_temp = mean(clim_OceanSODA_temp, na.rm = TRUE),
    clim_argo_temp = mean(clim_argo_temp, na.rm = TRUE),
    offset_clim = mean(offset_clim, na.rm = TRUE)
  ) %>%
  ungroup()

map +
  geom_tile(data = argo_OceanSODA_temp_SO_clim_2x2,
            aes(lon, lat, fill = clim_OceanSODA_temp)) +
  lims(y = c(-85, -25)) +
  scale_fill_viridis_c() +
  labs(x = 'lon',
       y = 'lat',
       fill = 'SST (ºC)',
       title = 'Monthly climatological \nOceanSODA SST (Jan 2013 - Dec 2020)') +
  theme(legend.position = 'right') +
  facet_wrap(~month, ncol = 2)
```

```{r polar_projection_SST_map, fig.asp=2}
# plot the climatological monthly OceanSODA SST on a polar projection 
basemap(limits = -32, data = argo_OceanSODA_temp_SO_clim_2x2) +   # change to polar projection
  geom_spatial_tile(data = argo_OceanSODA_temp_SO_clim_2x2,
                    aes(x = lon,
                        y = lat,
                        fill = clim_OceanSODA_temp),
                    linejoin = 'mitre',
                    col = 'transparent',
                    detail = 60)+
  scale_fill_viridis_c()+
  theme(legend.position = 'right')+
  labs(x = 'lon',
       y = 'lat',
       fill = 'SST (ºC)',
       title = 'monthly climatological \nOceanSODA SST (Jan 2013 - Dec 2020)')+
  facet_wrap(~month, ncol = 2)
```

## Monthly climatological Argo SST

Climatological Argo SST

```{r clim_map_argo_SST, fig.asp=2}
map +
  geom_tile(data = argo_OceanSODA_temp_SO_clim_2x2,
            aes(lon, lat, fill = clim_argo_temp)) +
  lims(y = c(-85, -25)) +
  scale_fill_viridis_c() +
  labs(x = 'lon',
       y = 'lat',
       fill = 'SST (ºC)',
       title = 'Monthly climatological \nArgo SST (Jan 2013 - Dec 2021)') +
  theme(legend.position = 'right') +
  facet_wrap(~month, ncol = 2)
```

```{r polar_projection_Argo_SST, fig.asp=2}

basemap(limits = -32, data = argo_OceanSODA_temp_SO_clim_2x2) +   # change to polar projection
  geom_spatial_tile(data = argo_OceanSODA_temp_SO_clim_2x2,
                    aes(x = lon,
                        y = lat,
                        fill = clim_argo_temp),
                    linejoin = 'mitre',
                    col = 'transparent',
                    detail = 60)+
  scale_fill_viridis_c()+
  theme(legend.position = 'right')+
  labs(x = 'lon',
       y = 'lat',
       fill = 'SST (ºC)',
       title = 'monthly climatological \nArgo SST (Jan 2013 - Dec 2021)')+
  facet_wrap(~month, ncol = 2)
```

## Timeseries of monthly OceanSODA SST

Evolution of monthly SST, for the three Southern Ocean RECCAP regions

```{r southern_ocean_biomes}

map +
  geom_raster(data = region_masks_all_seamask_2x2 %>%
                filter(seamask == 0),
              aes(x = lon, y = lat)) +
  geom_raster(data = region_masks_all_2x2 %>%
                filter(region == 'southern',
                       value != 0),
              aes(x = lon,
                  y = lat,
                  fill = value)) +
  labs(title = 'Southern Ocean RECCAP regions',
       fill = 'region')
```

```{r calculate_regional_SST_climatology}
# plot timeseries of monthly OceanSODA pH
argo_OceanSODA_temp_SO_clim_regional <- argo_OceanSODA_temp_SO %>%
  select(year, month, value, OceanSODA_temp, argo_temp_month) %>% 
  pivot_longer(c(OceanSODA_temp,argo_temp_month),
               values_to = "temp",
               names_to = "data_source") %>% 
  group_by(year, month, value, data_source) %>%  # compute regional mean OceanSODA pH for the three biomes
  summarise(temp = mean(temp, na.rm = TRUE)) %>%
  ungroup()
```

```{r plot_monthly_timeseries, fig.asp=2, eval=FALSE}

argo_OceanSODA_temp_SO_clim_regional %>%   
  ggplot(aes(x = year,
             y = temp,
             col = value)) +
  facet_grid(month ~ data_source) +
  geom_line() +
  geom_point() +
  labs(x = 'year',
       y = 'SST (ºC)',
       title = 'monthly mean SST (Jan 2013-Dec 2021, Southern Ocean)',
       col = 'region')
```

```{r yearly_timeseries_monthly_SST, fig.asp=2}

argo_OceanSODA_temp_SO_clim_regional %>%   
  filter(value != 0) %>%
  ggplot(aes(x = month,
             y = temp,
             group = year,
             col = as.character(year)))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 12, 2))+
  facet_grid(value~data_source)+
  lims(y = c(-5, 20))+
  labs(x = 'month',
       y = 'SST (ºC)',
       title = 'monthly mean SST (Jan 2013-Dec 2021, Southern Ocean)',
       col = 'year')
```

## Comparison between Argo and OceanSODA SST

Calculate the difference between Argo and OceanSODA SST values

Offset between in-situ monthly pH:

```{r calculate_in_situ_offset}

argo_OceanSODA_temp_SO <- argo_OceanSODA_temp_SO %>%
  mutate(offset = OceanSODA_temp - argo_temp_month)

argo_OceanSODA_temp_SO %>%
  # drop_na() %>%
  ggplot() +
  geom_hline(yintercept = 0, size = 1)+
  geom_point(aes(x = year_month, y = offset, col = value), size = 0.7, pch = 19) +
  scale_x_discrete(breaks = c('2013-01', '2014-01', '2015-01', '2016-01', '2017-01', '2018-01', '2019-01', '2020-01', '2021-01'))+
  labs(title = 'oceanSODA SST - Argo SST',
       x = 'date',
       y = 'offset (ºC)',
       col = 'region')
```

```{r oceanSODA_vs_argo_temp_plot}

argo_OceanSODA_temp_SO %>% 
  # drop_na() %>% 
  ggplot(aes(x = OceanSODA_temp, y = argo_temp_month))+
  # geom_point(pch = 19, size = 0.7)+
  geom_bin2d(aes(x = OceanSODA_temp, y = argo_temp_month), size = 0.3, bins = 60)+
  coord_fixed(ratio = 1,
              xlim = c(-3, 27),
              ylim= c(-3, 27)) +
  geom_abline(slope = 1, intercept = 0)+
  facet_wrap(~value)+
  labs(x = 'OceanSODA SST (ºC)',
       y = 'Argo SST (ºC)',
       title = 'Southern Ocean regional SST')

# test with basin and biome 
basinmask <-
  read_csv(paste(path_emlr_utilities,
                 "basin_mask_WOA18.csv",
                 sep = ""),
           col_types = cols("MLR_basins" = col_character()))

basinmask <- basinmask %>% 
  filter(MLR_basins == unique(basinmask$MLR_basins)[1]) %>% 
  select(lon, lat, basin_AIP)

argo_OceanSODA_temp_SO <- inner_join(argo_OceanSODA_temp_SO, basinmask)

argo_OceanSODA_temp_SO %>% 
  ggplot(aes(x = OceanSODA_temp, y = argo_temp_month))+
  geom_bin2d(aes(x = OceanSODA_temp, y = argo_temp_month), size = 0.3, bins = 60)+
  coord_fixed(ratio = 1, 
              xlim = c(-3, 27),
              ylim = c(-3, 27))+
  geom_abline(slope = 1, intercept = 0)+
  facet_grid(basin_AIP~value)+
  labs(x = 'Argo SST (ºC)', 
       y = 'OceanSODA SST (ºC)',
       title = 'Southern Ocean subregional SST')
```

Mean offset between in-situ OceanSODA SST and in-situ Argo SST

```{r mean_in_situ_SST_offset}

mean_insitu_offset <- argo_OceanSODA_temp_SO %>%
  group_by(year_month, value) %>% 
  summarise(mean_offset = mean(offset, na.rm = TRUE),
            std_offset = sd(offset, na.rm = TRUE))

mean_insitu_offset %>%
  # drop_na() %>%
  ggplot() +
  geom_hline(yintercept = 0, size = 1, col = 'red')+
  geom_point(aes(x = year_month, y = mean_offset, group = value, col = value), size = 0.7, pch = 19) +
  geom_line(aes(x = year_month, y = mean_offset, group = value, col = value))+
  geom_ribbon(aes(x = year_month, 
                  ymin = mean_offset-std_offset, 
                  ymax = mean_offset+std_offset, 
                  group = value, 
                  fill =value),
              alpha = 0.2)+
  scale_x_discrete(breaks = c('2013-01', '2014-01', '2015-01', '2016-01', '2017-01', '2018-01', '2019-01', '2020-01', '2021-01'))+
  # facet_wrap(~year)+
  labs(title = 'Mean offset (in situ oceanSODA SST - in situ Argo SST)',
       x = 'date',
       y = 'offset (ºC)',
       col = 'region',
       fill = '± 1 std')

```

Offset between climatological Argo and climatological OceanSODA SST:

```{r calculate_clim_offset}
# Offset between climatological argo and climatological OceanSODA SST 

argo_OceanSODA_temp_SO_clim <- inner_join(argo_OceanSODA_temp_SO_clim, region_masks_all_1x1_SO)
```

```{r plot_climatological_SST_offset, eval=FALSE}
argo_OceanSODA_temp_SO_clim %>% 
  # drop_na() %>%
  ggplot() +
  geom_point(aes(x = month, y = offset_clim, col = value), size = 0.7, pch = 19) +
  geom_hline(yintercept = 0, size = 1, col = 'red')+
  scale_x_continuous(breaks = seq(1, 12, 1))+
  labs(title = 'clim oceanSODA SST - clim Argo SST',
       x = 'month',
       y = 'offset (ºC)',
       col = 'region')
```

Mean offset between climatological OceanSODA SST and climatological Argo SST

```{r climatological_SST_offset_all_years}

mean_clim_offset <- argo_OceanSODA_temp_SO_clim %>% 
  group_by(month, value) %>% 
  summarise(mean_offset_clim = mean(offset_clim, na.rm = TRUE),
            std_offset_clim = sd(offset_clim, na.rm = TRUE))

mean_clim_offset %>% 
  ggplot()+
  geom_point(aes(x = month, y = mean_offset_clim, col = value))+
  geom_line(aes(x = month, y = mean_offset_clim, col = value))+
  geom_hline(yintercept = 0, col = 'red') +
  geom_ribbon(aes(x = month, 
                  ymin = mean_offset_clim - std_offset_clim, 
                  ymax = mean_offset_clim + std_offset_clim,
                  group = value, 
                  fill = value), 
              alpha = 0.2) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  labs(x = 'month',
       y = 'mean offset (ºC)',
       title = 'Mean offset (clim OceanSODA SST - clim Argo SST)', 
       col = 'region',
       fill = '± 1 std') 
```

Mapped offset between climatological OceanSODA pH and climatological Argo pH

```{r map_climatological_SST_offset, fig.asp=2}

# bin the offsets for better plotting  
# plot the offsets on a map of the Southern Ocean

argo_OceanSODA_temp_SO_clim_2x2 <- argo_OceanSODA_temp_SO_clim_2x2 %>% 
  mutate(offset_clim_binned = 
           cut(offset_clim, 
               breaks = c(-Inf, -0.025, -0.005, 0.000, 0.005, 0.025, 0.035, 0.05, Inf)))    # bin the offsets into intervals (create a discrete variable instead of continuous)
         # offset_clim_binned = as.factor(as.character(offset_clim_binned))) %>% 
  # drop_na()

map +
  geom_tile(data = argo_OceanSODA_temp_SO_clim_2x2,
            aes(lon, lat, fill = offset_clim_binned)) +
  lims(y = c(-85, -30)) +
  scale_fill_brewer(palette = 'RdBu', drop = FALSE) +
  labs(x = 'lon',
       y = 'lat',
       fill = 'offset (ºC)',
       title = 'clim OceanSODA SST - clim Argo SST') +
  theme(legend.position = 'right')+
  facet_wrap(~month, ncol = 2)
```

```{r map_climatological_SST_offsets_polar_map, fig.asp=2}

basemap(limits = -32, data = argo_OceanSODA_temp_SO_clim_2x2) +   # change to polar projection
  geom_spatial_tile(data = argo_OceanSODA_temp_SO_clim_2x2,
                    aes(x = lon,
                        y = lat,
                        fill = offset_clim_binned),
                    linejoin = 'mitre',
                    col = 'transparent',
                    detail = 60)+
  scale_fill_brewer(palette = 'RdBu', drop = FALSE)+
  theme(legend.position = 'right')+
  labs(x = 'lon',
       y = 'lat',
       fill = 'offset (ºC)',
       title = 'clim Ocean SODA SST - clim Argo SST')+
  facet_wrap(~month, ncol = 2)
```

### Basin separation

Using full OceanSODA data (even where there is no Argo data) Each RECCAP biome (1, 2, 3) is separated into basins (Atlantic, Pacific, Indian)

```{r add_in_basins_to_OceanSODA_temp}
basinmask <-
  read_csv(paste(path_emlr_utilities,
                 "basin_mask_WOA18.csv",
                 sep = ""),
           col_types = cols("MLR_basins" = col_character()))

basinmask <- basinmask %>% 
  filter(MLR_basins == unique(basinmask$MLR_basins)[1]) %>% 
  select(lon, lat, basin_AIP)

OceanSODA_temp_SO <- inner_join(OceanSODA_temp, region_masks_all_1x1) %>% 
  filter(region == 'southern')

OceanSODA_temp_SO <- inner_join(OceanSODA_temp_SO, basinmask) %>% 
  mutate(year = year(date_OceanSODA),
         month = month(date_OceanSODA)) %>% 
  mutate(date = format_ISO8601(date_OceanSODA, precision = 'ym'))
```

```{r monthly_OceanSODA_SST_basin}
# plot timeseries of monthly OceanSODA SST
OceanSODA_temp_SO_clim_subregional <- OceanSODA_temp_SO %>%
  filter(value != 0) %>% 
  group_by(year, month, value, basin_AIP) %>%  # compute regional mean OceanSODA pH for the three biomes
  summarise(temp = mean(temperature, na.rm = TRUE)) %>%
  ungroup()

# plot a timeseries of monthly average OceanSODA pH, per region and per basin
OceanSODA_temp_SO_clim_subregional %>% 
  ggplot(aes(x = month,
             y = temp,
             group = year,
             col = as.character(year)))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 12, 2))+
  facet_grid(value~basin_AIP)+
  labs(x = 'month',
       y = 'SST (ºC)',
       title = 'monthly mean OceanSODA SST (Jan 2013-Dec 2020, Southern Ocean basins)',
       col = 'year')
```

```{r monthly_OceanSODA_SST_basin_per_year}
OceanSODA_temp_SO_clim_subregional %>%  
  filter(value != 0) %>% 
  ggplot(aes(x = year,
             y = temp,
             col = value)) +
  facet_grid(month ~ basin_AIP) +
  geom_line() +
  geom_point() +
  labs(x = 'year',
       y = 'SST (ºC)',
       title = 'monthly mean OceanSODA SST (Jan 2013-Dec 2020, Southern Ocean basins)',
       col = 'region')
```

### Longitudinal separation 

Bin the SST data into 20º longitude bins
```{r regrid_temp_data_to_lon_bins}
OceanSODA_temp_SO_lon_binned <- OceanSODA_temp_SO %>%
  mutate(lon = cut(lon, seq(20, 380, 20), seq(30, 370, 20)),
         lon = as.numeric(as.character(lon))
  ) %>%
  group_by(lon, year, month, value) %>%
  summarise(
    OceanSODA_temp_binned = mean(temperature, na.rm = TRUE)
  ) %>%
  ungroup()
```

```{r plot_lon_binned_SST_data}

OceanSODA_temp_SO_lon_binned %>%
  # drop_na() %>% 
  filter(value != 0) %>% 
  ggplot(aes(x = month, y = OceanSODA_temp_binned, group = lon, col = as.factor(lon))) +
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 12, 2))+
  facet_grid(year~value)+
  labs(x = 'month',
       y = 'OceanSODA SST (ºC)',
       col = 'longitude bin')
```

