---
title: "OceanSODA-Argo pH"
author: "Pasqualina Vonlanthen & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Task 
Compare BGC-Argo pH data to pH from the OceanSODA surface data product 

```{r load_libraries}
# load in the necessary libraries
library(tidyverse)
library(argodata)
library(ggplot2)
library(lubridate)
library(ggOceanMaps)
```

# Load data 
Load in surface Argo pH and the OceanSODA pH, gridded to 1x1º 

```{r set_root_directories}

path_argo <- '/nfs/kryo/work/updata/bgc_argo_r_argodata'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")
path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"
```

```{r load_argo_oceanSODA_data}
# load in OceanSODA data and Argo pH
OceanSODA <- read_rds(file = paste0(path_argo_preprocessed, "/OceanSODA.rds"))

ph_surface_1x1 <- read_rds(file = paste0(path_argo_preprocessed, "/ph_surface_1x1.rds"))

# for plotting later, load in region and coastline information 
region_masks_all_seamask_2x2 <- read_rds(file = paste0(
  path_argo_preprocessed, "/region_masks_all_seamask_2x2.rds"))

region_masks_all_2x2 <- read_rds(file = paste0(path_argo_preprocessed, "/region_masks_all_2x2.rds"))
```

## Harmonise the two datasets 

Calculate monthly average pH for Argo pH for each lon/lat grid, centered on the 15th of each month, to match the format of OceanSODA
```{r create_monthly_argo_data}
argo_monthly_surface_ph_1x1 <- ph_surface_1x1 %>%
  mutate(day = rep(15, length(date)),  # change the date format to match OceanSODA
         .after = month) %>%
  unite(year, month, day,
        col = date,
        sep = '-',
        remove = FALSE) %>%
  mutate(date = ymd(date), .before = year) %>%
  group_by(date, lat, lon) %>%
  mutate(argo_ph_month = mean(ph_in_situ_total_adjusted, na.rm = TRUE), # calculate monthly mean argo parameters 
         argo_temp_month = mean(temp_adjusted, na.rm = TRUE),
         argo_psal_month = mean(psal_adjusted, na.rm = TRUE)) %>%
  ungroup() %>%
  select(date, year, month, day,
         lon, lat,
         float_serial_no, cycle_number,
         argo_temp_month,
         argo_psal_month,
         argo_ph_month,
         coast, region, value)
```

Join the two datasets 

```{r join_oceanSODA_argo}
argo_OceanSODA_1x1 <- left_join(argo_monthly_surface_ph_1x1, OceanSODA,
                            by = c('year', 'date', 'lat', 'lon')) %>%
  rename(OceanSODA_ph = ph_total,
         OceanSODA_ph_error = ph_total_uncert)

argo_OceanSODA_1x1 %>% 
  write_rds(file = paste0(path_argo_preprocessed, "/argo_OceanSODA_1x1.rds"))
```

# Southern Ocean surface pH 
The focus here is on Southern Ocean surface pH, south of 30ºS, as defined in the RECCAP biome regions

```{r extract_southern_ocean_data}

# keep only Southern Ocean data 
argo_OceanSODA_SO_1x1 <- argo_OceanSODA_1x1 %>%
  filter(region == 'southern',
         value != 0)
```

### Monthly climatological OceanSODA pH
Map monthly mean pH from the OceanSODA data product 
```{r set_global_theme}
theme_set(theme_bw())
```

```{r regrid_data}

# regrid the data into 2x2º, which is better for mapping 
argo_OceanSODA_SO_2x2 <- argo_OceanSODA_SO_1x1 %>%
  mutate(
    lat = cut(lat, seq(-90, 90, 2), seq(-89, 89, 2)),
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 2), seq(21, 379, 2)),
    lon = as.numeric(as.character(lon)))

```

```{r climatological_map}

# read in the map from updata
map <-
  read_rds(paste(path_emlr_utilities,
                 "map_landmask_WOA18.rds",
                 sep = ""))

# calculate average monthly pH between April 2014 and August 2021 
argo_OceanSODA_clim <- argo_OceanSODA_SO_2x2 %>% 
  group_by(lon, lat, month) %>% 
  summarise(clim_OceanSODA_ph = mean(OceanSODA_ph, na.rm = TRUE),
            clim_argo_ph = mean(argo_ph_month, na.rm = TRUE)) 

map +
  geom_tile(data = argo_OceanSODA_clim,
            aes(lon, lat, fill = clim_OceanSODA_ph)) +
  lims(y = c(-85, -25)) +
  scale_fill_viridis_c() +
  labs(x = 'lon',
       y = 'lat',
       fill = 'pH',
       title = 'Monthly climatological \nOceanSODA pH (Apr 2014 - Aug 2021)') +
  theme(legend.position = 'right') +
  facet_wrap(~month, ncol = 2)
```

```{r polar_projection_map}

# plot the climatological monthly OceanSODA pH on a polar projection 
basemap(limits = -32, data = argo_OceanSODA_clim) +   # change to polar projection
  geom_spatial_tile(data = argo_OceanSODA_clim,
                    aes(x = lon,
                        y = lat,
                        fill = clim_OceanSODA_ph),
                    linejoin = 'mitre',
                    col = 'transparent',
                    detail = 60)+
  scale_fill_viridis_c()+
  theme(legend.position = 'bottom')+
  labs(x = 'lon',
       y = 'lat',
       fill = 'pH',
       title = 'monthly climatological \nOceanSODA pH (Apr 2014 - Aug 2021)')+
  facet_wrap(~month)
```

### Timeseries of monthly OceanSODA pH 
Evolution of monthly surface pH, for the three Southern Ocean RECCAP regions 

```{r southern_ocean_biomes}

ggplot() +
  geom_raster(data = region_masks_all_seamask_2x2 %>%
                filter(seamask == 0),
              aes(x = lon, y = lat)) +
  geom_raster(data = region_masks_all_2x2 %>%
                filter(region == 'southern',
                       value != 0),
              aes(x = lon,
                  y = lat,
                  fill = value)) +
  coord_quickmap(expand = 0) +
  labs(title = 'Southern Ocean RECCAP regions',
       fill = 'region')
```

```{r plot_monthly_timeseries}

# plot timeseries of monthly OceanSODA pH
argo_OceanSODA_SO_2x2 %>%
  group_by(year, month, value) %>%  # compute regional mean OceanSODA pH for the three biomes
  mutate(OceanSODA_ph_region = mean(OceanSODA_ph, na.rm = TRUE)) %>%
  ggplot(aes(x = year,
             y = OceanSODA_ph_region,
             col = value)) +
  facet_wrap(~month) +
  geom_line() +
  geom_point() +
  labs(x = 'year',
       y = 'pH in situ (total scale)',
       title = 'monthly mean OceanSODA pH (Apr 2014-Aug 2021, Southern Ocean)',
       col = 'region')
```

```{r yearly_timeseries_monthly_oceanSODA_pH}

argo_OceanSODA_SO_2x2 %>%
  group_by(year, month, value) %>%   # compute regional mean OceanSODA pH for the three biomes
  mutate(OceanSODA_ph_region = mean(OceanSODA_ph, na.rm = TRUE)) %>%
  filter(year != 2014,
         year != 2021,
         value != 0) %>%
  ggplot(aes(x = month,
             y = OceanSODA_ph_region,
             group = year,
             col = as.character(year)))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 12, 2))+
  facet_wrap(~value)+
  labs(x = 'month',
       y = 'pH in situ (total scale)',
       title = 'monthly mean OceanSODA pH (Jan 2015-Dec 2020, Southern Ocean)',
       col = 'year')

```

### Comparison between Argo and OceanSODA pH

Calculate the difference between Argo and OceanSODA pH values 

In-situ monthly pH: 

```{r calculate_in_situ_offset}

argo_OceanSODA_SO_2x2 <- argo_OceanSODA_SO_2x2 %>%
  mutate(offset = OceanSODA_ph - argo_ph_month)

argo_OceanSODA_SO_2x2 %>%
  drop_na() %>%
  ggplot() +
  geom_point(aes(x = offset, y = date, col = value), size = 0.7, pch = 19) +
  geom_vline(xintercept = 0, size = 1, col = 'red')+
  labs(title = 'oceanSODA pH - Argo pH',
       x = 'offset (pH units)',
       y = 'date',
       col = 'region')
```

Offset between climatological Argo and climatological OceanSODA pH:

```{r calculate_clim_offset}
# Offset between climatological argo and climatological OceanSODA pH 

argo_OceanSODA_clim <- argo_OceanSODA_clim %>%
  mutate(offset_clim = clim_OceanSODA_ph - clim_argo_ph)

argo_OceanSODA_clim %>%
  drop_na() %>%
  ggplot() +
  geom_point(aes(x = offset_clim, y = month), size = 0.7, pch = 19) +
  geom_vline(xintercept = 0, size = 1, col = 'red')+
  scale_y_continuous(breaks = seq(1, 12, 1))+
  labs(title = 'clim oceanSODA pH - clim Argo pH',
       x = 'offset (pH units)',
       y = 'month',
       col = 'region')
```

```{r map_climatological_offset}
# plot the offsets on a map of the Southern Ocean
map +
  geom_tile(data = argo_OceanSODA_clim,
            aes(lon, lat, fill = offset_clim)) +
  lims(y = c(-85, -30)) +
  scale_fill_viridis_c() +
  labs(x = 'lon',
       y = 'lat',
       fill = 'offset (pH units)',
       title = 'clim OceanSODA ph - clim Argo pH') +
  theme(legend.position = 'right')+
  facet_wrap(~month, ncol = 2)

```

```{r map_climatological_offsets_polar_map}

basemap(limits = -32, data = argo_OceanSODA_clim) +   # change to polar projection
  geom_spatial_tile(data = argo_OceanSODA_clim,
                    aes(x = lon,
                        y = lat,
                        fill = offset_clim),
                    linejoin = 'mitre',
                    col = 'transparent',
                    detail = 60)+
  scale_fill_viridis_c()+
  theme(legend.position = 'bottom')+
  labs(x = 'lon',
       y = 'lat',
       fill = 'offset (pH units)',
       title = 'clim Ocean SODA pH - clim Argo pH')+
  facet_wrap(~month)

```


