---
title: "BGC-Argo Oxygen Data"
author: "Pasqualina Vonlanthen & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Task 

Explore BGC-Argo oxygen data through timeseries and climatological maps 

```{r loading_libraries, include=FALSE}
# load in the necessary libraries
library(tidyverse)
library(argodata)
library(ggplot2)
library(lubridate)
library(oce)
```

```{r set_root_directory}

path_argo <- '/nfs/kryo/work/updata/bgc_argo_r_argodata'
path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"
```

# Load oxygen data 

Load in delayed-mode, adjusted oxygen data from the BGC-Argo synthetic profile files 

```{r load_oxygen_data}

# set cache directory
argo_set_cache_dir(cache_dir = path_argo)

# periodically update the cached files
argo_update_global(max_global_cache_age = Inf)
argo_update_data(max_data_cache_age = Inf)

# load synthetic Argo files containing delayed-mode data between 2013 and now
bgc_subset <- argo_global_synthetic_prof() %>%
  argo_filter_data_mode(data_mode = 'delayed') %>%  # load in delayed-mode data
  argo_filter_date(date_min = '2013-01-01',
                   date_max = Sys.time())

# read in the oxygen data (with corresponding CTD variables)
oxy_data <- argo_prof_levels(
  path = bgc_subset,
  vars =
    c(
      'PRES_ADJUSTED',
      'PRES_ADJUSTED_QC',
      'PRES_ADJUSTED_ERROR',
      'PSAL_ADJUSTED',
      'PSAL_ADJUSTED_QC',
      'PSAL_ADJUSTED_ERROR',
      'TEMP_ADJUSTED',
      'TEMP_ADJUSTED_QC',
      'TEMP_ADJUSTED_ERROR',
      'DOXY_ADJUSTED',
      'DOXY_ADJUSTED_QC',
      'DOXY_ADJUSTED_ERROR'
    ),
  quiet = TRUE
)

# read in corresponding metadata
oxy_metadata <- argo_prof_prof(path = bgc_subset)

# merge the data and metadata
oxy_merge <-
  full_join(oxy_data, oxy_metadata)
```

```{r grid_oxygen_data}
oxy_merge <- oxy_merge %>%
  rename(lon = longitude,
         lat = latitude) %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon)) %>%
  mutate(
    lat = cut(lat, seq(-90, 90, 2), seq(-89.5, 89.5, 2)),
    lat = as.numeric(as.character(lat)),
    lon = cut(lon, seq(20, 380, 2), seq(20.5, 379.5, 2)),  # change to 2x2º grid
    lon = as.numeric(as.character(lon))
  ) %>%
  select(
    -c(profile_nitrate_qc:profile_up_radiance555_qc) 
  )        # remove profile_qc columns for non existent variables  

```

### Southern Ocean surface oxygen

Focus on surface oxygen (top 10 m of the watercolumn) in the Southern Ocean, south of 30ºS 

```{r select_relevant_oxygen_data}
# select only best pH data (with QC flag 1) below 30ºS, for the top 10 m of the watercolumn
oxy_surface <- oxy_merge %>%
  mutate(depth = swDepth(pres_adjusted, latitude = lat), .before = pres_adjusted) %>%
  filter(doxy_adjusted_qc == '1',   # keep only 'good' data
         lat <= -30,                   # keep only data at or south of 30ºS
         depth <= 10) %>%              # keep only data above or at 10 m depth
  mutate(
    year = year(date),     # separate the year and month from the date column
    month = month(date), .after = n_prof
  )

# check the correct latitudes, QC flags, and depth levels have been filtered
# max(oxy_surface$lat)
# min(oxy_surface$lat)
# table(oxy_surface$doxy_adjusted_qc)
# max(oxy_surface$depth)
# max(oxy_surface$date)
# min(oxy_surface$date)
```

#### Monthly climatological map

Create a map of climatological monthly oxygen values, from January 2013 to August 2021, for the region south of 30ºS 

```{r set_global_theme, include = FALSE}
theme_set(theme_bw())
```


```{r climatological_oxygen_map}

# average oxygen values in the top 10 m for each month in each 2 x 2º longitude/latitude grid 
oxy_mean <- oxy_surface %>%
  group_by(lat, lon, month) %>%
  summarise(oxy_ave_month = mean(doxy_adjusted))

# read in the map from updata
map <-
  read_rds(paste(path_emlr_utilities,
                 "map_landmask_WOA18.rds",
                 sep = ""))

# map a monthly climatology of surface oxygen (Jan 2013 - September 2021)
map +
  geom_tile(data = oxy_mean,
            aes(lon, lat, fill = oxy_ave_month)) +
  lims(y = c(-85, -25)) +
  scale_fill_gradientn(colors = oceColorsJet(n = oxy_mean$oxy_ave_month)) +
  labs(x = 'lon',
       y = 'lat',
       fill = 'dissolved oxygen \n(µmol kg-1)',
       title = 'Monthly average surface dissolved oxygen values (Jan 2013-Sep 2021)') +
  theme(legend.position = 'bottom')+
  facet_wrap(~month)

```

#### Monthly timeseries

Plot a timeseries of monthly-mean dissolved oxygen for the region south of 30ºS for the upper 10 m of the watercolumn 

```{r plot_monthly_oxygen_timeseries}

# plot a timeseries of monthly values over the whole southern ocean south of 30ºS
oxy_month <- oxy_surface %>%
  group_by(year, month) %>%
  summarise(oxy_ave = mean(doxy_adjusted))

# timeseries of monthly pH values over 2013-2021 (separate panels for each month)
oxy_month %>%
  ggplot(aes(x = year, y = oxy_ave)) +
  facet_wrap(~month) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(2013, 2021, 2))+
  labs(x = 'year', 
       y = 'dissolved O2 (µmol kg-1)', 
       title = 'monthly mean dissolved oxygen (Jan 2013-Sep 2021, south of 30ºS)')

```

Monthly average dissolved oxygen, per year (January 2013 - December 2020; plotting only full years), over the whole region south of 30ºS 

```{r monthly_oxygen_per_year}

# timeseries of monthly oxygen values for each year (separate years on the same plot)
oxy_month %>%
  filter(year != 2021) %>%    # keep only years with full data 
  ggplot(aes(x = month, y = oxy_ave, group = year, col = as.character(year)))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 12, 1))+
  labs(x = 'month',
       y = 'dissolved O2 (µmol kg-1)',
       title = 'monthly mean dissolved oxygen (Jan 2013-Dec 2020, south of 30ºS)',
       col = 'year')

```

### Northeast Pacific surface oxygen 

Focus on surface oxygen (upper 10 m) in the north-east Pacific (10ºN - 70ºN, -190ºE, -140ºE)

```{r select_nepacific_oxygen_data}

# select only best oxygen data (with QC flag 1) between 10 and 70ºN, and 190 and 140ºW, for the top 10 m of the watercolumn
oxy_nepacific <- oxy_merge %>%
  mutate(depth = swDepth(pres_adjusted, latitude = lat), .before = pres_adjusted) %>%
  filter(doxy_adjusted_qc == '1',   # keep only 'good' data
        between(lat, 10, 70),
        between(lon, 190, 240),                   # keep only data for the NE Pacific
        depth <= 10) %>%              # keep only data above or at 10 m depth
  mutate(
    year = year(date),     # separate the year and month from the date column
    month = month(date), .after = n_prof
  )
# longitudes larger than -180ºE are lon-380 

```

#### Monthly climatological map

Create a map of climatological monthly surface oxygen values, in the north-east Pacific ocean (10ºN - 70ºN, -190ºE, -140ºE), for January 2013 - August 2021

```{r climatological_oxygen_map_nepacific}

# average oxygen values in the top 10 m for each month in each 2 x 2º longitude/latitude grid 
oxy_mean_nepacific <- oxy_nepacific %>%
  group_by(lat, lon, month) %>%
  summarise(oxy_ave_month = mean(doxy_adjusted))

# map a monthly climatology of surface oxygen (Jan 2013 - August 2021)
map +
  geom_tile(data = oxy_mean_nepacific,
            aes(lon, lat, fill = oxy_ave_month)) +
  lims(y = c(5, 60), 
       x = c(180, 250)) +
  scale_fill_gradientn(colors = oceColorsJet(n = oxy_mean_nepacific$oxy_ave_month)) +
  labs(x = 'lon',
       y = 'lat',
       fill = 'dissolved oxygen \n(µmol kg-1)',
       title = 'Monthly average surface dissolved oxygen (Jan 2013-Aug 2021)') +
  theme(legend.position = 'right')+
  facet_wrap(~month)

```

#### Monthly timeseries

Timeseries of monthly mean oxygen for the northeast Pacific from January 2013 to August 2021

```{r timeseries_nepacific_oxygen}

# plot a timeseries of monthly values over the whole NE Pacific
oxy_month_nepacific <- oxy_nepacific %>%
  group_by(year, month) %>%
  summarise(oxy_ave = mean(doxy_adjusted))

# timeseries of monthly pH values over 2013-2021 (separate panels for each month)
oxy_month_nepacific %>%
  ggplot(aes(x = year, y = oxy_ave)) +
  facet_wrap(~month) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(2013, 2021, 2))+
  labs(x = 'year', 
       y = 'dissolved O2 (µmol kg-1)', 
       title = 'monthly mean dissolved oxygen (Jan 2013-Aug 2021, NE Pacific)')
```

Timeseries of monthly surface oxygen, per year, in the NE Pacific 

```{r timeseries_nepacific_oxygen_per_year}

# timeseries of monthly oxygen values for each year (separate years on the same plot)
oxy_month_nepacific %>%
  filter(year != 2021) %>%    # keep only years with full data
  ggplot(aes(x = month, y = oxy_ave, group = year, col = as.character(year)))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 12, 1))+
  labs(x = 'month',
       y = 'dissolved O2 (µmol kg-1)',
       title = 'monthly mean dissolved oxygen (Jan 2013-Dec 2020, NE Pacific)',
       col = 'year')

```

