---
title: "BGC-Argo pH Data"
author: "Pasqualina Vonlanthen & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Task

Explore BGC-Argo pH data through timeseries and monthly climatological maps

```{r loading_libraries, include=FALSE}
# load in the necessary libraries
library(tidyverse)
library(argodata)
library(ggplot2)
library(lubridate)
library(oce)
library(ggOceanMaps)
```

```{r set_root_directory}
path_argo <- '/nfs/kryo/work/updata/bgc_argo_r_argodata'
path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"
```

# Load pH data

Load in delayed-mode adjusted pH data from the data files created in Loading Data

```{r load_preprocessed_data}
# keep only pH data and associated CTD variables 
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")

# load in surface pH data 
ph_surface_2x2 <- read_rds(file = paste0(path_argo_preprocessed, '/ph_surface_2x2.rds'))

# ph_merge <-
#   read_rds(file = paste0(path_argo_preprocessed, "/bgc_merge.rds")) %>% 
#   select(
#     -c(doxy_adjusted:nitrate_adjusted_error),
#     -c(profile_doxy_qc, profile_nitrate_qc)
#   )
```

### Southern Ocean surface pH

The focus here is on surface pH (in the top 20 m of the watercolumn), in the region south of 30ºS

```{r select_relevant_ph_data}

# select only best pH data (with QC flag 1) below 30ºS, for the top 20 m of the watercolumn
ph_surface_SO <- ph_surface_2x2 %>%
  filter(lat <= -30)

# check the correct latitudes, QC flags, and depth levels have been filtered
# max(ph_surface$lat)
# min(ph_surface$lat)
# table(ph_surface$ph_in_situ_total_adjusted_qc)
# max(ph_surface$depth)
# min(ph_surface$date)
# max(ph_surface$date)
```

#### Monthly climatological map

Create a climatological monthly map of surface pH, in a 2x2º longitude/latitude grid, for the region south of 30ºS (monthly pH averaged over April 2014-August 2021)

```{r set_global_theme, include=FALSE}

theme_set(theme_bw())

```

```{r climatological_monthly_map}

# average pH values in the top 20 m for each month in each 2 x 2º longitude/latitude grid 
ph_mean_SO <- ph_surface_SO %>%
  group_by(lat, lon, month) %>%
  summarise(ph_ave_month = mean(ph_in_situ_total_adjusted))

# read in the map from updata
map <-
  read_rds(paste(path_emlr_utilities,
                 "map_landmask_WOA18.rds",
                 sep = ""))

# map a monthly climatology of pH (April 2014 - August 2021)
map +
  geom_tile(data = ph_mean_SO,
            aes(lon, lat, fill = ph_ave_month)) +
  lims(y = c(-85, -25)) +
  scale_fill_viridis_c() +
  labs(x = 'lon',
       y = 'lat',
       fill = 'pH',
       title = 'Monthly average pH (Apr 2014 - Aug 2021)') +
  theme(legend.position = 'bottom') +
  facet_wrap(~month)


```

```{r using_ggOceanMap}

basemap(limits = -30, data = ph_mean_SO) +   # change to polar projection 
  geom_spatial_tile(data = ph_mean_SO, 
            aes(x = lon,
                y = lat,
                fill = ph_ave_month),
            linejoin = 'mitre',
            col = 'transparent',
            detail = 60)+
  scale_fill_viridis_c()+
  theme(legend.position = 'bottom')+
  labs(x = 'lon',
       y = 'lat',
       fill = 'pH',
       title = 'monthly average pH (Apr 2014 - Aug 2021)')+
  facet_wrap(~month)
```

#### Monthly timeseries

Plot timeseries of monthly pH values, averaged over the whole region south of 30ºS

```{r plot_monthly_ph_timeseries}

# plot a timeseries of monthly values over the whole southern ocean south of 30ºS
ph_month_SO <- ph_surface_SO %>%
  group_by(year, month) %>%
  summarise(ph_ave = mean(ph_in_situ_total_adjusted))

# timeseries of monthly pH values over 2014-2021 (separate panels for each month)
ph_month_SO %>%
  ggplot(aes(x = year, y = ph_ave)) +
  facet_wrap(~month) +
  geom_line() +
  geom_point() +
  labs(x = 'year', 
       y = 'pH in situ (total scale)', 
       title = 'monthly mean pH (Apr 2014-Aug 2021, south of 30ºS)')
```

```{r weird_plot, eval=FALSE}

#all months on one plot in different colors (not very nice plot)
# ph_month_SO %>%
#   ggplot(aes(x = year, y = ph_ave, group = month, col = as.character(month))) +
#   geom_line() +
#   geom_point() +
#   labs(x = 'year', y = 'pH in situ (total scale)', title = 'monthly mean pH (Apr 2014-Aug 2021)')
```

Plot the monthly average pH, per year (from Jan 2015 - Dec 2020), over the whole region south of 30ºS

```{r monthly_pH_per_year}

# timeseries of monthly pH values for each year (separate years on the same plot)
ph_month_SO %>%
  filter(year != 2014,
         year != 2021) %>%    # remove the two years that are missing data (keep only data for full years)
  ggplot(aes(x = month, y = ph_ave, group = year, col = as.character(year)))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 12, 1))+
  labs(x = 'month',
       y = 'pH in situ (total scale)',
       title = 'monthly mean pH (Jan 2015-Dec 2020, south of 30ºS)',
       col = 'year')
```

```{r yearly_average_pH, eval=FALSE}

# calculate a yearly average ph (one ph value per year, for the whole domain)
ph_year_SO <- ph_surface_SO %>%
  group_by(year) %>%
  summarise(ph_ave = mean(ph_in_situ_total_adjusted))

# plot a timeseries of the yearly average pH value (one value per year)
ph_year_SO %>%
  ggplot(aes(x = year, y = ph_ave))+
  lims(y = c(8.03, 8.06))+
  geom_line()+
  geom_point()+
  labs(x = 'year',
       y = 'pH in situ (total scale)',
       title = 'yearly mean pH (Apr 2014-Aug 2021, south of 30ºS)')

```

### Northeast Pacific surface pH

Focus on surface pH in the northeast Pacific Ocean (10ºN - 70ºN, -190ºE - -140ºE)

```{r select_northeast_pacific_pH_data}

# select only best pH data (with QC flag 1) between 10 and 70ºN, and 190 and 140ºW, for the top 20 m of the watercolumn
ph_nepacific <- ph_surface_2x2 %>%
  mutate(depth = swDepth(pres_adjusted, latitude = lat), .before = pres_adjusted) %>%
  filter(ph_in_situ_total_adjusted_qc == '1',   # keep only 'good' data
        between(lat, 10, 70),
        between(lon, 190, 240),               # NE Pacific 
        depth <= 20) %>%              # keep only data above or at 20 m depth
  mutate(
    year = year(date),     # separate the year and month from the date column
    month = month(date), .after = n_prof
  )
# longitudes larger than -180ºE are lon-380

```

#### Monthly climatological map

Create a map of climatological monthly surface pH values, in the north-west Pacific ocean (10ºN - 70ºN, -190ºE, -140ºE), for

```{r northeast_pacific_climatological_pH_map}

# average pH values in the top 20 m for each month in each 2 x 2º longitude/latitude grid 
ph_mean_nepacific <- ph_nepacific %>%
  group_by(lat, lon, month) %>%
  summarise(ph_ave_month = mean(ph_in_situ_total_adjusted))

# map a monthly climatology of surface pH (Jan 2013 - August 2021)
map +
  geom_tile(data = ph_mean_nepacific,
            aes(lon, lat, fill = ph_ave_month)) +
  lims(y = c(5, 60), 
       x = c(180, 250)) +
  scale_fill_viridis_c() +
  labs(x = 'lon',
       y = 'lat',
       fill = 'pH',
       title = 'Monthly average pH (Jan 2013-Aug 2021)') +
  theme(legend.position = 'right')+
  facet_wrap(~month)

```

```{r ggOceanMaps_Pacific, eval=FALSE}
# using the ggOceanMaps package
basemap(limits = c(-180, -110, 7, 60), data = ph_mean_nepacific) +
  geom_spatial_tile(data = ph_mean_nepacific, 
              aes(x = lon,
                  y = lat, 
                  fill = ph_ave_month))+
  scale_fill_viridis_c()+
  facet_wrap(~month) + 
  labs(x = 'lon', 
       y = 'lat',
       fill = 'pH',
       title = 'Monthly average pH (Jan 2013-Aug 2021)')
# haven't figured out why the data isn't being plotted 
```

#### Monthly timeseries

Timeseries of monthly mean pH, averaged over the whole NE-Pacific region (10ºN - 70ºN, -190ºE - -140ºE), in the upper 20 m of the watercolumn.

```{r monthly_pH_timeseries_pacific}
# plot a timeseries of monthly values over the whole southern ocean south of 30ºS
ph_month_nepacific <- ph_nepacific %>%
  group_by(year, month) %>%
  summarise(ph_ave = mean(ph_in_situ_total_adjusted))

# timeseries of monthly pH values over 2014-2021 (separate panels for each month)
ph_month_nepacific %>%
  ggplot(aes(x = year, y = ph_ave)) +
  facet_wrap(~month) +
  scale_x_continuous(breaks = seq(2013, 2021, 2)) +
  geom_line() +
  geom_point() +
  labs(x = 'year', 
       y = 'pH in situ (total scale)', 
       title = 'monthly mean pH (Jan 2013-Aug 2021, NE Pacific)')
```

Monthly average pH, per year, over the NE Pacific region

```{r monthly_pH_timeseries_per_year_nepacific}

# timeseries of monthly pH values for each year (separate years on the same plot)
ph_month_nepacific %>%
  filter(year != 2016,
         year != 2021) %>%    # remove the two years that are missing data (keep only data for full years)
  ggplot(aes(x = month, y = ph_ave, group = year, col = as.character(year)))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 12, 1))+
  labs(x = 'month',
       y = 'pH in situ (total scale)',
       title = 'monthly mean pH (Jan 2013-Dec 2020, NE Pacific)',
       col = 'year')
```
