---
title: "BGC-Argo pH Data"
author: "Pasqualina Vonlanthen & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r set_options_global, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

# Task

Explore BGC-Argo pH data through timeseries and monthly climatological maps

```{r loading_libraries, include=FALSE}
# load in the necessary libraries
library(tidyverse)
library(ggplot2)
library(lubridate)
library(oce)
library(ggOceanMaps)
```

```{r set_root_directory}
path_argo <- '/nfs/kryo/work/updata/bgc_argo_r_argodata'
path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"
path_basin_mask <- "/nfs/kryo/work/updata/reccap2/"
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")
```

# Load pH data

Load in delayed-mode adjusted pH data from the data files created in Loading Data

```{r load_preprocessed_data}
# keep only pH data and associated CTD variables 

# load in surface pH data 
# ph_surface_2x2 <- read_rds(file = paste0(path_argo_preprocessed, '/ph_surface_2x2.rds'))
# 
# ph_surface_1x1 <- read_rds(file = paste0(path_argo_preprocessed, "/ph_surface_1x1.rds"))

ph_surface <- read_rds(file = paste0(path_argo_preprocessed, "/bgc_merge_flag_A.rds")) %>% 
  filter(between(depth, 0, 20)) %>% 
  mutate(year = year(date),
         month = month(date))

# region_masks_all_seamask_2x2 <- read_rds(file = paste0(
#   path_argo_preprocessed, "/region_masks_all_seamask_2x2.rds"))
# 
# region_masks_all_2x2 <- read_rds(file = paste0(path_argo_preprocessed, "/region_masks_all_2x2.rds"))

nm_biomes <- read_rds(file = paste0(path_argo_preprocessed, "/nm_biomes.rds"))
```

```{r set_global_theme, include=FALSE}

theme_set(theme_bw())

```

### Southern Ocean surface pH

The focus here is on surface pH (in the top 20 m of the watercolumn), in the region south of 30ºS

```{r select_relevant_ph_data}

ph_surface_SO <- ph_surface %>% 
  filter(lat <= - 30)

# check the correct latitudes, QC flags, and depth levels have been filtered
#max(ph_surface_SO$lat)
#min(ph_surface_SO$lat)
# table(ph_surface_SO$ph_in_situ_total_adjusted_qc)
# max(ph_surface_SO$depth)
# min(ph_surface_SO$date)
# max(ph_surface_SO$date)
```

#### pH offset with depth

Plot the difference between in-situ observed pH and the profile-mean surface pH for the upper 20 m. This difference represents the variability of the surface pH values with respect to the mean surface pH of the upper 20 m.

```{r SO_offset_from_the_mean}

# calculate the mean pH for each surface profile 
mean_profile_ph <- ph_surface_SO %>% 
  group_by(platform_number, cycle_number) %>% 
  mutate(mean_prof_ph = mean(ph_in_situ_total_adjusted, na.rm = TRUE), 
         .before = depth) %>% 
  ungroup() %>% 
  mutate(offset = ph_in_situ_total_adjusted-mean_prof_ph,   
         .after = mean_prof_ph) # subtract the mean profile pH from the measured in situ pH

mean_profile_ph %>%
  ggplot()+
  geom_point(aes(x = offset, y = depth, col = as.character(year)), size = 0.3, pch = 19) +
  scale_y_reverse()+
  geom_vline(xintercept = 0, col = 'red', size = 0.6)+
  labs(x = 'offset (pH units)',
       y = 'depth (m)',
       col = 'year',
       title = 'in situ pH - mean profile pH')
```

Bin the pH data into 2m-depth intervals and calculate the offset for each pH observation in each depth interval relative to the profile-mean pH

```{r offset_SO_binned_by_depth}

# bin the ph values into 2m bins and calculate the offset for each 2m bin 

mean_profile_ph_binned <- ph_surface_SO %>% 
  mutate(depth = cut(depth, seq(0, 20, 2), seq(1, 19, 2)),
         depth = as.numeric(as.character(depth))) %>% 
  group_by(platform_number, cycle_number) %>% 
  mutate(mean_prof_ph = mean(ph_in_situ_total_adjusted, na.rm = TRUE),
         .before = depth) %>% 
  ungroup() %>% 
  mutate(offset = ph_in_situ_total_adjusted-mean_prof_ph, 
         .after = mean_prof_ph) 

# plot the offset of the depth-binned values 
mean_profile_ph_binned %>%
  ggplot()+
  geom_point(aes(x = offset, y = depth, col = as.character(year)), size = 0.3, pch = 19) +
  scale_y_reverse()+
  geom_vline(xintercept = 0, col = 'red', size = 0.6)+
  labs(x = 'offset (pH units)',
       y = 'depth (m)',
       col = 'year',
       title = 'in situ pH - mean profile pH (2m depth bins)')

```

Mean offset for each 2m depth bin

```{r pH_offset_mean_binned}

# bin the ph values into 2m bins and calculate the offset for each 2m bin 
profile_ph_binned_ave <- ph_surface_SO %>% 
  mutate(depth = cut(depth, seq(0, 20, 2), seq(1, 19, 2)),
         depth = as.numeric(as.character(depth))) %>% 
  group_by(platform_number, cycle_number) %>% 
  mutate(mean_prof_ph = mean(ph_in_situ_total_adjusted, na.rm = TRUE),
         .before = depth) %>% 
  ungroup() %>% 
  mutate(offset = ph_in_situ_total_adjusted-mean_prof_ph, 
         .after = mean_prof_ph) %>% 
  group_by(depth) %>% 
  summarise(mean_offset = mean(offset))

# plot the offset of the depth-binned values 
profile_ph_binned_ave %>%
  ggplot()+
  geom_point(aes(x = mean_offset, y = depth), size = 1, pch = 19) +
  geom_line(aes(x = mean_offset, y = depth))+
  scale_y_reverse()+
  geom_vline(xintercept = 0, col = 'red', size = 1)+
  labs(x = 'mean offset (pH units)',
       y = 'depth (m)',
       col = 'year',
       title = 'in situ pH - mean profile pH (2m depth bins)')

```

#### Monthly climatological map

Create a climatological monthly map of surface pH, in a 2x2º longitude/latitude grid, for the region south of 30ºS (monthly pH averaged over April 2014-December 2021)

```{r climatological_monthly_map, fig.asp=2}

# average pH values in the top 20 m for each month in each 2 x 2º longitude/latitude grid 
ph_clim_SO <- ph_surface_SO %>%
  group_by(lat, lon, month) %>%
  summarise(ph_clim_month = mean(ph_in_situ_total_adjusted))

# read in the map from updata
map <-
  read_rds(paste(path_emlr_utilities,
                 "map_landmask_WOA18.rds",
                 sep = ""))

# map a monthly climatology of pH (April 2014 - December 2021)
map +
  geom_tile(data = ph_clim_SO,
            aes(lon, lat, fill = ph_clim_month)) +
  lims(y = c(-85, -25)) +
  scale_fill_viridis_c() +
  labs(x = 'lon',
       y = 'lat',
       fill = 'pH',
       title = 'Monthly climatological \nArgo pH (Apr 2014 - Dec 2021)') +
  theme(legend.position = 'right') +
  facet_wrap(~month, ncol = 2)


```

```{r using_ggOceanMaps, fig.asp=2, eval=FALSE}

basemap(limits = -32, data = ph_clim_SO) +   # change to polar projection 
  geom_spatial_tile(data = ph_clim_SO, 
            aes(x = lon,
                y = lat,
                fill = ph_clim_month),
            linejoin = 'mitre',
            col = 'transparent',
            detail = 60)+
  scale_fill_viridis_c()+
  theme(legend.position = 'bottom')+
  labs(x = 'lon',
       y = 'lat',
       fill = 'pH',
       title = 'monthly climatological \nArgo pH (Apr 2014 - Dec 2021)')+
  facet_wrap(~month, ncol = 2)
```

#### Monthly timeseries

Timeseries of monthly mean pH values, over the three different Southern Ocean regions (separated based on Mayot biomes):

```{r map_southern_ocean_biomes}

# plot the region separations on a map 

map +
  geom_raster(data = nm_biomes, 
              aes(x = lon, 
                  y = lat, 
                  fill = biome_name)) +
  labs(title = 'Southern Ocean Mayot biomes', 
       fill = 'biome')
```

```{r plot_monthly_ph_timeseries}

# plot a timeseries of monthly values over the whole southern ocean south of 30ºS

ph_surface_SO <- inner_join(ph_surface_SO, nm_biomes)

ph_month_SO <- ph_surface_SO %>%
  group_by(year, month, biome_name) %>%
  summarise(ph_ave = mean(ph_in_situ_total_adjusted))

# timeseries of monthly pH values over 2014-2021 (separate panels for each month)
ph_month_SO %>%
  ggplot(aes(x = year, 
             y = ph_ave, 
             group = biome_name, 
             col = biome_name)) +
  facet_wrap(~month) +
  geom_line() +
  geom_point() +
  labs(x = 'year', 
       y = 'pH in situ (total scale)', 
       title = 'monthly mean Argo pH (Apr 2014-Dec 2021, Southern Ocean)', 
       col = 'region')

```

```{r weird_plot, eval=FALSE}

#all months on one plot in different colors (not very nice plot)
# ph_month_SO %>%
#   ggplot(aes(x = year, y = ph_ave, group = month, col = as.character(month))) +
#   geom_line() +
#   geom_point() +
#   labs(x = 'year', y = 'pH in situ (total scale)', title = 'monthly mean pH (Apr 2014-Aug 2021)')
```

Plot the monthly average pH, per year (from Jan 2015 - Dec 2020), for each Southern Ocean RECCAP region (1, 2, 3)

```{r monthly_pH_per_year}

# timeseries of monthly pH values for each year (separate years on the same plot)
ph_month_SO %>%
  # filter(year != 2014) %>%    # remove the year that is missing data 
  ggplot(aes(x = month, 
             y = ph_ave, 
             group = year,
             col = as.character(year)))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 12, 2))+
  facet_wrap(~biome_name)+
  labs(x = 'month',
       y = 'pH in situ (total scale)',
       title = 'monthly mean Argo pH (Jan 2015-Dec 2021, Southern Ocean regions)',
       col = 'year')
```

```{r yearly_average_pH, eval=FALSE}

# calculate a yearly average ph (one ph value per year, for the whole domain)
ph_year_SO <- ph_surface_SO %>%
  group_by(year, biome_name) %>%
  summarise(ph_ave = mean(ph_in_situ_total_adjusted))

# plot a timeseries of the yearly average pH value (one value per year)
ph_year_SO %>%
  ggplot(aes(x = year, y = ph_ave, group = biome_name, col = biome_name))+
  geom_line()+
  geom_point()+
  labs(x = 'year',
       y = 'pH in situ (total scale)',
       title = 'yearly mean Argo pH (Apr 2014-Aug 2021, south of 30ºS)', 
       col = 'region')

```

### Northeast Pacific surface pH

Focus on surface pH in the northeast Pacific Ocean (10ºN - 70ºN, -190ºE - -140ºE)

```{r select_northeast_pacific_pH_data, eval=FALSE}

# select only pH databetween 10 and 70ºN, and 190 and 140ºW, for the top 20 m of the watercolumn

ph_nepacific <- ph_surface %>%
  filter(between(lat, 10, 70),
         between(lon, 190, 240)) 
# longitudes larger than -180ºE are lon-380

```

#### Monthly climatological map

Create a map of climatological monthly surface pH values, in the north-west Pacific ocean (10ºN - 70ºN, -190ºE, -140ºE), for

```{r northeast_pacific_climatological_pH_map, eval=FALSE}

# average pH values in the top 20 m for each month in each 2 x 2º longitude/latitude grid 
ph_mean_nepacific <- ph_nepacific %>%
  group_by(lat, lon, month) %>%
  summarise(ph_ave_month = mean(ph_in_situ_total_adjusted))

# map a monthly climatology of surface pH (Jan 2013 - August 2021)
map +
  geom_tile(data = ph_mean_nepacific,
            aes(lon, lat, fill = ph_ave_month)) +
  lims(y = c(5, 60), 
       x = c(180, 250)) +
  scale_fill_viridis_c() +
  labs(x = 'lon',
       y = 'lat',
       fill = 'pH',
       title = 'Monthly average pH (Jan 2013-Aug 2021)') +
  theme(legend.position = 'right')+
  facet_wrap(~month)

```

```{r ggOceanMaps_Pacific, eval=FALSE}

# using the ggOceanMaps package
basemap(limits = c(-180, -110, 7, 60), data = ph_mean_nepacific) +
  geom_spatial_tile(data = ph_mean_nepacific, 
              aes(x = lon,
                  y = lat, 
                  fill = ph_ave_month))+
  scale_fill_viridis_c()+
  facet_wrap(~month) + 
  labs(x = 'lon', 
       y = 'lat',
       fill = 'pH',
       title = 'Monthly average pH (Jan 2013-Aug 2021)')
# haven't figured out why the data isn't being plotted 
```

#### Monthly timeseries

Timeseries of monthly mean pH, averaged over the whole NE-Pacific region (10ºN - 70ºN, -190ºE - -140ºE), in the upper 20 m of the watercolumn.

```{r monthly_pH_timeseries_pacific, eval=FALSE}

# plot a timeseries of monthly values over the whole southern ocean south of 30ºS

ph_month_nepacific <- ph_nepacific %>%
  group_by(year, month) %>%
  summarise(ph_ave = mean(ph_in_situ_total_adjusted))

# timeseries of monthly pH values over 2014-2021 (separate panels for each month)
ph_month_nepacific %>%
  ggplot(aes(x = year, y = ph_ave)) +
  facet_wrap(~month) +
  scale_x_continuous(breaks = seq(2013, 2021, 2)) +
  geom_line() +
  geom_point() +
  labs(x = 'year', 
       y = 'pH in situ (total scale)', 
       title = 'monthly mean pH (Jan 2013-Aug 2021, NE Pacific)')
```

Monthly average pH, per year, over the NE Pacific region

```{r monthly_pH_timeseries_per_year_nepacific, eval=FALSE}

# timeseries of monthly pH values for each year (separate years on the same plot)
ph_month_nepacific %>%   
  ggplot(aes(x = month, y = ph_ave, group = year, col = as.character(year)))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 12, 1))+
  labs(x = 'month',
       y = 'pH in situ (total scale)',
       title = 'monthly mean pH (Jan 2013-Dec 2020, NE Pacific)',
       col = 'year')
```
