---
title: "SO BGC temperature cluster analysis"
author: "David Stappard & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r set_options_global, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

# Tasks

This markdown file carries out cluster analysis on previously created bgc temperature anomaly profiles.

The cluster_analysis_determine_k chunk is used to give an indication of an appropriate number of clusers and this can then set the option opt_num_clusters. Chunk cluster_analysis_cluster_details carried out the cluster analysis and the results are used in subsequent figures.

The cluster analysis is carried out on base anomaly profiles and well as adjusted profiles. Adjusted profiles have been normalised by dividing each anomaly in the profile by the max anomaly across the profile. This results in all Adjusted profiles having values in the rage -1 to +1.

A restriction is added that just SO profiles are considered. Also the profile must have a valid pH profile associated.

cluster_analysis_base.Rmd - Child markdown file that does cluster analysis and can if set overlay additional attributes on clusters.

cluster_analysis_extreme.Rmd - Child markdown file that does cluster analysis under surface extremes conditions.

Dependencies
------------

temp_anomaly_va.rds - bgc preprocessed folder, created by temp_align_climatology

OceanSODA_global_SST_anomaly_field_01.rds (or _02.rds) - bgc preprocessed folder, created by extreme_temp

```{r loading_libraries, include=FALSE}

library(tidyverse)
library(argodata)
library(lubridate)
library(gridExtra)
library(gsw)
library(ggforce)
library(ggpmisc)
library(tidymodels)

```

## Set directories

location of pre-prepared data

```{r set_updata_root_directory, include=FALSE}

path_argo <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/bgc_argo'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")

path_core_argo <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/core_argo_r_argodata'
path_core_preprocessed <- paste0(path_core_argo, "/preprocessed_core_data")

path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"

```

## Category options

What category of data is the cluster analysis related to

```{r set_category_options}

# opt_category
opt_category <- "bgc_temp_SO"

# # Global
# opt_lat_min <- -90
# opt_lat_max <- 90
# opt_lon_min <- 20
# opt_lon_max <- 380
# 
# # spatial restrictions
# # Southern ocean
# opt_lat_min <- -90
# opt_lat_max <- -30
# opt_lon_min <- 20
# opt_lon_max <- 380
# 
# # Mapping latitude limits
# opt_map_lat_limit <- c(-85, 85) # global
# opt_map_lat_limit <- c(-85,-30) # SO

```

## Analysis options

Define options that are used to determine the nature of the cluster analysis that is carried out.

```{r set_cluster_options}

# Options

# opt_num_clusters
# How many clusters are used in the cluster analysis for each depth 1 (600 m), 2 (1000 m) and 3 (1500 m)
opt_num_clusters_min <- c(8, 8, 4)
opt_num_clusters_max <- c(8, 8, 5)
# Which profile range is used
opt_profile_range <- 3

# options relating to cluster analysis
opt_n_start <- 15
opt_max_iterations <- 500
opt_n_clusters <- 14 # Max number of clusters to try when determining optimal number of clusters

# opt_extreme_determination
# 1 - based on the trend of de-seasonal data - we believe this results in more summer extremes where variation tend to be greater.
# 2 - based on the trend of de-seasonal data by month. grouping is by lat, lon and month.
opt_extreme_determination <- 2

# Options associated with profiles under surface extreme conditions
extreme_type <- c('L', 'N', 'H')
opt_num_clusters_ext_min <- c(4, 4, 4)
opt_num_clusters_ext_max <- c(5, 5, 5)

# Option related to normalising the anomaly profiles.
# TRUE - anomaly profiles are normalised by the surface anomaly. Every depth anomaly is divided by the surface anomaly.
#      - The is only carried out for profiles where the abs(surface temp) > 1.
#      - This analysis is carried out in addition to the analysis on base anomaly profiles.  
# FALSE - The normalisation process is not carried out. 
opt_norm_anomaly <- TRUE

```


```{r set_global_theme, include=TRUE}

theme_set(theme_bw())

map <-
  read_rds(paste(path_emlr_utilities,
                 "map_landmask_WOA18.rds",
                 sep = ""))

```


## Preparation

Prepare data for cluster analysis

```{r prep_bgc_temp}

if (opt_category == "bgc_temp_SO") {

  # ---------------------------------------------------------------------------------------------
  # spatial restrictions
  # ---------------------------------------------------------------------------------------------
  # Southern Ocean
  opt_lat_min <- -90
  opt_lat_max <- -30
  opt_lon_min <- 20
  opt_lon_max <- 380
  
  # Mapping latitude limits
  opt_map_lat_limit <- c(-85, -30) # SO
    
  # ---------------------------------------------------------------------------------------------
  # read data - bgc temperature must have a ph profile
  # ---------------------------------------------------------------------------------------------
  # read data, applying geographical limits and standardize field names.
  anomaly_va <-
    read_rds(file = paste0(path_argo_preprocessed, "/temp_anomaly_va.rds")) %>%
    filter (lat >= opt_lat_min &
            lat <= opt_lat_max &
            lon >= opt_lon_min &
            lon <= opt_lon_max) %>%
    select(file_id,
           date,
           year,
           month,
           lat,
           lon,
           profile_range,
           depth, 
           prof_measure = temp,
           clim_measure = clim_temp,
           anomaly
           )

  # Additional ph requirement restriction
  bgc_ph <- read_rds(file = paste0(path_argo_preprocessed, "/pH_bgc_va.rds"))
  anomaly_va <- inner_join(anomaly_va, bgc_ph %>% distinct(file_id)) 
  
  # ---------------------------------------------------------------------------------------------
  # read data extreme data for later use
  # ---------------------------------------------------------------------------------------------
  # load previously created OceanSODA extreme data. date, position and nature of extreme
  if (opt_extreme_determination == 1){
    extreme_data <- read_rds(file = paste0(path_argo_preprocessed, "/OceanSODA_global_SST_anomaly_field_01.rds")) %>%
      select(lon, lat, date, extreme_flag = temp_extreme)
  } else if (opt_extreme_determination == 2){
    extreme_data <- read_rds(file = paste0(path_argo_preprocessed, "/OceanSODA_global_SST_anomaly_field_02.rds")) %>%
      select(lon, lat, date, extreme_flag = temp_extreme)
  }

  # ---------------------------------------------------------------------------------------------
  # Associated data restrictions and formatting
  # ---------------------------------------------------------------------------------------------
  # What is the max depth of each profile_range
  opt_max_depth <- c(600, 1000, 1500)
  
  # opt_measure_label, opt_xlim and opt_xbreaks are associated with formatting
  opt_measure_label <- "temperature anomaly (°C)"
  opt_xlim <- c(-4.5, 4.5)
  opt_xbreaks <- c(-4, -2, 0, 2, 4)
  
  # adjusted to be in scale -1 to 1
  opt_measure_label_adjusted <- "adjusted temperature anomaly"
  opt_xlim_adjusted <- c(-1, 1)
  opt_xbreaks_adjusted <- c(-1.0, -0.5, 0, 0.5, 1.0)

  # oxygen formatting
  opt_doxy_measure_label <- expression("dissolved oxygen anomaly ( µmol kg"^"-1"~")")

  # pH formatting
  opt_ph_h_plus_measure_label <- expression("[H]"^"+" ~ "anomaly")

  # Under extreme analysis
  opt_extreme_analysis <- TRUE
  
}

```

```{r source_cluster_analysis_base_child, include = FALSE}


cluster_analysis_base <-
  knitr::knit_expand(
    file = here::here("analysis/child/cluster_analysis_base.Rmd")
  )


```

`r knitr::knit(text = unlist(cluster_analysis_base))`


```{r source_cluster_analysis_extreme_child, include = FALSE}


cluster_analysis_extreme <-
  knitr::knit_expand(
    file = here::here("analysis/child/cluster_analysis_extreme.Rmd")
  )


```

`r knitr::knit(text = unlist(cluster_analysis_extreme))`
