---
title: "Temperature core cluster analysis"
author: "David Stappard & Jens Daniel Müller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r set_options_global, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

# Tasks

This markdown file carries out cluster analysis on previously created core temperature anomaly profiles.

The cluster_analysis_determine_k chunk is used to give an indication of an appropriate number of clusers and this can then set the option opt_num_clusters. Chunk cluster_analysis_cluster_details carried out the cluster analysis and the results are used in subsequent figures.


```{r loading_libraries, include=FALSE}

library(tidyverse)
library(argodata)
library(lubridate)
library(gridExtra)
library(gsw)
library(ggforce)
library(ggpmisc)
library(tidymodels)

```

## Set directories

location of pre-prepared data

```{r set_updata_root_directory, include=FALSE}

path_argo <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/bgc_argo'
path_argo_preprocessed <- paste0(path_argo, "/preprocessed_bgc_data")

path_core_argo <- '/nfs/kryo/work/datasets/ungridded/3d/ocean/floats/core_argo_r_argodata'
path_core_preprocessed <- paste0(path_core_argo, "/preprocessed_core_data")

path_emlr_utilities <- "/nfs/kryo/work/jenmueller/emlr_cant/utilities/files/"

```

## Set options

Define options that are used to determine what analysis is done

```{r set_options}

# Options

# opt_analysis_type
# opt_analysis_type = 1 do analysis to determine number of clusters to use
# opt_analysis_type = 2 do cluster analysis and further analysis on the identified clusters clusters
opt_analysis_type <- 2

# opt_num_clusters
# How many clusters are used in the cluster analysis for each depth 1 (600 m), 2 (1000 m) and 3 (1500 m)
opt_num_clusters <- c(8, 8, 8)
# What is the max depth of each profile_range
opt_max_depth <- c(600, 1000, 1500)
# Which profile range is used
opt_profile_range <- 3

# opt_measure_label, opt_xlim and opt_xbreaks are associated formatting
opt_measure_label <- "temperature anomaly (°C)"
opt_xlim <- c(-4, 4)
opt_xbreaks <- c(-4, -2, 0, 2, 4)

# options relating to cluster analysis
opt_n_start <- 15
opt_max_iterations <- 500
opt_n_clusters <- 14 # Max number of clusters to try when determining optimal number of clusters

```


```{r set_global_theme, include=TRUE}

theme_set(theme_bw())

map <-
  read_rds(paste(path_emlr_utilities,
                 "map_landmask_WOA18.rds",
                 sep = ""))

```

## Cluster analysis

### Preperation

Prepare data for cluster analysis

```{r cluster_analysis_prep}

temp_anomaly_va <- read_rds(file = paste0(path_core_preprocessed, "/temp_anomaly_va.rds"))


# select just 1500m profiles
temp_anomaly_va <- temp_anomaly_va %>% 
  filter(profile_range == opt_profile_range & depth <= opt_max_depth[opt_profile_range])

# Simplified table ready to pivot
temp_anomaly_va_id <- temp_anomaly_va %>%
  select(file_id,
         depth,
         anomaly,
         year, 
         month, 
         lat, 
         lon)

# Check duplicate values and remove those profiles
check_duplicates <- temp_anomaly_va_id %>%
  group_by(file_id, depth) %>%
  summarise(count_measures = n()) %>%
  filter(count_measures > 1) %>%
  ungroup()

check_duplicates <- check_duplicates %>%
  select (file_id, count_measures)

temp_anomaly_va_id2 <-
  full_join(temp_anomaly_va_id, check_duplicates)

temp_anomaly_va_id <- temp_anomaly_va_id2 %>%
  filter(is.na(count_measures)) %>%
  select(file_id,
         depth,
         anomaly,
         year,
         month,
         lat,
         lon)

rm(temp_anomaly_va_id2)
                            
# wide table with each depth becoming a column
temp_anomaly_va_wide <- temp_anomaly_va_id %>%
  select(file_id, depth, anomaly) %>%
  pivot_wider(names_from = depth, values_from = anomaly)

# Drop any rows with missing values N/A caused by gaps in climatology data
temp_anomaly_va_wide <- temp_anomaly_va_wide %>% 
  drop_na()

# Table for cluster analysis
points <- temp_anomaly_va_wide %>%
  column_to_rownames(var = "file_id")

```

### Number of clusters

```{r cluster_analysis_determine_k}

if (opt_analysis_type == 1) {


  # cluster analysis - What k? try between 1 and opt_n_clusters clusters
  kclusts <- 
  tibble(k = 1:opt_n_clusters) %>%
  mutate(
    kclust = map(k, ~kmeans(points, .x, iter.max = opt_max_iterations, nstart = opt_n_start)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, points)
  )
  

  # cluster analysis data
  clusters <-
    kclusts %>%
    unnest(cols = c(tidied))
  
  assignments <-
    kclusts %>%
    unnest(cols = c(augmented))
  
  clusterings <-
    kclusts %>%
    unnest(cols = c(glanced))
  
  # What cluster works best?
  clusterings %>%
  ggplot(aes(k, tot.withinss)) +
    geom_line() +
    geom_point() +
    scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14))

}

```

### Cluster analysis

```{r cluster_analysis_cluster_details}

if (opt_analysis_type == 2) {

  set.seed(1)
  kclusts <-
    tibble(k = opt_num_clusters[opt_profile_range]) %>%
    mutate(
      kclust = map(k, ~ kmeans(points, .x, iter.max = opt_max_iterations, nstart = opt_n_start)),
      tidied = map(kclust, tidy),
      glanced = map(kclust, glance),
      augmented = map(kclust, augment, points)
    )
  
  profile_id <-
    kclusts %>%
    unnest(cols = c(augmented)) %>%
    select(file_id = .rownames,
           cluster = .cluster) %>% 
    mutate(file_id = as.numeric(file_id),
           cluster = as.character(cluster))
  
  # Add cluster to temp_anomaly_va
  temp_anomaly_cluster <- full_join(temp_anomaly_va, profile_id)
  
  # Plot cluster mean
  temp_anomaly_cluster <- temp_anomaly_cluster %>% 
    filter(!is.na(cluster))
  
  # Plot cluster mean
  anomaly_cluster_mean <- temp_anomaly_cluster %>%
    group_by(cluster, depth) %>%
    summarise(
      count_cluster = n(),
      anomaly_mean = mean(anomaly, na.rm = TRUE),
      anomaly_sd = sd(anomaly, na.rm = TRUE)
    ) %>%
    ungroup()
  
  anomaly_cluster_mean_year <- temp_anomaly_cluster %>%
    group_by(cluster, depth, year) %>%
    summarise(
      count_cluster = n(),
      anomaly_mean = mean(anomaly, na.rm = TRUE),
      anomaly_sd = sd(anomaly, na.rm = TRUE)
    ) %>%
    ungroup()
  
  anomaly_year_mean <- temp_anomaly_cluster %>%
    group_by(cluster, year) %>%
    summarise(
      count_cluster = n(),
      anomaly_mean = mean(anomaly, na.rm = TRUE),
      anomaly_sd = sd(anomaly, na.rm = TRUE)
    ) %>%
    ungroup()
  
  anomaly_year_mean <- anomaly_year_mean %>%
    group_by(year) %>%
    summarise(anomaly_mean = mean(anomaly_mean, na.rm = TRUE)) %>%
    ungroup ()
  
  # create figure
  anomaly_cluster_mean %>%
    ggplot() +
    geom_path(aes(x = anomaly_mean,
                  y = depth)) +
    geom_ribbon(aes(
      xmax = anomaly_mean + anomaly_sd,
      xmin = anomaly_mean - anomaly_sd,
      y = depth
    ),
    alpha = 0.2) +
    geom_vline(xintercept = 0) +
    scale_y_reverse() +
    facet_wrap( ~ cluster) +
    coord_cartesian(xlim = opt_xlim) +
    scale_x_continuous(breaks = opt_xbreaks) +
    labs(
      title = paste0('Overall mean anomaly profiles by cluster'),
      x = opt_measure_label,
      y = 'depth (m)'
    )
  
}

```

### Cluster mean by year

```{r cluster_mean_year}

if (opt_analysis_type == 2) {

  # cluster means by year
  anomaly_cluster_mean_year %>%
    mutate(year = as.factor(year)) %>%
    ggplot() +
    geom_path(aes(x = anomaly_mean,
                  y = depth,
                  col = year)) +
    geom_vline(xintercept = 0) +
    scale_y_reverse() +
    facet_wrap( ~ cluster) +
    coord_cartesian(xlim = opt_xlim) +
    scale_x_continuous(breaks = opt_xbreaks) +
    scale_color_viridis_d() +
    labs(
      title = paste0('Overall mean anomaly profiles by cluster'),
      x = opt_measure_label,
      y = 'depth (m)'
    )
  
}

    
```

### Cluster by year

count of each cluster by year

```{r cluster_by_year}

if (opt_analysis_type == 2) {

  # Determine profile count by cluster and year
  cluster_by_year <- temp_anomaly_cluster %>% 
    count(cluster, year,
          name = "count_cluster")

  year_min <- min(cluster_by_year$year)
  year_max <- max(cluster_by_year$year)
  
  # create figure
  cluster_by_year %>% 
    ggplot(aes(x = year, y = count_cluster, col = cluster, group=cluster))+
    geom_point() +
    geom_line() +
    scale_x_continuous(breaks = seq(year_min, year_max, 2)) +
    scale_color_brewer(palette = 'Dark2')+
    labs(x = 'year', 
         y = 'number of profiles', 
         col = 'cluster',
         title = 'Count of profiles by year and cluster')

}

```

### Cluster by month

count of each cluster by month of year

```{r cluster_by_month}

if (opt_analysis_type == 2) {

  # Determine profile count by cluster and year
  cluster_by_year <- temp_anomaly_cluster %>% 
    count(cluster, month,
          name = "count_cluster")

  # create figure
  cluster_by_year %>% 
    ggplot(aes(x = month, y = count_cluster, col = cluster, group=cluster))+
    geom_point() +
    geom_line() +
    scale_x_continuous(breaks = seq(1, 12, 2)) +
    scale_color_brewer(palette = 'Dark2')+
    labs(x = 'month', 
         y = 'number of profiles', 
         col = 'cluster',
         title = 'Count of profiles by month and cluster')

}

```

### Cluster spatial

location of each cluster on map, spatial analysis

```{r spatial_cluster}

if (opt_analysis_type == 2) {

  # create figure
map +
  geom_tile(data = temp_anomaly_cluster, 
            aes(x = lon, 
                y = lat, 
                fill = cluster))+
  lims(y = c(-85, -30))+
  scale_fill_brewer(palette = 'Dark2')+
  labs(title = 'cluster spatial distribution')

  # create figure
map +
  geom_tile(data = temp_anomaly_cluster,
            aes(x = lon,
                y = lat,
                fill = cluster)) +
  lims(y = c(-85, -30)) +
  scale_fill_brewer(palette = 'Dark2') +
  facet_wrap( ~ cluster, ncol = 2) +
  labs(title = 'cluster spatial distribution')

  # create figure
map +
  geom_tile(data = temp_anomaly_cluster %>% 
              count(lat, lon, cluster),
            aes(x = lon,
                y = lat,
                fill = n)) +
  scale_fill_viridis_c(option = "cividis", direction = -1, trans = "log10") +
  facet_wrap( ~ cluster, ncol = 2) +
  labs(title = 'cluster spatial distribution')

}

```

### Cluster spatial year

location of each cluster on map, spatial analysis by year

```{r spatial_cluster_year}

if (opt_analysis_type == 2) {

  # create figure
  temp_anomaly_cluster %>%
    group_split(year) %>%
    map(
      ~ map +
        geom_tile(data = .x,
                  aes(
                    x = lon,
                    y = lat,
                    fill = cluster
                  )) +
        scale_fill_brewer(palette = 'Dark2') +
        facet_wrap( ~ cluster, ncol = 2) +
        labs(title = paste0(
          'cluster spatial distribution ', unique(.x$year)
        ))
    )

}

```

## Cluster by surface Extreme

```{r cluster_analysis_prep}

# Carried out for 1500m profiles
opt_profile_range = 3
extreme_type <- c('L', 'N', 'H')
num_clusters <- c(5, 6, 5)

# ---------------------------------------------------------------------------------------------
# read data
# ---------------------------------------------------------------------------------------------
# load previously created OceanSODA extreme data. date, position and nature of extreme
temp_extreme <- read_rds(file = paste0(path_argo_preprocessed, "/OceanSODA_global_SST_anomaly_field.rds")) %>%
  select(lon, lat, date, temp_extreme)

# read data
temp_anomaly_va <- read_rds(file = paste0(path_core_preprocessed, "/temp_anomaly_va.rds")) %>%
  mutate(date = ymd(format(date, "%Y-%m-15")))

# Add the OceanSODA extreme condition
temp_anomaly_va <- left_join(temp_anomaly_va, temp_extreme)

for (i in 1:3) {

  # ---------------------------------------------------------------------------------------------
  # Preparation
  # ---------------------------------------------------------------------------------------------
  # select profile based on profile_range and he appropriate max depth
  temp_anomaly_va_id <- temp_anomaly_va %>% 
    filter(profile_range == opt_profile_range & depth <= opt_max_depth[opt_profile_range] & temp_extreme == extreme_type[i])
  
  # Simplified table ready to pivot
  temp_anomaly_va_id <- temp_anomaly_va_id %>%
    select(file_id,
           depth,
           anomaly,
           year, 
           month, 
           lat, 
           lon)
  
  # wide table with each depth becoming a column
  temp_anomaly_va_wide <- temp_anomaly_va_id %>%
    select(file_id, depth, anomaly) %>%
    pivot_wider(names_from = depth, values_from = anomaly)
  
  # Drop any rows with missing values N/A caused by gaps in climatology data
  temp_anomaly_va_wide <- temp_anomaly_va_wide %>% 
    drop_na()
  
  # Table for cluster analysis
  points <- temp_anomaly_va_wide %>%
    column_to_rownames(var = "file_id")
  
  # ---------------------------------------------------------------------------------------------
  # cluster analysis
  # ---------------------------------------------------------------------------------------------
  set.seed(1)
  kclusts <-
    tibble(k = num_clusters[i]) %>%
    mutate(
      kclust = map(k, ~ kmeans(points, .x, iter.max = opt_max_iterations, nstart = opt_n_start)),
      tidied = map(kclust, tidy),
      glanced = map(kclust, glance),
      augmented = map(kclust, augment, points)
    )
  
  profile_id <-
    kclusts %>%
    unnest(cols = c(augmented)) %>%
    select(file_id = .rownames,
           cluster = .cluster) %>% 
    mutate(file_id = as.numeric(file_id),
           cluster = as.character(cluster))
  
  # Add cluster to temp_anomaly_va
  temp_anomaly_cluster <- full_join(temp_anomaly_va, profile_id)
  
  # Plot cluster mean
  temp_anomaly_cluster <- temp_anomaly_cluster %>% 
    filter(!is.na(cluster))
  
  # Plot cluster mean
  anomaly_cluster_mean <- temp_anomaly_cluster %>%
    group_by(cluster, depth) %>%
    summarise(
      count_cluster = n(),
      anomaly_mean = mean(anomaly, na.rm = TRUE),
      anomaly_sd = sd(anomaly, na.rm = TRUE)
    ) %>%
    ungroup()
  
  anomaly_cluster_mean_year <- temp_anomaly_cluster %>%
    group_by(cluster, depth, year) %>%
    summarise(
      count_cluster = n(),
      anomaly_mean = mean(anomaly, na.rm = TRUE),
      anomaly_sd = sd(anomaly, na.rm = TRUE)
    ) %>%
    ungroup()
  
  anomaly_year_mean <- temp_anomaly_cluster %>%
    group_by(cluster, year) %>%
    summarise(
      count_cluster = n(),
      anomaly_mean = mean(anomaly, na.rm = TRUE),
      anomaly_sd = sd(anomaly, na.rm = TRUE)
    ) %>%
    ungroup()
  
  anomaly_year_mean <- anomaly_year_mean %>%
    group_by(year) %>%
    summarise(anomaly_mean = mean(anomaly_mean, na.rm = TRUE)) %>%
    ungroup ()
  
  if (i == 1) {
    anomaly_cluster_mean_ext <- anomaly_cluster_mean %>% mutate(temp_extreme_order = i, temp_extreme = extreme_type[i])
    anomaly_cluster_mean_year_ext <- anomaly_cluster_mean_year %>% mutate(temp_extreme_order = i, temp_extreme = extreme_type[i])
    anomaly_year_mean_ext <- anomaly_year_mean %>% mutate(temp_extreme_order = i, temp_extreme = extreme_type[i])
    temp_anomaly_cluster_ext <- temp_anomaly_cluster %>% mutate(temp_extreme_order = i, temp_extreme = extreme_type[i])
  } else {
    anomaly_cluster_mean_ext <- rbind(anomaly_cluster_mean_ext, anomaly_cluster_mean %>% mutate(temp_extreme_order = i, temp_extreme = extreme_type[i]))
    anomaly_cluster_mean_year_ext <- rbind(anomaly_cluster_mean_year_ext, anomaly_cluster_mean_year %>% mutate(temp_extreme_order = i, temp_extreme = extreme_type[i]))
    anomaly_year_mean_ext <- rbind(anomaly_year_mean_ext, anomaly_year_mean %>% mutate(temp_extreme_order = i, temp_extreme = extreme_type[i]))
    temp_anomaly_cluster_ext <- rbind(temp_anomaly_cluster_ext, temp_anomaly_cluster_ext <- temp_anomaly_cluster %>% mutate(temp_extreme_order = i, temp_extreme = extreme_type[i]))
  }

}

```

### Cluster mean

```{r cluster_mean_extreme}

# create figure of mean clusters
anomaly_cluster_mean_ext %>%
  group_split(temp_extreme_order) %>%
  map(
    ~  ggplot(data = .x, ) +
      geom_path(aes(x = anomaly_mean,
                    y = depth)) +
      geom_ribbon(
        aes(
          xmax = anomaly_mean + anomaly_sd,
          xmin = anomaly_mean - anomaly_sd,
          y = depth
        ),
        alpha = 0.2
      ) +
      geom_vline(xintercept = 0) +
      scale_y_reverse() +
      facet_wrap(~ cluster) +
      coord_cartesian(xlim = opt_xlim) +
      scale_x_continuous(breaks = opt_xbreaks) +
      labs(
        title = paste0(
          'Overall mean anomaly profiles by cluster \n',
          'Surface extreme: ',
          unique(.x$temp_extreme)
        ),
        x = opt_measure_label,
        y = 'depth (m)'
      )
  )


```

### Clusters mean by year

```{r cluster_mean_extreme_year}

# cluster means by year
anomaly_cluster_mean_year_ext %>%
  mutate(year = as.factor(year)) %>%
  group_split(temp_extreme_order) %>%
  map(
    ~  ggplot(data = .x,) +
      geom_path(aes(
        x = anomaly_mean,
        y = depth,
        col = year
      )) +
      geom_vline(xintercept = 0) +
      scale_y_reverse() +
      facet_wrap(~ cluster) +
      coord_cartesian(xlim = opt_xlim) +
      scale_x_continuous(breaks = opt_xbreaks) +
      scale_color_viridis_d() +
      labs(
        title = paste0(
          'Overall mean anomaly profiles by cluster by year \n',
          'Surface extreme: ',
          unique(.x$temp_extreme)
        ),
        x = opt_measure_label,
        y = 'depth (m)'
      )
  )

```

### Cluster by year

count of each cluster by year

```{r cluster_by_year}

# Determine profile count by extreme and cluster and year
# Count the measurements
cluster_by_year <- temp_anomaly_cluster_ext %>% 
  count(file_id, temp_extreme_order, temp_extreme, cluster, year,
        name = "count_cluster")
# Convert to profiles
cluster_by_year <- cluster_by_year %>% 
  count(temp_extreme_order, temp_extreme, cluster, year,
        name = "count_cluster")

year_min <- min(cluster_by_year$year)
year_max <- max(cluster_by_year$year)

# create figure
cluster_by_year %>%
  group_split(temp_extreme_order) %>%
  map(
    ~ ggplot(
      data = .x,
      aes(
        x = year,
        y = count_cluster,
        col = cluster,
        group = cluster
      )
    ) +
      geom_point() +
      geom_line() +
      scale_x_continuous(breaks = seq(year_min, year_max, 2)) +
      scale_color_brewer(palette = 'Dark2') +
      labs(
        x = 'year',
        y = 'number of profiles',
        col = 'cluster',
        title = paste0(
          'Count of profiles by year and cluster \n',
          'Surface extreme: ',
          unique(.x$temp_extreme)
        )
      )
  )

```

### Cluster by month

count of each cluster by month of year

```{r cluster_by_month}

# Determine profile count by cluster and year
# Count the measurements
cluster_by_year <- temp_anomaly_cluster_ext %>% 
  count(file_id, temp_extreme_order, temp_extreme, cluster, month,
        name = "count_cluster")
# Convert to profiles
cluster_by_year <- cluster_by_year %>% 
  count(temp_extreme_order, temp_extreme, cluster, month,
        name = "count_cluster")

# create figure
cluster_by_year %>%
  group_split(temp_extreme_order) %>%
  map(
    ~ ggplot(
      data = .x,
      aes(
        x = month,
        y = count_cluster,
        col = cluster,
        group = cluster
      )
    ) +
      geom_point() +
      geom_line() +
      scale_x_continuous(breaks = seq(1, 12, 2)) +
      scale_color_brewer(palette = 'Dark2') +
      labs(
        x = 'month',
        y = 'number of profiles',
        col = 'cluster',
        title = paste0(
          'Count of profiles by month and cluster \n',
          'Surface extreme: ',
          unique(.x$temp_extreme)
        )
      )
  )


```

### Cluster spatial

location of each cluster on map, spatial analysis

```{r spatial_cluster_combined}

# create figure combined
temp_anomaly_cluster_ext %>%
  group_split(temp_extreme_order) %>%
  map(
    ~ map +
      geom_tile(data = .x,
                aes(
                  x = lon,
                  y = lat,
                  fill = cluster
                )) +
      scale_fill_brewer(palette = 'Dark2') +
      labs(
        title = paste0(
          'cluster spatial distribution \n',
          'Surface extreme: ',
          unique(.x$temp_extreme)
        )
      )
  )

```

### Spatial by cluster

location of each cluster on map, spatial analysis

```{r spatial_cluster}

# create figure by cluster
temp_anomaly_cluster_ext %>%
  group_split(temp_extreme_order) %>%
  map(
    ~ map +
      geom_tile(data = .x,
                aes(
                  x = lon,
                  y = lat,
                  fill = cluster
                )) +
      scale_fill_brewer(palette = 'Dark2') +
      facet_wrap( ~ cluster, ncol = 2) +
      labs(
        title = paste0(
          'cluster spatial distribution \n',
          'Surface extreme: ',
          unique(.x$temp_extreme)
        )
      )
  )

```

### Spatial count

location of each cluster on map, spatial analysis

```{r spatial_cluster}

# create figure
temp_anomaly_cluster_ext %>%
  group_split(temp_extreme_order) %>%
  map(
    ~ map +
      geom_tile(data = .x %>%
                  count(lat, lon, cluster),
                aes(
                  x = lon,
                  y = lat,
                  fill = n
                )) +
      scale_fill_viridis_c(
        option = "cividis",
        direction = -1,
        trans = "log10"
      ) +
      facet_wrap(~ cluster, ncol = 2) +
      labs(
        title = paste0(
          'cluster spatial distribution \n',
          'Surface extreme: ',
          unique(.x$temp_extreme)
        )
      )
  )


```

